### 2018-07-16 data-science-math/Homework 1.Rmd
Initial commit

---
title: "CUNY SPSS Data Science Math Homework 1"
author: "Mike Silva"
date: "July 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Profit is the difference between revenue and cost.  Revenue is the quantity sold (q) multiplied by the price(p) and the costs is given at $10 times the quantity sold, thus
$$P(q,p) = R(q,p) - C(q) = qp - 10q$$
The quantity sold is given as a function of the price
$$q(p) = 80-2p$$
So rewriting the equation for profit as a function of the price
$$P(p) = R(p) - C(p) = (80-2p)p - 10(80-2p) = (80p-2p^2) - (800-20p) = -2p^2 + 100 - 800 $$
The retailer's profit function is
$$P(p) = -2p^2 + 100 - 800 $$

## Question 2
Given $f(x)=8x^3+7x^2-5$:

### Step 1
$$f(3) = 8(3)^3+7(3)^2-5 = (8*27)+(7*9)-5 = 216+63-5 = 274$$

### Step 2
$$f(-1) = 8(-2)^3+7(-2)^2-5 = (8*-8)+(7*4)-5 = -64+28-5 = -41$$

### Step 3
$$f(x+c) = 8(x+c)^3+7(x+c)^2-5$$