### 2018-07-17 data-science-math/Homework 1.Rmd
Additional work

---
title: "CUNY SPSS Data Science Math Homework 1"
author: "Mike Silva"
date: "July 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(Deriv)
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1**

Profit is the difference between revenue and cost.  Revenue is the quantity sold (q) multiplied by the price(p) and the costs is given at $10 times the quantity sold, thus
$$P(q,p) = R(q,p) - C(q) = qp - 10q$$
The quantity sold is given as a function of the price
$$q(p) = 80-2p$$
So rewriting the equation for profit as a function of the price
$$P(p) = R(p) - C(p)$$
$$ (80-2p)p - 10(80-2p)$$
$$ (80p-2p^2) - (800-20p)$$
$$-2p^2 + 100p - 800 $$

**Question 2**

Given $f(x)=8x^3+7x^2-5$:

**Step 1**
$$f(3) = 8(3)^3+7(3)^2-5 = (8\cdot27)+(7\cdot9)-5 = 216+63-5 = 274$$

**Step 2**
$$f(-1) = 8(-2)^3+7(-2)^2-5 = (8\cdot-8)+(7\cdot4)-5 = -64+28-5 = -41$$

**Step 3**
$$f(x+c) = 8(x+c)^3+7(x+c)^2-5$$
Verifying my math with R:
```{r}
f2 <- function(x){8*x**3+7*x**2-5}
f2(3)
f2(-2)
```

**Question 3 **

**Step 1**
$$\lim_{x\to1^-} f(x) = 2$$

**Step 2**
$$\lim_{x\to1^+} f(x) = -5$$

**Step 3**
$$\lim_{x\to1} f(x) = \mbox{ Does not exist} $$

**Question 4**
Find the derivative of $f(x)=-2x^3$

$$f'(x)=-6x^2$$
Verifying my math in R:
```{r}
f4 <- function(x){-2*x**3}
Deriv(f4)
```

**Question 5**
$$f(x) = \frac{-8}{x^2}$$
$$f'(x) = \frac{16}{x^3}$$
Verify with R:
```{r}
f5 <-function(x){-8/(x**2)}
Deriv(f5)
```

**Question 6**
$$g(x) 5\sqrt[3]{x}$$

$$g'(x) \frac{5}{3x^\frac{2}{3}}=\frac{5}{3}\cdot\frac{1}{x^\frac{2}{3}}$$
```{r}
f6 <- function(x){5*x**(1/3)}
Deriv(f6)

5/3

2/3
```

**Question 7**

$$y=-2x^\frac{9}{8}$$

$$y'=\frac{-9x^\frac{1}{8}}{4}=\frac{-9}{4}\cdot x^\frac{1}{8}$$
```{r}
f7 <- function(x){-2*x**(9/8)}
Deriv(f7)

-9/4
1/8
```

**Question 8**
$$\frac{35-40}{4-0}=\frac{-5}{4} = -1\frac{1}{4}$$

**Question 9**
$$C(x)=630+2.4x$$
$$\mbox{Avg Cost} = \frac{\mbox{Cost}}{\mbox{Unit}}=\frac{C(x)}{x}=\frac{630+2.4x}{x}$$

**Question 10**
$$f(x)=(-2x^{-2} +1)(-5x + 9)$$

Applying the quotient rule where $f(x)=-2x^{-2}+1$ and $g(x)=(-5x+9)$

$$f'(x) = \frac{4}{x^3}(-5x+9) + -5(2x^{-2} + 1) = \frac{36}{x^3} - \frac{20}{x^2} + \frac{10}{x^2} - 5=\frac{36}{x^3}-\frac{10}{x^2}-5$$
Checking my math with R:
```{r}
f10 <- function(x){(-2*x**-2+1)*(-5*x+9)}
Deriv(f10)
# Evaluating at 10
f10_prime <- function(x){(36/x**3)-(10/x**2)-5}
f10_prime(10)
f10_prime <- Deriv(f10)
f10_prime(10)
```

**Question 11**
$$f(x)=\frac{5x^\frac{1}{2}+7}{x^3 + 1}$$

Applying the quotient rule where $f(x)=5\sqrt{x}$ and $g(x)=x^3+1$

$$f'(x) = \frac{\Big(\frac{5}{2\sqrt{x}}\Big)(x^3+1)-(3x^2(5x^\frac{1}{2}+7))}{(x^3+1)^2}$$
Simplified to 
$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^6+2x^3 + 1)}$$
Again checking with R:
```{r}
f11 <- function(x){(7+5*x**(1/2))/(1+x**3)}
Deriv(f11)
# Evaluate at 5
fprime11 <- function(x){(-42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*(x**6+2*x**3+1))}
fprime11(5)
fprime11 <- Deriv(f11)
fprime11(5)
```

**Question 12**
$$f(x)=(3x^{-3}-8x+6)^\frac{4}{3}$$

$$f'(x) = \frac{4}{3}(-\frac{9}{x^4}-8) \Big(\frac{3}{x^3}-8x+6\Big)^\frac{1}{3}$$

Checking my math in R:
```{r}
f12 <- function(x){(3*x**-3-(8*x)+6)**(4/3)}
Deriv(f12)
# Evaluate at 1
f12_prime <- function(x){(4/3)*((-9/x**4)-8)*(((3/x**3)-8*x+6)**(1/3))}
f12_prime(1)
f12_prime <- Deriv(f12)
f12_prime(1)
```

*Question 13*
$$f(t) =\frac{550t^2}{\sqrt{t^2+15}}$$

$$f'(t)=$$
```{r}
f13 <- function(t){(550*t**2)/(t+15)**(1/2)}
Deriv(f13)
```