### 2018-07-19 data-science-math/Homework 1.Rmd
More work

---
title: "CUNY SPSS Data Science Math Homework 1"
author: "Mike Silva"
date: "July 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(Deriv)
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1

Profit is the difference between revenue and cost.  Revenue is the quantity sold (q) multiplied by the price(p) and the costs is given at $10 times the quantity sold, thus
$$P(q,p) = R(q,p) - C(q) = qp - 10q$$
The quantity sold is given as a function of the price
$$q(p) = 80-2p$$
So rewriting the equation for profit as a function of the price
$$P(p) = R(p) - C(p)$$
$$ (80-2p)p - 10(80-2p)$$
$$ (80p-2p^2) - (800-20p)$$
$$-2p^2 + 100p - 800 $$

******

### Question 2

Given $f(x)=8x^3+7x^2-5$:

#### Step 1

$$f(3) = 8(3)^3+7(3)^2-5 = (8\cdot27)+(7\cdot9)-5 = 216+63-5 = 274$$

#### Step 2

$$f(-1) = 8(-2)^3+7(-2)^2-5 = (8\cdot-8)+(7\cdot4)-5 = -64+28-5 = -41$$

#### Step 3

$$f(x+c) = 8(x+c)^3+7(x+c)^2-5$$
Verifying my math with R:
```{r}
f2 <- function(x){8*x**3+7*x**2-5}
f2(3)
f2(-2)
```

******

### Question 3

#### Step 1

$$\lim_{x\to1^-} f(x) = 2$$

#### Step 2

$$\lim_{x\to1^+} f(x) = -5$$

#### Step 3

$$\lim_{x\to1} f(x) = \mbox{ Does not exist} $$

******

### Question 4

Find the derivative of $f(x)=-2x^3$

$$f'(x)=-6x^2$$
Verifying my math in R:
```{r}
f4 <- function(x){-2*x**3}
Deriv(f4)
```

******

### Question 5

$$f(x) = \frac{-8}{x^2}$$
$$f'(x) = \frac{16}{x^3}$$

Verify with R:
```{r}
f5 <-function(x){-8/(x**2)}
Deriv(f5)
```

******

### Question 6
$$g(x) 5\sqrt[3]{x}$$

$$g'(x) \frac{5}{3x^\frac{2}{3}}=\frac{5}{3}\cdot\frac{1}{x^\frac{2}{3}}$$

Verify with R:
```{r}
f6 <- function(x){5*x**(1/3)}
Deriv(f6)

5/3

2/3
```

******

### Question 7

$$y=-2x^\frac{9}{8}$$

$$y'=\frac{-9x^\frac{1}{8}}{4}=\frac{-9}{4}\cdot x^\frac{1}{8}$$

Verify with R:
```{r}
f7 <- function(x){-2*x**(9/8)}
Deriv(f7)

-9/4
1/8
```

******

### Question 8

$$\frac{35-40}{4-0}=\frac{-5}{4} = -1\frac{1}{4}$$

******

### Question 9

$$C(x)=630+2.4x$$
$$\mbox{Avg Cost} = \frac{\mbox{Cost}}{\mbox{Unit}}=\frac{C(x)}{x}=\frac{630+2.4x}{x}$$

******

### Question 10

$$f(x)=(-2x^{-2} +1)(-5x + 9)$$

Applying the quotient rule where $f(x)=-2x^{-2}+1$ and $g(x)=(-5x+9)$

$$f'(x) = \frac{4}{x^3}(-5x+9) + -5(2x^{-2} + 1) = \frac{36}{x^3} - \frac{20}{x^2} + \frac{10}{x^2} - 5=\frac{36}{x^3}-\frac{10}{x^2}-5$$

Verify with R:
```{r}
f10 <- function(x){(-2*x**-2+1)*(-5*x+9)}
Deriv(f10)
# Evaluating at 10
f10_prime <- function(x){(36/x**3)-(10/x**2)-5}
f10_prime(10)
f10_prime <- Deriv(f10)
f10_prime(10)
```

******

### Question 11

$$f(x)=\frac{5x^\frac{1}{2}+7}{x^3 + 1}$$

Applying the quotient rule where $f(x)=5\sqrt{x}+7$ and $g(x)=x^3+1$ so $f'(x)=\frac{5}{2\sqrt{x}}$ and $g'(x)=3x^2$

$$f'(x) = \frac{\Big(\frac{5}{2\sqrt{x}}\Big)(x^3+1)-(3x^2(5\sqrt{x}+7))}{(x^3+1)^2}$$
$$f'(x) = \frac{\frac{5x^3+5}{2\sqrt{x}}-3x^2 (5\sqrt{x}+7)}{(x^3+1)^2}$$

Simplified to 
$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^3 + 1)^2}$$

Again checking with R:
```{r}
f11 <- function(x){(5*x**(1/2)+7)/(x**3+1)}
Deriv(f11)
# Evaluate at 5
fprime11 <- function(x){-1*(42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*((x**3+1)**2))}
fprime11(5)
fprime11 <- Deriv(f11)
fprime11(5)
```

******

### Question 12

$$f(x)=(3x^{-3}-8x+6)^\frac{4}{3}$$

$$f'(x) = \frac{4}{3}(-\frac{9}{x^4}-8) \Big(\frac{3}{x^3}-8x+6\Big)^\frac{1}{3}$$

Checking my math in R:
```{r}
f12 <- function(x){(3*x**-3-(8*x)+6)**(4/3)}
Deriv(f12)
# Evaluate at 1
f12_prime <- function(x){(4/3)*((-9/x**4)-8)*(((3/x**3)-8*x+6)**(1/3))}
f12_prime(1)
f12_prime <- Deriv(f12)
f12_prime(1)
```

******

### Question 13

$$f(t) =\frac{550t^2}{\sqrt{t^2+15}}$$

After factoring out the 550 constant and applying the quotient rule where $f(x)=t^2$ and $g(x)=\sqrt{t^2+15}$

$$550\cdot f'(x) = \frac{(2t)(\sqrt{t^2+15})-(t^2)\Big(\frac{t}{\sqrt{t^2+15}}\Big)}{t^2+15} = \frac{2t\sqrt{t^2+15} -\frac{t^3}{\sqrt{t^2+15}}}{t^2+15}$$

Multiplying by 550 and simplifying
$$f'(x) = \frac{1100t\sqrt{t^2+15} -\frac{550t^3}{\sqrt{t^2+15}}}{t^2+15}=\frac{550t(t^2+30)}{(t^2+15)^\frac{3}{2}}$$

Verifying with R:
```{r}
f13 <- function(t){(550*t**2)/((t**2+15)**(1/2))}
Deriv(f13)
f13_prime <- function(t){(550*t*(t**2+30))/((t**2+15)**(3/2))}
f13_prime(3)
f13_prime <- Deriv(f13)
f13_prime(3)
```

The answer is 547

******

### Question 14

$$N(t) = 1000(6+0.1t)^\frac{1}{2}=1000\sqrt{0.1t+6}$$

#### Step 1
$$N(3) = 1000(6+0.1(3))^\frac{1}{2}=39690$$

#### Step 2
Let $u=0.1t+6$ so $\sqrt{u}$ becomes $\frac{1}{2\sqrt{u}}$ because of the power rule

and due to the chain rule $u'=0.1$.  Substitute

$$1000\cdot N'(t) = \frac{0.1\cdot0.5}{\sqrt{0.1t+6}}=\frac{0.05}{\sqrt{0.1t+6}}$$
$$ N'(t)=\frac{1000\cdot 0.05}{\sqrt{0.1t+6}}=\frac{50}{\sqrt{0.1t+6}}$$
$$N'(5)=\frac{50}{\sqrt{0.1(5)+6}}=\frac{50}{\sqrt{6.5}}=19.6116135138\approx20	$$
20 is roughly how many additional people attend in the 5th week of the season

Checking my math with R:
```{r}
f14 <- function(t){1000*(6+0.1*t)**(1/2)}
Deriv(f14)
f14_prime <- function(t){50/(6.5**(1/2))}
f14_prime(5)
f14_prime <- Deriv(f14)
f14_prime(5)
```

******

### Question 15

$$3x^3+4y^3=77$$

#### Step 1

$$\frac{d}{dx}(3x^3+4y^3)=\frac{d}{dx}(77)$$

$$\frac{d}{dx}(3x^3)+\frac{d}{dx}(4y^3)=0$$
$$9x^2+12y^2\frac{dy}{dx}=0$$

$$12y^2\frac{dy}{dx}=-9x^2$$

$$\frac{dy}{dx}=\frac{-9x^2}{12y^2} = -\frac{3x^2}{4y^2}$$

#### Step 2

The slope of the tangent line at (3,-1) is

$$-\frac{3(3)^2}{4(-1)^2}=-\frac{3(9)}{4(1)}=-\frac{7}{4}$$

******

### Question 16

$$f(x)=\frac{x+3}{x-8}$$

Differentiate using the quotient rule where $f(x)=x+3$ and $g(x)=x-8$

$$f'(x)=\frac{(1)(x-8)-(1)(x+3)}{(x-8)^2}=\frac{(x-8)-(x+3)}{(x-8)^2}=\frac{x-8-x-3}{(x-8)^2}=-\frac{11}{(x-8)^2}$$

Checking my work with R:
```{r}
f16 <- function(x){(x+3)/(x-8)}
Deriv(f16)
f16_prime <- function(x){-11/((x-8)^2)}
f16_prime(20)
f16_prime <- Deriv(f16)
f16_prime(20)
```

$$f''(x)= \frac{22}{(x-8)^3}$$

The critical point is when x=8.  f(x) is decreasing when x < 8 and increasing when x > 8.

******

### Question 17

$$F(t) = 14 + \frac{367t^2}{t^2+100}$$

#### Step 1

$$367 \cdot F'(t) = \frac{t^2}{(t^2+100)^2}=\frac{(2t(t^2+100))(-2t^2\cdot t)}{(t^2+100)^2}=\frac{2t(t^2+100)(-2t^3)}{(t^2+100)^2}$$

Multiply it by 367

$$f'(x)=367 \cdot \frac{2t(t^2+100)(-2t^3)}{(t^2+100)^2}=\frac{734t(t^2+100)-734t^3}{(t^2+100)^2}=\frac{73400t}{(t^2+100)^2}$$

In the interest of turning my assignment on time some of the steps have been excluded

$$f''(x)=-\frac{73400(3t^2-100)}{(t^2+100)^3}$$

$F'(t) > 0$ from $(0,\infty)$ and $F'(t) < 0$ from $(-\infty, 0)$ which doesn't make sense in this case.

#### Step 2

$$\lim_{t\to\infty} F(t) = 381^\circ F$$

******

### Question 18

$$f(x)=\frac{x^3}{3340000}-\frac{7x^2}{9475}+\frac{42417727x}{1265860000}+\frac{1}{33}$$

Again in the interest of time I will skip the steps of the calculation:

$$f'(x) = \frac{3x^2}{3340000}-\frac{14x}{9475}+\frac{42417727}{1265860000}$$

$$0 = \frac{3x^2}{3340000}-\frac{14x}{9475}+\frac{42417727}{1265860000}$$

```{r}
f18 <- function(x){}
f18_prime <- function(x){(3*x**2/3340000)-(14*x/9475)+(42417727/1265860000)}
x <- seq(0, 30, by=.01)
y <- f18_prime(x)
i <- stats::uniroot(f18_prime,c(-30,30))
plot(y~x,type="l", col="blue")
abline(h=0, col="black")
abline(v=0, col="black")
abline(v=i$root, col="red")
points(i$root, 0, col="red", pch=21)
```

```{r}
i$root
```

******

### Question 19

$$f(x) = 7x^2+28x-35$$

$$f'(x) = 14x+28$$

$$0 = 14x+28 $$

$$-28 = 14x$$

$$x=-\frac{28}{14}=-2$$

$$f(-2) = (7(-2)^2)+(28(-2))-35 = 28-56-35=-63$$

$$\mbox minimum = (-2,-63)$$

Checking with R:
```{r}
f19 <- function(x){7*x**2+28*x-35}
f19_prime <- Deriv(f19)
i <- stats::uniroot(f19_prime,c(-30,30))
i$root
f19(i$root)
```

```{r}
x <- seq(-30, 30, by=.01)
y <- f19(x)
i <- stats::uniroot(f19_prime,c(-30,30))
plot(y~x,type="l", col="blue")
abline(h=0, col="black")
abline(v=0, col="black")
points(i$root, 0, col="red", pch=21)
```

******

### Question 20

$$f(x)=-6x^3+27x^2+180x$$

$$f'(x)=-18x^2+54x+180$$

$$0=-18x^2+54x+180$$

$$0=-18(x^2-3x-10)$$

$$x^2-3x-10=0$$

$$(x+2)(x-5)=0$$

The function is minimized/maximized when $x = -2$ or $x = 5$.  Plugging in and solving for y:

$$f(-2)=-6(-2)^3+27(-2)^2+180(-2)=48+108-360=-204$$
$$f(-2)=-6(5)^3+27(5)^2+180(5)=-750+650+900=800$$

$$f''(x)=-36x+54$$
$$f''(-2)=-36(-2)+54=72+54=126>0$$
$$f''(5)=-36(5)+54=180+54=234>0$$

So the critical points are $(-2, -204)$ and $(5, 800)$.  $(-2, -204)$ is a local min and $(5, 800)$

```{r}
f20 <- function(x){(-6*x**3)+(27*x**2)+(180*x)}
f20_prime <- Deriv(f20)
x <- seq(-30, 30, by=.01)
y <- f20(x)
#i <- stats::uniroot(f20_prime,c(-30,30))
plot(y~x,type="l", col="blue")
abline(h=0, col="black")
abline(v=0, col="black")
#points(i$root, 0, col="red", pch=21)
```

******

### Question 21
Inventory cost is defined as the cost of holding stock and the cost of ordering.  The inventory cost as a function of the iron lot size is:

$$C(i)=\frac{120}{i}(6+4.5(i))+(1.60\frac{i}{2})$$
This function assumes the irons are selling at a constant rate (no seasonality) throughout the year.  Simplifing the function:

$$C(i)=\frac{620}{i}+ \frac{540i}{i}+(0.8i) = \frac{620}{i}+0.8i+540$$

Since we want to minimize this cost we take the first derivative and set it equal to zero.

$$C'(i)=-\frac{720}{i^2} +0.8$$

$$0=-\frac{720}{i^2} +0.8$$
$$\frac{720}{i^2}=0.8$$
$$720=0.8i^2$$
$$i^2=\frac{720}{0.8}=900$$
$$i=\sqrt{900}=30$$
So the **lot size is 30 irons**.  Since we are expecting to sell 120 irons per year this would require **4 orders**.

******

### Question 22

$$Volume = 18432ft^2 = x\cdot x \cdot y \Rightarrow y=\frac{18432}{x^2}$$
The cost is the sum of the sides, top and bottom.  The sides and top material cost $3.00 per square foot and the bottom is $5.00.

$$C(x,y) = 3(4xy) + 3x^2 + 5x^2 = 12xy+8x^2$$
Substituting in y:
$$ C(x)=\frac{12x\cdot 18432}{x^2}+8x^2=\frac{221184}{x}+8x^2$$
Minimize the cost function:
$$ C'(x)=-\frac{221184}{x^2}+16x$$
$$ 0=-\frac{221184}{x^2}+16x\Rightarrow 0=-221184+16x^3 \Rightarrow 0=-13824 + x^3 $$
$$ x^3 = 13824 \Rightarrow x=\sqrt[3]{13824} = 24$$
Plugging 24 into the volume equation and solving for y:
$$18432 = x^2y\Rightarrow 18432=(24)^2 y\Rightarrow 18432=576 y \Rightarrow y=\frac{18432}{576} \Rightarrow y = 32$$
Now checking everything with R:
```{r}
f22 <- function(x){
  (3*(4*x*(18432/(x**2)))) + (3*x**2) + (5*x**2)
}
f22_prime <- Deriv(f22)
library(rootSolve)
x22 <- uniroot.all(f22_prime, c(0,100))
x22
y22 <- function(x){18432/(x**2)}
y22(x22)
```

******

### Question 23

******

### Question 24

******

### Question 25

******

### Question 26

******

### Question 27

******

### Question 28

******

### Question 29

******
### Question 30

$$ \int_{-7}^2 x\sqrt{x+7}dx$$

For this problem $u=x$ and $dv=\sqrt{x+7}dx$.  Calculate $du$ and $v$:

$du=1dx$ and $v = \int \sqrt{x+7}dx = \frac{2}{3}(x+7)^\frac{3}{2}$

Plugging it into the integration by parts formula:

$$ \int_a^b udv = uv]_a^b -\int_a^b vdu$$

$$ \int_{-7}^2x\sqrt{x+7}dx]_{-7}^2 -\int_{-7}^2 \frac{2}{3}(x+7)^\frac{3}{2}\cdot 1dx$$
Integrate the right hand side

$$ \frac{2x}{3}(x+7)^\frac{3}{2}]_{-7}^2 -\frac{4}{15}(x+7)^\frac{5}{2}]_{-7}^2 $$

Plugging in the -7 and 2 yeilds -28.8 as shown below:
```{r}
f30 <- function(a, b){
  left <- function(x){
    ((2*x)/3)*(x+7)**(3/2)
  }
  right <- function(y){
    (4/15)*(y+7)**(5/2)
  }
  (left(b)-left(a))-(right(b)-right(a))
}

f30(-7,2)
```
******

### Question 31

#### Step 1
In the interest of time I will solve this question by building a simple model
```{r}
bounce <- function(total_times, drop_height, debug=F){
  rebound = .22 # Amount of the rebound as a share of drop height
  times_ball_hit_the_ground = 1
  total_distance = drop_height
  if(debug){
    message(paste(total_distance, 'meters after ball hits the ground',times_ball_hit_the_ground,'times'))
  }
  while(times_ball_hit_the_ground < total_times){
    drop_height = rebound * drop_height 
    total_distance = total_distance + (2 * drop_height) #going Up and down
    times_ball_hit_the_ground = times_ball_hit_the_ground + 1
    if(debug){message(paste(total_distance, 'meters after ball hits the ground',times_ball_hit_the_ground,'times'))}
  }
  if(!debug){
  message(paste(total_distance, 'meters after ball hits the ground',times_ball_hit_the_ground,'times'))
  }
}

bounce(20, 46, T)
```
So the ball traveled 71.94872 meters when it hit the ground on the 20th time.  I would note that the additional distance is very small after the 7th

#### Step 2
To evaluate the total distance as the number of bounces heads to infinity I will cheat and plug in a really large number and see what it tends toward.
```{r}
bounce(1000000, 46)
```

So it is tending toward 72 meters.

******

### Question 34