### 2018-08-28 DATA607/Week 1 Assignment/Week_1_Assignment.Rmd
More work to Week 1 Assignment

---
title: "CUNY SPS DATA-607 Week 1 Assignment"
subtitle: "Loading Data into a Data Frame"
author: "Mike Silva"
date: "August 27, 2018"
output:
  html_document:
    keep_md: yes
---

## Description

The assignment is to load the Mushroom Dataset from the [UCI repository](https://archive.ics.uci.edu/ml/datasets/Mushroom) into a data frame, rename the columns into something meaningful, replace the abbreviations used in the data, and subset the columns of the data frame.  It must include the column that indicated if the mushroom is edible or poisonous and three or four other columns.

## Step 1: Acquire the Data

The first step will be to download the data to the local environment:

```{r acquire-data}
download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data', 'agaricus-lepiota.data')
```

## Step 2: Load the Data into Data Frame

Now that we have the dataset locally we will create a data frame:

```{r load-data}
df <- read.table('agaricus-lepiota.data', sep=',', stringsAsFactors=FALSE)
head(df)
```

According to the documentation we should have 8,124 rows and 23 columns of data.

```{r data-size}
dim(df)
```

## Step 3: Rename the columns

Our data checks out, but the names are not very useful.

```{r bad-names}
names(df)
```

I will rename the columns by using the dplyr package:

```{r rename-columns, warning=FALSE}
library(dplyr)

df <- df %>%
  rename(type = V1,
         cap_shape = V2,
         cap_surface = V3,
         cap_color = V4,
         bruises = V5,
         odor = V6,
         gill_attachment = V7,
         gill_spacing = V8,
         gill_size = V9,
         gill_color = V10,
         stalk_shape = V11,
         stalk_root = V12,
         stalk_surface_above_ring = V13,
         stalk_surface_below_ring = V14,
         stalk_color_above_ring = V15,
         stalk_color_below_ring = V16,
         veil_type = V17,
         veil_color = V18,
         ring_number = V19,
         ring_type = V20,
         spore_print_color = V21,
         population = V22,
         habitat = V23)
```

## Step 4: Subsetting Data Frame

Before changing the data to be more meaningful I want to subset the data frame.  I would like to use this data to predict if a mushroom is poisonous or not.  According to the documentation the odor, spore print color, stalk surface below ring, and stalk color above ring are 4 attributes that can predict if a mushroom is poisonous with a 99.9% accuracy for these 23 species.

```{r subset-data}
df <- df %>%
  select(type, odor, spore_print_color, stalk_color_below_ring, stalk_color_above_ring)
```

Now that we have these subsets, let's examine the values:

```{r subset-tables}
lapply(df, function(x) table(x)) 
```

## Step 5: Recoding Variables

Now that we have the subset of columns we can recode the variables.  I will accomplish this with dplyr and recode statements.

```{r recode-variables}
df <- df %>%
  mutate(type = recode(type, p="Poisonous", e="Edible"),
         odor = recode(odor, a="Almond", c="Creosote", f="Foul", l="Anise", m="Musty", n="None", p="Pungent", s="Spicy", y="Fishy"),
         spore_print_color = recode(spore_print_color, b="Buff", h="Chocolate", k="Black", n="Brown", o="Orange", r="Green", u="Purple", w="White", y="Yellow"),
         stalk_color_below_ring = recode(stalk_color_below_ring, b="Buff", c="Cinnamon", e="Red", g="Gray", n="Brown", o="Orange", p="Pink", w="White", y="Yellow"),
         stalk_color_above_ring = recode(stalk_color_above_ring, b="Buff", c="Cinnamon", e="Red", g="Gray", n="Brown", o="Orange", p="Pink", w="White", y="Yellow"))
```

Now to examine the final product:

```{r final-product}
head(df)
tail(df)
```