### 2018-09-07 DATA607/Week_2_Assignment.Rmd
Initial commit of Week 2 Assignment R Markdown

---
title: "Week 2 Assignment"
author: "Mike Silva"
date: "September 1, 2018"
output: html_document
---

## Description

Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

The films used in this study were selected by Googling "2018 recent films" and getting the first 6 films listed in the search results.

## What's Been Done So Far

The 'Week 2 Assignment.sql' script has been executed which has set up the flix database with the ratings.

## Connect To The Database

Using settings in a hidden file, connect to the database.

```{r connect,message=FALSE}
library(RMySQL)
library(ggplot2)
library(dplyr)
source('hidden/MySQL Settings.R')
conn <- dbConnect(MySQL(), user=user, password=password, dbname='flix', host=host)
```

## Pull Data From The Database

Pull data via the connection.

```{r pull_data}
sql <- 'SELECT `films`.`title`, 
`reviewers`.`name`, 
`ratings`.`rating`
FROM `ratings`
JOIN `films` ON `films`.`id` = `ratings`.`films_id`
JOIN `reviewers` ON `reviewers`.`id` = `ratings`.`reviewers_id`'
df <- dbGetQuery(conn, sql)

# Disconnect from the database
disconnected <- dbDisconnect(conn)
```

## Exploratory Data Analysis

Now that the data is loaded into a dataframe let's take a look at the data.

```{r first_look}
dim(df)
names(df)
```

Now for a preview of the data

```{r preview}
head(df)
```

```{r viz}
ggplot(df, aes(x=title, y=rating, fill=title)) + 
  geom_boxplot() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

The data is centered arround 3 with the exception of Annihilaton and Deadpool 2 which have a lower rating.

```{r}
df %>%
  group_by(title) %>%
  summarise(min=min(rating),
            mean=mean(rating),
            median=median(rating),
            max=max(rating))
```