### 2018-09-07 DATA607/Hands on Lab 2-3.Rmd
Initial commit

---
title: "Hands on Lab 2-3: Combining Data from Multiple Sources"
author: "Mike Silva"
date: "September 7, 2018"
output: html_document
---

## Description

Your task is to create an R dataframe that shows rates of tuberculosis infection by country. You should use the information from the tb database and the attached population.csv file. Your R dataframe should have the following columns:

Country, Year, Rate

Where Rate is defined as Cases/Population.

## Step 1: Read in the Data

First I will extract the TB cases from the MySQL database.  I will make use of a MySQL Settings file which hold authentication credentials.  It is not included in the [GitHub repo](https://github.com/mikeasilva/CUNY-SPS/tree/master/DATA607) for obvious reasons.

```{r connect,message=FALSE}
library(RMySQL)
library(ggplot2)
library(dplyr)
source('hidden/MySQL Settings.R')
conn <- dbConnect(MySQL(), user=user, password=password, dbname='tb', host=host)
```