### 2018-09-11 DATA607/Week_3_Assignment.Rmd
Linted Week 3 Assignment for DATA 607

---
title: "Week 3 Assignment"
author: "Mike Silva"
date: "September 9, 2018"
output: html_document
---

## Description

The following problems come from chapter 8, regular expressions and essential string functions from "Automated Data Collection with R"

### Problem 3

```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

name
```

(a) Use the tools of this chapter to rearrange  the vector so that all elements conform to the standard *first_name* *last_name*.

```{r}
get_first_last <- function(n){
  # Strip out title and middle name
  n <- str_replace(n, "(Rev. )|(C. )|(Dr. )", "")
  # Reorder if comma is present
  split_n <- unlist(str_split(n, ", "))
  if (length(split_n) == 2){
    n <- paste(split_n[2], split_n[1])
  }
  return(n)
}

first_name_last_name <- as.vector(sapply(name, get_first_last))

first_name_last_name
```

(b) Construct a logical vector indicating whether a character has a title (i.e. Rev. and Dr.)

```{r}
# Look for two letters followed by a period and a space
has_title <- str_count(name, "\\w{2}. ") > 1
has_title
```

(c) Construct a logical vector indicating whether a character has a second name.

```{r}
# Find a space followed by one letter then a period then a space
has_second_name <- str_count(name, " \\w{1}. ") == 1
has_second_name
```

## Problem 4

Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression

| Question | Regular Expression | Type of String | Examples |
|----------|--------------------|----------------|----------|
| (a) | [0-9]+\\$ | Numbers followed by a dollar sign| 100$, 25$ |
| (b) | \\b[a-z]{1,4}\\b | A one to four lower-case character string  | help, bat | 
| (c) | .*?\\.txt$ | Alphanumeric characters followed by .txt | readme.txt, 1.txt |
| (d) | \\d{2}/\\d{2}/\\d{4} | A date in a DD/MM/YYYY or MM/DD/YYYY format | 12/31/2018 |
| (e) | <(.+?)>.+?</\\1> | Text between to HTML tags | `<b>This is bold</b>` |