### 2018-10-02 DATA607/Project_2.Rmd
Initial commit of DATA 607 Project 2

---
title: "DATA 607 Project 2"
author: "Mike Silva"
date: "October 1, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float: true
---

## Description

This project involves transforming three untidy datasets into tidy datasets using dplyr and tidyr.  Once the data is transformed a brief exploratory data analysis will be preformed.  Since we are using dplyr and tidyr we will load them now (all other libraries will be loaded as needed).

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
```

## 2017 United Nations International Migrant Stock

The first dataset that will undergo the cleanup is the 2017 International Migrant Stock data produced by the United Nations Department of Economic and Social Affairs Population Division.  They produce multiple datasets.  This project will use the immigrants by origin and destination.

```{r echo=FALSE}
# [http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf](http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf)
# 
# This dataset provides estimates of the international migrant stock by age, sex and origin for the mid-point (1 July) of each year: 1990, 1995, 2000,
# 2005, 2010, 2015 and 2017. 
# 
# Most of the data used to estimate the international migrant stock by country or area were obtained from population censuses. Additionally, population registers and nationally representative surveys provided information on the number and composition of international migrants.
# 
# In estimating the international migrant stock, international migrants have been equated with the foreignborn population whenever this information is available, which is the case in most countries or areas. In most countries lacking data on place of birth, information on the country of citizenship of those enumerated was available, and was used as the basis for the identification of international migrants, thus effectively equating, in these cases, international migrants with foreign citizens.
# 
# Equating international migrants with foreign citizens when estimating the migrant stock has important shortcomings. In countries where citizenship is conferred on the basis of jus sanguinis, people who were born in the country of residence may be included in the number of international migrants even though they may have never lived abroad. Conversely, persons who were born abroad and who naturalized in their 
```

### Data Acquisition

We will download the migration stock by destination and origin data to our local system.

```{r}
file_name <- "data/UN_MigrantStockByOriginAndDestination_2017.xlsx"
if(!file.exists(file_name)){
  download.file("http://www.un.org/en/development/desa/population/migration/data/estimates2/data/UN_MigrantStockByOriginAndDestination_2017.xlsx", file_name)
}
```

### Parsing Untidy Data

I will begin by reading in a few rows and examing them to determine what actions need to be taken.

```{r, results='asis'}
library(openxlsx)
library(kableExtra)
read.xlsx(file_name, sheet = "Table 1", rows = 1:25) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
  scroll_box(width = "100%", height = "200px")
```

#### Developing a Workplan

There is a lot of data at the begining of the file that needs to be overlooked.  When I read in the data I will skip down to the row after the one that begins with "year".  The country names are in that row.  Thre are ".." that will need to be changed to NAs.  The data will need to be renamed. I won't need the sort order, code or type of data columns so I will drop them.

#### Execution of the Plan

```{r}
un_df <- read.xlsx('data/UN_MigrantStockByOriginAndDestination_2017.xlsx', 'Table 1', startRow = 16, na.strings = "..") %>%
  rename(Year = X1,
         Sort.order = X2,
         Destination = X3,
         Notes = X4,
         Code = X5,
         Type.of.data = X6) %>%
  select(-Sort.order, -Code, -Type.of.data)
```

Now that these steps have been taken let's take a look at the data frame to determine what to do next.

```{r, echo=FALSE}
un_df[1:10,] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
  scroll_box(width = "100%", height = "200px")
```

The Notes column is not NA on the rows I want to keep.  There are regional aggregates that I don't want to include too.  These aggregates have letters in the Notes column.  If I try to coorce them into numeric data they will turn into NAs and I can drop all the rows in one fell swoop.  I will then drop the Notes, Total, Other.North and Other.South columns.

```{r, warning=FALSE}
un_df <- un_df %>%
  mutate(Notes = as.numeric(Notes)) %>% # Introduces NAs where desired
  drop_na(Notes) %>%
  select(-Notes:-Other.South)
```

Now we are ready to change the data from wide to long.

```{r}
un_df <- un_df %>%
  gather(key = Origin, value = Migrants, -Year, -Destination) %>%
  drop_na(Migrants)
```

Now I need to clean up the Origins columns.  There are periods that represent spaces.  When there are two periods it needs to be changed into a period and space.  An example of that is Micronesia.(Fed..States.of).

```{r}
un_df <- un_df %>%
  mutate(Origin = str_replace_all(Origin, "[.]", " ")) %>%
  mutate(Origin = str_replace_all(Origin, "  ", ". "))
```

And here is the results:

```{r, echo=FALSE}
un_df[1:10,] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
  scroll_box(width = "100%", height = "200px")
```

Text