### 2018-10-05 DATA606/Silva-Proposal.Rmd
Initial commit of DATA 606 Proposal

---
title: "DATA 606 Data Project Proposal"
author: "Mike Silva"
date: "October 3, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 4
    toc_float: true
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data

```

### Research question

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

### Cases

**What are the cases, and how many are there?**

### Data collection

**Describe the method of data collection.**

### Type of study

**What type of study is this (observational/experiment)?**

### Data Source

**If you collected the data, state self-collected. If not, provide a citation/link.**

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

### Relevant summary statistics

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```