### 2018-10-11 DATA607/Week_7_Assignment.Rmd
Initial commit of Week 7 Assignment

---
title: "Week 7 Assignment"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(rmdformats)
library(kableExtra)
```

```{r knitr_init, echo=FALSE, message=FALSE, eval=FALSE}
library(dplyr)
library(knitr)
library(rmdformats)
library(kableExtra)

## Global options
options(max.print="75")
# opts_chunk$set(echo=FALSE,
# 	             cache=TRUE,
#                prompt=FALSE,
#                tidy=TRUE,
#                comment=NA,
#                message=FALSE,
#                warning=FALSE)
opts_knit$set(width=75)
```

```{r, echo=FALSE}
# A Prefix nulling hook.
# Taken from https://stackoverflow.com/questions/22524822/how-can-i-remove-the-prefix-index-indicator-1-in-knitr-output

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
    x <- gsub( "##", "", x)
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

## Description

In this assignment I have selected three books that I enjoyed reading and have entered information these three books in separately created files.  One encodes the information in HTML, XML and JSON formats.  I am to pull in the data into R and answer if the three data frames are identical.

### Book Files

I created the files by hand and saved the [HTML](https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html), [XML](https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml) and [JSON](https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json) files to my GitHub repository.  I have also loaded *dplyr* into my R environment.

## Reading HTML File into R

Here's what the HTML file looks like:

##### Table 1. Preview of books.html
```{r, null_prefix=TRUE, echo=FALSE}
print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html"), quote=FALSE)
```

I will use rvest to extract the data from the table.

```{r, echo=TRUE, message=FALSE}
library(rvest)

html_df <- read_html("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html") %>%
  html_node("table") %>%
  html_table()
```

Here's what the HTML data looks like after processed by rvest:

##### Table 2. Rvest Data Frame of books.html
```{r, echo=FALSE}
library(kableExtra)
html_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

## Reading XML File into R

Now it's time to work on the XML file.  Here's what the XML file looks like:

##### Table 3. Preview of books.xml
```{r, null_prefix=TRUE, echo=FALSE}
print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml"), quote=FALSE)
```

I will use xml2 to extract the data from the XML file.

```{r, echo=TRUE}
library(xml2)
books_xml <- read_xml("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml")

X1 <- books_xml %>%
  xml_children() %>%
  xml_name()

X2 <- books_xml %>%
  xml_children() %>%
  xml_text()

xml_df <-data.frame(X1, X2)
```

Here's what the XML data looks like after passing through the above process:

##### Table 4. Data Frame of books.xml
```{r, echo=FALSE}
xml_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

## Reading JSON File into R

Here's what the JSON file looks like:

##### Table 5. Preview of books.json
```{r, null_prefix=TRUE, echo=FALSE}
print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json"), quote=FALSE)
```

I will use jsonlite to extract the data from the file.

```{r, echo=TRUE}
library(jsonlite)
json_df <- fromJSON("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json") %>%
  unlist() %>%
  as.data.frame()

# Change the name
names(json_df) <- c("X2")
# Pull the row names into X1
json_df <- data.frame(X1 = row.names(json_df), json_df)
# Drop the row names
rownames(json_df) <- c()
```

Here's what the JSON data looks like after processed by rvest:

##### Table 6. Data Frame of books.json
```{r, echo=FALSE}
json_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

## Are the Three Data Frames Identical?

The three data frames are close.  I processed the XML file so that it is identical to