### 2018-10-22 DATA607/Week_9_Assignment.Rmd
Modified Week_9_Assignment.Rmd
Linted document and made minor changes

---
title: "CUNY SPS DATA-607 Week 9 Assignment"
author: "Mike Silva"
date: "22 October 2018"
output:
  rmdformats::readthedown:
    highlight: kate
---

## Description

The assignment is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

## Data Acquisition

I have chosen to pull the top stories from the upshot section of the paper.  I will request the data from the API

```{r}
library(httr)
source("hidden/NYT API.R") # Load my API key into api_key
base_url <- "https://api.nytimes.com/svc/topstories/v2/upshot.json"
url <- paste0(base_url, "?api-key=", api_key)
response <- GET(url)
```

## Data Transformation

Now the JSON response needs to be transformed into a data frame:

```{r, message=FALSE}
library(jsonlite)
json <- fromJSON(content(response, as = "text"))
df <- data.frame( json$results )
```

## Results

The following is the data pulled from the New York Times API:

##### Table 1. Response from the API
```{r, echo=FALSE, message=FALSE}
library(kableExtra)
library(dplyr)
df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```