### 2018-10-30 DATA607/Project_4.Rmd
Initial commit

---
title: "Document Classification"
date: "29 October 2018"
output:
  rmdformats::readthedown:
    highlight: kate
---

## Description

This project seeks to build a classifier that can predict if an email is spam or ham.  It will be using the Apache Spam Assassin public corpus as the dataset.

### Data Aquisition

The first step is to download the files.  I am selecting the most recent releases of spam and easy ham.

```{r file_download}
if(!file.exists("data/20050311_spam_2.tar.bz2") && !file.exists("data/20050311_spam_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar.bz2")
}

if(!file.exists("data/20030228_easy_ham_2.tar.bz2") & !file.exists("data/20030228_easy_ham_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar.bz2")
}
```

Now that the files have been aquired, they need to be decompressed.

```{r decompress_files}
if(!dir.exists("data/spam_2")){
  R.utils::bunzip2("data/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar")
  untar("data/20050311_spam_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/spam_2/cmds")
}

if(!dir.exists("data/easy_ham_2")){
  R.utils::bunzip2("data/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar")
  untar("data/20030228_easy_ham_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/easy_ham_2/cmds")
}
```

## Approach

```{r}
spam_docs <- list.files("data/spam_2")
ham_docs <- list.files("data/easy_ham_2")
```

## Processing Pipeline

### Split up Sentance

### Remove Punctuation and Stop Words

### Stemming