### 2018-10-30 DATA607/Project_4.Rmd
More work on Project 4

---
title: "Document Classification"
date: "29 October 2018"
output:
  rmdformats::readthedown:
    highlight: kate
---

## Description

This project seeks to build a classifier that can predict if an email is spam or ham.  It will be using the Apache Spam Assassin public corpus as the dataset.

### Data Aquisition

The first step is to download the files.  I am selecting the most recent releases of spam and easy ham.

```{r file_download}
if(!file.exists("data/20050311_spam_2.tar.bz2") && !file.exists("data/20050311_spam_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar.bz2")
}

if(!file.exists("data/20030228_easy_ham_2.tar.bz2") & !file.exists("data/20030228_easy_ham_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar.bz2")
}
```

Now that the files have been aquired, they need to be decompressed.

```{r decompress_files}
if(!dir.exists("data/spam_2")){
  R.utils::bunzip2("data/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar")
  untar("data/20050311_spam_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/spam_2/cmds")
}

if(!dir.exists("data/easy_ham_2")){
  R.utils::bunzip2("data/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar")
  untar("data/20030228_easy_ham_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/easy_ham_2/cmds")
}
```

## Processing Pipeline

### Train/Test Split

First we are going to split the data into training and testing sets using an 80:20 split.

```{r}
set.seed(12345)
training_percent <- 0.8

train_test_split <- function(data, training_percent){
  n <- round(length(data)*training_percent, digits = 0)
  train <- sort(sample(data, n, replace = FALSE))
  test <- data[!data %in% train]
  return(list(test=test, train=train))
}

spam <- train_test_split(list.files("data/spam_2"), training_percent)
ham <- train_test_split(list.files("data/easy_ham_2"), training_percent)
```

The following table summarizes the number of messages included in each set:

##### Table 1. Spam and Ham Message Counts by Data Set
|      |       Training       |        Testing      |
|------|----------------------|---------------------|
| Spam |`r length(spam$train)`|`r length(spam$test)`|
| Ham  |`r length(ham$train)` |`r length(ham$test)` |

### Pull out Email Contents

This model will classify only on the basis of the content.  The header information will not be taken into consideration.  In order to illustrate the process I will take the first message in the spam training set as an example.

```{r}
example_file <- spam$train[1]

get_email_contents <- function(file_name, path){
 full_path = paste0(path, file_name)
 con <- file(full_path)
 readLines(con)
}

```

### Split up Sentance

### Remove Punctuation and Stop Words

### Stemming