### 2018-11-15 DATA606/Silva-Proposal.Rmd
Updated proposal

---
title: "DATA 606 Data Project Proposal"
author: "Mike Silva"
date: "October 3, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 4
    toc_float: true
---

## Research Question

What effect does the age of a home and its square footage have on its sales price in Monroe County, New York?

## Data Collection

The sales price data were collected through governmental reporting.  The price data is through the NYS Office of Real Property Services and provided to the general public by Monroe County through their real property portal.  The housing attributes is compiled by the NYS GIS clearinghous from data provided by local assessors.

## Type of Study

This study will be an observational study.

## Dependent Variable

The response variable is the housing price.  This study will use the latest sales price available. This is a quantitative variable.  Since the housing attributes are as of 2017 we will only include sales that occured in 2016-2018 (assuming the homes haven't changed much since 2017).

## Independent Variable

The two variables will be the square footage of the living space (quantitative) and the era the home was built (qualitative).  These eras will follow the groups used by the Trulia's Chief Economist in [this article](https://www.forbes.com/sites/trulia/2013/05/02/american-homes-by-decade/#16c559821045).

## Data Source

### Sales Price Data

Monroe County's Real Property Portal. (2018).  Database was synced with NYS RPS on 22 October 2018. Scraped from [https://www.monroecounty.gov/etc/rp](https://www.monroecounty.gov/etc/rp) on 10 November 2018.

### Housing Attributes

NYS GIS Clearinghouse. (2017).  Monroe County Tax Parcel Centroids [Data file].  Retrieved from [http://gis.ny.gov/gisdata/inventories/details.cfm?DSID=1300](http://gis.ny.gov/gisdata/inventories/details.cfm?DSID=1300).

## Data Preparation

The Monroe County Real Property portal has been scraped and the data has been stored in an SQLite database.

```{r get_the_database, message=FALSE}
library(DBI)
library(dplyr)

# Download the database
if(!file.exists("Monroe Real Property Data.db")){
  if(!file.exists("Monroe Real Property Data.db.tar.gz")){
    # Download this specific commit from my GitHub repo
    download.file("https://github.com/mikeasilva/CUNY-SPS/raw/9a2023b57a466550646c9190076a53d51ef09ef2/data/Monroe%20Real%20Property%20Data.db.tar.gz", "Monroe Real Property Data.db.tar.gz")
  }
  # Decompress the file
  untar("Monroe Real Property Data.db.tar.gz")
}
```

Now that the data is downloaded it's time to read it in.

```{r read_db}
# Connect to the database
con <- dbConnect(RSQLite::SQLite(), "Monroe Real Property Data.db")

# Get the latest unduplicated sales price data
results <- dbSendQuery(con, "SELECT MAX(sale_date) AS sale_date, SWIS, SBL FROM sales_data GROUP BY SWIS, SBL HAVING sale_date > '2015-12-31'")
df <- dbFetch(results)

dbClearResult(results)
```

There are `r nrow(df)` sales prices in the data frame.

```{r sales_data}
df <- dbReadTable(con, "sales_data") %>% # Pull all the records
  distinct() %>% # remove any duplicates
  merge(df) %>% # Merge with the targeted sales observations
  rename(Price = price)
```

The count of observations is now `r nrow(df)`.  There was some duplication introduced through the merging process.  I will drop all the observations with more than one price.

```{r deduplication}
df <- df %>%
  select(SWIS, SBL) %>% # Select the unique identifiers
  group_by(SWIS, SBL) %>% 
  summarise(count = n()) %>% # Get the number of records
  ungroup() %>%
  filter(count == 1) %>% # Only keep the observations with one price
  select(-count) %>% # Drop the count
  merge(df) # subset the price observations
```

There are `r nrow(df)` observations in the data frame.  Now that we have the dependent variable, it is time to add in the housing attributes.

```{r housing_attributes}
# Add in the housing attributes
df <- dbReadTable(con, "property_info") %>% # Pull all the records
  distinct() %>% # remove any duplicates
  merge(df) # Merge with the targeted sales observations

dbDisconnect(con)
```

Now that the housing attributes are added in I will filter out observations that are odd.  Some variable have 0 when there isn't data.

```{r filter_out_crap}
df <- df %>% # First filter out some obviously weird data
  filter(YR_BLT > 0) %>%
  filter(SQFT_LIV > 0) %>%
  filter(NBR_KITCHN > 0) %>%
  filter(NBR_BEDRM > 0)
```

The last step is to change the year built quantitative variable into a qualitative variable.

```{r finalize_df}
get_era <- function(year){
  if(year < 1940){
    return("1. Before 1940s")
  } else if(year < 1980){
    return("2. Post-War Boom")
  }else{
    return("3. Last 30+ Years")
  }
}

df <- df %>% 
  rowwise() %>%
  mutate(Era = get_era(YR_BLT)) %>%
  ungroup() %>%
  mutate(Era = as.factor(Era)) %>%
  rename(Size = SQFT_LIV) %>%
  select(Price, Size, Era)
```

## Cases

Each case is a residential property in Monroe County, New York.  There are **`r nrow(df)`** cases in the data set.

## Relevant Summary Statistics

Here's the summary of the data

##### Table 1: Summary of Data
```{r summary, comment=NA, message=FALSE, echo=FALSE}
library(kableExtra)
summary(df)
```

There are extreme price and size values on both ends.  There will probably need to be some more filtering of the the data.  Let's explore the data with some visualizations.

```{r hist, echo=FALSE}
library(ggplot2)
library(ggthemes)

ggplot(df, aes(Price)) + 
  geom_histogram(bins = 100) + 
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight(base_size = 12, base_family = "sans")
```

##### Figure 1: Price and Size Scatterplot
```{r echo=FALSE}
ggplot(df, aes(Price, Size)) + 
  geom_point() + 
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight(base_size = 12, base_family = "sans")
```

This looks encouraging.  I probably should remove the few high price outliers from the data set.

##### Figure 2: Price and Era Boxplot
```{r era_by_price_boxplot, echo=FALSE}
ggplot(df, aes(x=Era, y=Price)) + 
  geom_boxplot() + 
  theme_fivethirtyeight(base_size = 12, base_family = "sans")
```

This visualization is not very helpful because of the outliers.  Let's look at the same data in tabular form.

##### Table 2: Price Summary by Era
```{r price_by_era_table, echo=FALSE}
df %>%
  group_by(Era) %>%
  summarise(min = min(Price), mean = mean(Price), median = median(Price), max = max(Price)) %>%
  kable() %>%
  kable_styling()
```

Again it looks like more work will need to be done on the price data to clean up the records that are more transactional than capturing the value of a home.  Is a home really worth $2.00?