### 2019-02-19 DATA605/MSilva_Assign4.Rmd
Work on assignment 4

---
title: "DATA 605 Assignment 4"
author: "by Michael Silva"
output: html_document
---

## Problem Set 1

**Given a 3 × 2 matrix A**

\[ A=\begin{bmatrix}
  1 & 2 & 3 \\
  -1 & 0 & 4
  \end{bmatrix}
\]

**write code in R to compute $X = AA^T$ and $Y = A^TA$.**

```{r}
A <- matrix(c(1, -1, 2, 0, 3, 4), 2)
X <- A %*% t(A)
Y <- t(A) %*% A
```

**Then, compute the eigenvalues and eigenvectors of X and Y using the built-in commans in R.**

```{r}
X_eigen <- eigen(X)
Y_eigen <- eigen(Y)
```

**Then, compute the left-singular, singular values, and right-singular vectors of A using the svd command.**

```{r}
A_svd <- svd(A)
```

**Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y.**

```{r}
A_svd$u
X_eigen$vectors
```

One might observe a difference in signs.  According to [this post](https://stat.ethz.ch/pipermail/r-help/2003-July/036504.html) on the R help forum by Roger Peng:

>That is true, but using different compilers on the same machine and the 
>same version of R may give different signs for the eigenvectors.  The 
>moral is, don't rely on the signs of eigenvectors!  (This *is* on the help 
>page.)

```{r}
A_svd$v
Y_eigen$vectors[, 1:2] # The third value is basically zero
```

**In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A.**

```{r}
X_eigen$values
Y_eigen$values[1:2]
A_svd$d ^ 2
```

## Problem Set 2

**Using the procedure outlined in section 1 of the weekly handout, write a function to compute the inverse of a well-conditioned full-rank square matrix using co-factors. In order to compute the co-factors, you may use built-in commands to compute the determinant.**

```{r}
myinverse <- function(A){
  # Initialize the cofactors matrix
  matrix_of_cofactors <- matrix(NA, nrow = nrow(A), ncol = ncol(A))
  for (i in c(1:nrow(A))){
    for (j in c(1:ncol(A))){
      # Get the determinant of the matrix except the current column and row
      matrix_of_cofactors[i, j] <- det(A[-i, -j]) * (-1) ^ (i + j)
    }
  }
  # Transpose all elements of the previous matrix of cofactors
  adjugate <- t(matrix_of_cofactors)
  # Divide the ajugate by the determinant of the original matrix
  inverse <- adjugate / det(A)
  return(inverse)
}

A <- matrix(c(3, 2, 0, 0, 0, 1, 2, -2, 1), 3)
B <- myinverse(A)

A %*% B
```