### 2019-02-25 DATA605/MSilva_Assign5.Rmd
Initial commit

---
title: "DATA 605 Assignment 5"
author: "by Michael Silva"
output: html_document
---

Choose independently two numbers B and C at random from the interval [0, 1] with uniform density.

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
set.seed(12345)
n <- 1000000
B <- runif(n, min = 0, max = 1)
C <- runif(n, min = 0, max = 1)
data <- data.frame(B = B, C = C)
```

Prove that B and C are proper probability distributions.

```{r, echo=FALSE}
hist(data$B, main = "Histogram of B", xlab = "")
```

```{r, echo=FALSE}
hist(data$C, main = "Histogram of C", xlab = "")
```

Note that the point (B,C) is then chosen at random in the unit square.

Find the probability that

## (a) B + C < 1/2

```{r, message=FALSE}
data <- data %>%
  mutate(a = ifelse(B + C < 1/2, 1, 0))
sum(data$a) / n
```

This is represented by the area in blue

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(a))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (b) BC < 1/2

```{r}
data <- data %>%
  mutate(b = ifelse(B * C < 1/2, 1, 0))
sum(data$b) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(b))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (c) |B - C| < 1/2

```{r}
data <- data %>%
  mutate(c = ifelse(abs(B - C) < 1/2, 1, 0))
sum(data$c) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(c))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (d) max(B, C) < 1/2

```{r}
data <- data %>%
  mutate(d = ifelse(pmax(B, C) < 1/2, 1, 0))
sum(data$d) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(d))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (e) min(B, C) < 1/2

```{r}
data <- data %>%
  mutate(e = ifelse(pmin(B, C) < 1/2, 1, 0))
sum(data$e) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(e))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```