### 2019-03-16 DATA605/MSilva_Assign7.Rmd
More work on Assignment 7

---
title: "DATA 605 Week 7 Assignment"
author: "Mike Silva"
output: html_document
---

## Question 1

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.

```{r}
set.seed(42)
n <- 100000
k <- 50
draws <- 10
Y <- c()
for (i in 1:n){
  Xn <- sample(1:k, draws, TRUE)
  Y <- c(Y, min(Xn))
}
```

```{r, echo=FALSE}
library(ggplot2)

ggplot(data.frame(table(Y)), aes(Y, Freq)) + 
  geom_bar(stat = "identity", fill = "steelblue", width = 1) +
  theme_minimal() +
  ylab("Count") + 
  ggtitle("Distribution of Y's")
```

It's a long tail distribution like a power law distribution.

## Question 2

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.)

### Part A

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a geometric.**

The probability the machine will fail within ten years is 1.  The probability that the machine will fail within k years is found using:

$$P(x=k) = (1-p)^{k-1} \times p$$
where p is the probability of the event happening that year.  We are given that p is 1/10.  The probability it will fail in the first eight years is displayed below

```{r echo=FALSE}
n_failures_before_success <- 8
probability_of_success <- 1/10

p_ns <- c()
for (n in 1:n_failures_before_success){
  p_n <- probability_of_success * (1 - probability_of_success)^(n-1)
  p_ns <- c(p_ns, p_n)
}

ggplot(data.frame(year = 1:n_failures_before_success, p = p_ns), aes(year,p)) + 
  geom_bar(stat = "identity", fill = "steelblue", width = 1) +
  geom_text(aes(label = round(p,3)), vjust = 1.6, color = "white", size = 3.5) +
  theme_minimal() +
  ggtitle("Probability of Failure in the First 8 Years")
```

The probability of failure after the eight years will be one minus the sum of all the probabilities above (the area in blue).  Otherwise stated:

$$P(\text{failure after 8 years}) = 1 - \sum_{k=1}^{8} \frac{1}{10}\times\Big(\frac{9}{10}\Big)^{k-1} = 0.4304672$$

```{r}
n_failures_before_success <- 8
probability_of_success <- 1/10

p <- 0
for (n in 1:n_failures_before_success){
  p <- p + probability_of_success * (1 - probability_of_success) ^ (n - 1)
}

p_geometric <- 1 - p

message(paste("Probability that the machine will fail after 8 years:", p_geometric))
```

The [expected value](https://math.stackexchange.com/questions/2368304/how-do-you-calculate-the-expected-value-of-geometric-distribution-without-diffre) of a geometric distribution is

$$\text{Expected Value} = \frac{1}{P({success})} = \frac{1}{\frac{1}{10}} = 10$$

```{r}
expected_value_geom <- 1 / probability_of_success

message(paste("Expected value:", expected_value_geom))
```

The [standard deviation](https://www.youtube.com/watch?v=kDFiNt8nxSo) is:

$$\text{Standard Deviation} = \sqrt{\frac{1-P(success)}{P(success)^2}} = \sqrt{\frac{1-0.1}{0.1^2}} = \sqrt{\frac{0.9}{0.01}} = \sqrt{90} \approx 9.486833$$

```{r}
variance <- (1 - probability_of_success) / (probability_of_success ^ 2)

standard_deviation_geom <- sqrt(variance)

message(paste("Standard deviation:", standard_deviation_geom))
```

### Part B

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as an exponential.**

The probability the machine fails after 8 years is

$$P(\text{failure after 8 years}) = e^\frac{-8}{10} = e^{-0.8}=0.449329$$

```{r}
p_exponential <- exp(1) ^ (-n_failures_before_success /expected_value_geom)

message(paste("Probability that the machine will fail after 8 years:", p_exponential))
```

The expected value is found using the same formula as above as demonstrated in [this video](https://www.youtube.com/watch?v=GJ2klfD0Q3g). Therefore

$$\text{Expected Value} = \frac{1}{P({success})} = \frac{1}{\frac{1}{10}} = 10$$

```{r}
expected_value_exp <- 1 / probability_of_success

message(paste("Expected value:", expected_value_exp))
```

The [standard deviation](https://www.youtube.com/watch?v=GJ2klfD0Q3g) is:

$$\text{Standard Deviation} = \sqrt{\frac{1}{P(success)^2}} = \sqrt{\frac{1}{0.1^2}} = \sqrt{\frac{1}{0.01}} = \sqrt{100} = 10$$

```{r}
standard_deviation_exp <- sqrt(1 / (probability_of_success ^ 2))

message(paste("Standard deviation:", standard_deviation_exp))
```

### Part C

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)**

The binomial probability formula is

$$P(x) = \frac{N!}{x!(N-x)!}\pi^x(1-\pi)^{N-x}$$

The probability of the maching failing after eight years is 1 minus the probability it does not fail in the first 8 years.  Otherwise put 0 successes(x) in the first 8 years (N).  The probability of failure is 1/10 therefore:

$$P(0) = \frac{8!}{0!(8-0)!}\frac{1}{10}^0(1-\frac{1}{10})^{8-0} = \frac{40320}{40320}\times 1 \times \frac{9}{10}^8 = 1 \times 1 \times 0.9^8 = 0.4304672$$

```{r}
p_binomial <- pbinom(0, 8, 0.1)
message(paste("Probability that the machine will fail after 8 years:", p_binomial))
```

The [expected value](http://onlinestatbook.com/2/probability/binomial.html) is the number of trials times the probability of success per trial.

```{r}
expected_value_binom <- n_failures_before_success * probability_of_success

message(paste("Expected value:", expected_value_binom))
```

The [standard deviation](http://onlinestatbook.com/2/probability/binomial.html) is found using:

$$\sigma = \sqrt{N\pi(1-\pi)}$$

```{r}
standard_deviation_binom <- sqrt(n_failures_before_success * probability_of_success*(1 - probability_of_success))

message(paste("Standard deviation:", standard_deviation_binom))
```

## Part D

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a Poisson**

The Poisson formula is:

$$P(x) = \frac{\lambda^xe^{-\lambda}}{x!}= \frac{10^8e^{-10}}{8!} =\frac{4539.993}{40320} = 0.112599$$

```{r}
x <- 8
mu <- 10
p_poisson <- ((mu^x) * (exp(1)^-mu)) / factorial(x)
message(paste("Probability that the machine will fail after 8 years:", p_poisson))
```

For the Poisson distribution the mean and standard deviation are the same.

```{r}
expected_value_poisson <- mu

message(paste("Expected value:", expected_value_poisson))
```

The variance is equal to mu so

```{r}
standard_deviation_poisson <- sqrt(mu)
```

### Summary

The following table summarizes all the information found above

```{r echo=FALSE, message=FALSE}
library(kableExtra)
library(dplyr)

data.frame(Model = c("Geometric", "Exponential", "Binomial", "Poisson"),
           Probability = c(p_geometric, p_exponential, p_binomial, p_poisson),
           ex = c(expected_value_geom, expected_value_exp, expected_value_binom, expected_value_poisson),
           sd = c(standard_deviation_geom, standard_deviation_exp, standard_deviation_binom, standard_deviation_poisson)) %>%
  rename("Expected Value" = ex,
         "Standard Deviation" = sd) %>%
  kable() %>%
  kable_styling()
```