### 2019-03-18 DATA605/MSilva_Assign8.Rmd
Initial commit of Week 8 assignment

---
title: "DATA 605 Week 8 Assignment"
author: "Mike Silva"
output: html_document
---

## Question 11 on Pg. 303

A company buys 100 lightbulbs, each of which has an exponential lifetime of
1000 hours. What is the expected time for the first of these bulbs to burn
out?

## Question 14 on Pg. 303

Assume that X1 and X2 are independent random variables, each having an
exponential density with parameter λ. Show that Z = X1 − X2 has density

$$fz(z)=(1/2)\lambda e ^{-\lambda \mid z \mid}$$

## Question 1 Pg. 320-321

 Let X be a continuous random variable with mean µ = 10 and variance
$\sigma^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the
following probabilities.

### (a) P(|X − 10| ≥ 2).

### (b) P(|X − 10| ≥ 5).

### (c) P(|X − 10| ≥ 9).

### (d) P(|X − 10| ≥ 20).