### 2019-03-21 DATA605/MSilva_Assign8.Rmd
More work on Assignment 8

---
title: "DATA 605 Week 8 Assignment"
author: "Mike Silva"
output: html_document
---

## Question 11 on Pg. 303

**A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?**

The exponential formula is:

$$N(t) = N_0 e ^ {-\lambda t} $$

lambda would be 1 failure every 1000 hours.  Otherwise stated:

$$\lambda = \frac{1}{1000}$$

Now we need to plug in the variables and solve for t

$$100 e ^ {-\frac{1}{1000}t} = 99 \rightarrow e ^ {-\frac{1}{1000}t} = \frac{99}{100} \rightarrow -\frac{1}{1000}t = ln\bigg(\frac{99}{100}\bigg) \rightarrow t = ln\bigg(\frac{99}{100}\bigg)\times-1000 \approx 10.05034$$

**Answer: 10.05034 hours**

## Question 14 on Pg. 303

**Assume that X1 and X2 are independent random variables, each having an exponential density with parameter λ. Show that Z = X1 − X2 has density**

$$fz(z)=(1/2)\lambda e ^{-\lambda \mid z \mid}$$

## Question 1 Pg. 320-321

**Let X be a continuous random variable with mean µ = 10 and variance $\sigma^2$ = 100/3. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.**

### (a) P(|X − 10| ≥ 2).

### (b) P(|X − 10| ≥ 5).

### (c) P(|X − 10| ≥ 9).

### (d) P(|X − 10| ≥ 20).