### 2019-04-02 DATA605/MSilva_Assign10.Rmd
Completion of Assignment 10

---
title: "DATA 605 Week 10 Assignment"
author: "Mike Silva"
output: html_document
---

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.  A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if

(a) he bets 1 dollar each time (timid strategy).

**About 2%**

```{r timid}
states <- c("0", "1", "2", "3", "4", "5", "6", "7", "8")
transition_matrix <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0, 
                              0.6, 0, 0.4, 0, 0, 0, 0, 0, 0,
                              0, 0.6, 0, 0.4, 0, 0, 0, 0, 0,
                              0, 0, 0.6, 0, 0.4, 0, 0, 0, 0,
                              0, 0, 0, 0.6, 0, 0.4, 0, 0, 0,
                              0, 0, 0, 0, 0.6, 0, 0.4, 0, 0,
                              0, 0, 0, 0, 0, 0.6, 0, 0.4, 0,
                              0, 0, 0, 0, 0, 0, 0.6, 0, 0.4,
                              0, 0, 0, 0, 0, 0, 0, 0, 1), 
                            nrow = 9, 
                            byrow = TRUE,
                            dimnames = list(states, states))

starting_state <- c(0, 1, 0, 0, 0, 0, 0, 0, 0)

n <- 1000

for (i in 1:n){
  starting_state <- starting_state %*% transition_matrix
}
starting_state
```

```{r, echo = FALSE, eval=FALSE}
library(markovchain)
timid <- new("markovchain", states = states, byrow = TRUE, transitionMatrix = transition_matrix, name = "Timid")

c(0, 1, 0, 0, 0, 0, 0, 0, 0) * timid^1000
```

(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).

**About 6.4%**

```{r bold}
transition_matrix <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0, 
                              0.6, 0, 0.4, 0, 0, 0, 0, 0, 0,
                              0.6, 0, 0, 0, 0.4, 0, 0, 0, 0,
                              0.6, 0, 0, 0, 0, 0, 0.4, 0, 0,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0, 0, 0, 0, 0, 0, 0, 0, 1), 
                            nrow = 9, 
                            byrow = TRUE,
                            dimnames = list(states, states))

starting_state <- c(0, 1, 0, 0, 0, 0, 0, 0, 0)

for (i in 1:n){
  starting_state <- starting_state %*% transition_matrix
}
starting_state
```

(c) Which strategy gives Smith the better chance of getting out of jail?

**The bold strategy**