### 2019-04-13 DATA605/MSilva_Assign11.Rmd
Initial commit of week 11 work

---
title: "DATA 605 Week 11 Assignment"
author: "Mike Silva"
output: html_document
---

Using the "cars" dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

## Visualization

```{r echo=FALSE}
library(ggplot2)
library(ggthemes)

ggplot(cars, aes(speed, dist)) +
  geom_point() +
  ylab("Stopping Distance (Feet)") +
  xlab("Speed (MPH)") + 
  theme_economist_white()
```

There appears to be a positive linear relationship between the speed of a car and the stopping distance.

## Model Creation & Evaluation

Let's fit a line to the data

```{r}
model <- lm(dist ~ speed, cars)
```

And let's visualize it

```{r echo=FALSE}
lm_eqn <- function(m){
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = unname(coef(m)[1]),
              b = unname(coef(m)[2]),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq))
}

ggplot(cars, aes(speed, dist)) +
  geom_point() +
  geom_abline(slope = model$coefficients[2], intercept = model$coefficients[1], color="red") +
  ylab("Stopping Distance (Feet)") +
  xlab("Speed (MPH)") + 
  geom_label(x = 10, y = 115, label.size = NA, label = lm_eqn(model), parse = TRUE) +
  theme_economist_white()
```

```{r comment=NA}
summary(model)
```

Every additional MPH in speed requires about 4 feet of stopping distance.  The standard error on this variable is about 9.5 times smaller than the estimate which is good.  The P-value of this coefficient is very small indicating that it is statistically significant.  This model explains about 65% of the variation.

## Residual Analysis

```{r echo=FALSE}
ggplot(cars, aes(resid(model))) +
  geom_density() +
  ylab("Density") +
  xlab("Residuals") + 
  geom_vline(xintercept = 0, linetype="dotted", color = "red") +
  theme_economist_white()
```

From the summary above and the chart of the residuals above we can note the residuals range from -29 to 43 with a median at -2.27 and the first and third quartiles are -9.5 and 9.2 respectively.  From that it would seem they are largely normaly distributed with a bit of an outlier on the the positive side.

```{r echo=FALSE}
ggplot(cars, aes(fitted(model),resid(model))) +
  geom_point() +
  ylab("Residuals") +
  xlab("Fitted Values") + 
  theme_economist_white()
```

There are three observations that look like outliers on the positive side of the distribution.  Outside of this the variation is fairly uniformly distributed.  Now for the Q-Q plot.

```{r echo=FALSE}
gg_qqplot <- function(fitted.lm) {
   # extract residuals from lm object
   res = residuals(fitted.lm)

   # calculate slope and interncept for qqline
   slope = (quantile(res, .75) - quantile(res, .25)) / (qnorm(.75) - qnorm(.25))
   intercept = quantile(res,.25) - slope*qnorm(.25)
   qq_line = data.frame(intercept = intercept, slope = slope)

   # generate ggplot for qqplot
   qq_plot <- ggplot(data = fitted.lm) +
      stat_qq(aes(sample = res)) +
      labs(x = "Theoretical Quantile", y = "Standardized Residual") +
      geom_abline(data = qq_line ,aes(intercept = intercept ,slope = slope), color = "red") +
    theme_economist_white()   
   qq_plot
}

gg_qqplot(model)
```

Again we see that most of the residuals lay on the line except for three observations which seem to be outliers.