### 2019-04-20 DATA605/MSilva_Assign12.Rmd
More work on Assignment 12

---
title: "DATA 605 Week 12 Assignment"
author: "Mike Silva"
output: html_document
---

## Part 1 - Original Data

### Data Acquisition/Wrangling

```{r}
df <- read.csv("who.csv")
```

Here's the first 5 observations of the data

```{r echo=FALSE, message=FALSE}
library(dplyr)
library(kableExtra)
df %>%
  head(5) %>%
  kable() %>%
  kable_styling()
```

### Visualization

```{r echo=FALSE}
library(ggplot2)
library(ggthemes)

ggplot(df, aes(TotExp, LifeExp)) +
  geom_point() +
  ylab("Avg Life Expectancy (years)") +
  xlab("Avg Personal and Government Expenditures (US dollars)") + 
  theme_economist_white()
```

### Model Building

```{r}
model <- lm(LifeExp~TotExp, df)
```

### Model Evaluation

The F statistic is `r summary(model)$fstatistic["value"]`.  It can give a yes-no response to "can I reject $H_0$".  In this case F statistic indicates that there's a relationship that is statistically significant and $R^2$ will indicate the strength of the relationship.

The $R^2$ is `r summary(model)$r.squared`.  It explains how much of the variation in life expectancy is explained by the total expenditures.  Only 25% of the variation is explained which is not a very good fit.

The standard error for total expenditures is `r summary(model)$coefficients[1,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[1,"Estimate"]` +/- `r summary(model)$coefficients[1,"Std. Error"] * 1.96` at the 95% confidence level.

The P-value for total expenditures is `r summary(model)$coefficients[1,"Pr(>|t|)"]` which indicates the beta is statistically significant.

## Part 2 - Transformed Data

### Data Acquisition/Wrangling

```{r message=FALSE}
library(dplyr)
df <- df %>%
  mutate(TransformedTotExp = TotExp^.06,
         TransformedLifeExp = LifeExp^4.6)
```

### Visualization

```{r echo=FALSE}
ggplot(df, aes(TransformedTotExp, TransformedLifeExp)) +
  geom_point() +
  ylab("Avg Life Expectancy (years) to the 4.6 power") +
  xlab("Avg Personal and Government Expenditures (US dollars) to the 0.06 power") + 
  theme_economist_white()
```

### Model Building

```{r}
model <- lm(TransformedLifeExp~TransformedTotExp, df)
```

### Model Evaluation

The F statistic (`r summary(model)$fstatistic["value"]`) is larger than before and indicates that there's a relationship that is statistically significant. The $R^2$ (`r summary(model)$r.squared`) means about 73% of the variation is now explained which is a much better fit.

The standard error for total expenditures is `r summary(model)$coefficients[1,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[1,"Estimate"]` +/- `r summary(model)$coefficients[1,"Std. Error"] * 1.96` at the 95% confidence level.

The P-value for the total expenditures is `r summary(model)$coefficients[1,"Pr(>|t|)"]` now which indicates the estimate is statistically significant.

This model is definately a better fit although the I personally find it harder to explain to less statistically litterate users due to the transformation.

## Part 3

```{r, results='asis'}
forecast_life_expectancy <- function(TransformedTotExp){
  d <- data.frame(TransformedTotExp = c(TransformedTotExp))
  prediction <- predict(model, d)
  cat(paste("Life Expectancy:", round(unname(log(prediction, 4.6)),1), "years"))
}
forecast_life_expectancy(1.5)
```

```{r, results='asis'}
forecast_life_expectancy(2.5)
```

## Part 4

### Model Building

```{r}
model <- lm(LifeExp ~ PropMD + TotExp + I(PropMD * TotExp), df)
```

### Model Evaluation

The F statistic (`r summary(model)$fstatistic["value"]`) is smaller than the model in Part 2.  It still indicates that there's a relationship that is statistically significant. The $R^2$ (`r summary(model)$r.squared`) means about 36% of the variation is explained by this model which is worse than the prvious model.

The standard error for Prop MD is `r summary(model)$coefficients[1,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[1,"Estimate"]` +/- `r summary(model)$coefficients[1,"Std. Error"] * 1.96` at the 95% confidence level.

The standard error for total expenditures is `r summary(model)$coefficients[2,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[2,"Estimate"]` +/- `r summary(model)$coefficients[2,"Std. Error"] * 1.96` at the 95% confidence level.

The standard error for the interaction term is `r summary(model)$coefficients[3,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[3,"Estimate"]` +/- `r summary(model)$coefficients[3,"Std. Error"] * 1.96` at the 95% confidence level.

The P-value for the first term is `r summary(model)$coefficients[1,"Pr(>|t|)"]`  which indicates the estimate is statistically significant.  The second term is `r summary(model)$coefficients[2,"Pr(>|t|)"]`  which indicates the estimate is statistically significant as is the third term (`r summary(model)$coefficients[3,"Pr(>|t|)"]` ).

This model is not as good as the second model.

## Part 5

```{r, results='asis'}
forecast_life_expectancy <- function(PropMD, TotExp){
  d <- data.frame(PropMD = c(PropMD), TotExp = c(TotExp))
  prediction <- predict(model, d)
  cat(paste("Life Expectancy:", round(unname(prediction),1), "years"))
}
forecast_life_expectancy(0.03, 14)
```

This forecast does not seem realistic because the average live expectancy is not that high.