### 2019-04-21 DATA605/MSilva_Exercise12.Rmd
Added Week 12 Exercise

---
title: "DATA 605 Week 12"
author: "Mike Silva"
date: "16 April 2019"
output: 
  rmdformats::readthedown:
    highlight: kate
---

## Introduction

As part of the Fundamentals of Computational Mathematics course work at the CUNY School of Professional Studies, we need to build a multiple regression model.  The model needs at least one quadratic term, one dichotomous term, and one dichotomous vs. quantitative interaction term.  For this exercise I will be using the clasic Iris data set.

## Data Acquisition

I will be predicting the petal length.  I know that there is a strong relationship between the sepal length and the petal length (`r cor(iris$Sepal.Length, iris$Petal.Length)` correlation).  I will be using this in the model as a quadratic term.

Because of my knowledge of this dataset I know that the setosa species of iris is distinct from the versicolor and virginica species.  I will be using that as the dichotomous term.

I want to differentiate the versicolor and viginica group and will use the petal width as the quantitative interaction term.

```{r, message=FALSE}
library(dplyr)
df <- iris %>%
  mutate(Sepal.Length.Squared = Sepal.Length ^ 2) %>% # Quadratic term
  mutate(Not.Setosa = ifelse(Species == "setosa", 0, 1)) %>% # Dichotomous term
  mutate(Petal.Width.Interaction = Not.Setosa * Petal.Width) %>% # Dichotomous vs. quantitative interaction
  select(Petal.Length, Sepal.Length.Squared, Not.Setosa, Petal.Width.Interaction) %>%
  arrange(Petal.Length)
```

The end result is the following data

```{r echo=FALSE, message=FALSE}
library(kableExtra)
df %>%
  head(10) %>%
  kable() %>%
  kable_styling()
```

## Visualization

It's a bit harder to visualize this model because of the number of dimensions but here's the clearest viz I was able to work up:

```{r echo = FALSE}
library(ggplot2)
df %>%
  mutate(Not.Setosa = ifelse(Not.Setosa == 1, "Yes", "No")) %>%
  mutate(Not.Setosa = as.factor(Not.Setosa)) %>%
  ggplot(aes(Sepal.Length.Squared, Petal.Length, shape=Not.Setosa, color=Petal.Width.Interaction)) +
  geom_point() +
  ylab("Petal Length") +
  xlab("Sepal Length Squared") +
  scale_color_distiller(palette = "Spectral")
```

## Model Creation

Now to create the multivariate model

```{r}
model <- lm(Petal.Length ~ ., df)
```

## Model Evaluation

```{r comment=NA}
summary(model)
```

This model preformed very well.  The adjusted R-squared is `r summary(model)$adj.r.squared` which is phenominal!  All of the terms are statistically significant.

### Residual Analysis

Let's take a look at the residuals

```{r, comment=NA}
summary(resid(model))
```

The mean and median are close to zero which is good.  All of the other data hint that the residuals are fairly normally distributed.  Let's take a look at the plots:

```{r, echo=FALSE}
library(ggfortify)
autoplot(model)
```

There are a couple observations that are outliers but generally everything looks great!

## Conclusion

This model explains almost all of the variablitiy.  The R-squared is about 0.97.  The residual plots do not raise any concern.  The Q-Q is fairly linear.  The coefficients are statistically significant.