### 2019-04-23 DATA605/MSilva_Assign13.Rmd
Completed Question 3

---
title: "DATA 605 Week 13"
author: "Mike Silva"
date: "April 23, 2019"
output: 
  rmdformats::readthedown:
    highlight: kate
---

## Question 1
**Use integration by substitution to solve the integral below.**

$$ \int{4e^{-7x}dx}$$

The integral of a constant times a function is the constant times the integral of the function so

$$ \int{4e^{-7x}dx} \rightarrow 4 \int{e^{-7x}dx}$$

Let u = -7x

$$ du = -7dx \rightarrow \frac{du}{-7}=dx\rightarrow dx = -\frac{1}{7}du$$

Substituded in

$$ 4 \int{e^{-7x}dx} \rightarrow 4\int{-\frac{e^u}{7}du} \rightarrow -\frac{4}{7}\int{e^u du}$$

Since  $\int{e^u du} = e^u$ we get

$$ -\frac{4}{7}\int{e^u du} \rightarrow -\frac{4}{7}e^u$$

Substitute for u we get

$$ -\frac{4}{7}e^{-7x} + C$$

## Question 2

**Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac{dN}{dt}=-\frac{3150}{t^4}-220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N(t) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.**

$$\int{-\frac{3150}{t^4}-220 dt}\rightarrow-220t-3150\frac{1}{-3t^3}\rightarrow-220t+\frac{1050}{t^3} + C$$

To solve for the constant we solve for $C$ given that $N(t) = 6530$

$$6530 = -220(1)+\frac{1050}{(1)^3} + C \rightarrow 6530 = -220 + 1050 + C \rightarrow 6530 = 830 + C \rightarrow 5700 = C$$

The function is $N(t)=\frac{1050}{t^3}-220t+5700$

## Question 3

**Find the total area of the red rectangles in the figure below, where the equation of the line is $f(x) = 2x-9$.**

![](MSilva_Assign13_Q3.png)

The area under the line is a triangle from 4.5 to 8.5.  The height is found by pluging 8.5 into the function.

$$f(8.5)=2(8.5)-9=17-9=8$$

The area of the triangle is 1/2 base x height so

$$\frac{1}{2}(8.5 - 4.5) \times 8 = \frac{1}{2}(4) \times 8 = 2 \times 8 = 16$$

So the area is 16.  Verifying with R

```{r, comment=NA}
fx <- function(x){
  2*x-9
}
integrate(fx, 4.5, 8.5)
```

## Question 4

**Find the area of the region bounded by the graphs of the given equations **

$$y=x^2-2x-2, y=x+2$$

## Question 5

**A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.**

## Question 6

**Use integration by parts to solve the integral below.**

$$\int{ln(9x) \times x^6 dx}$$

## Question 7

**Determine whether f(x) is a probability density function on the interval $[1, e^6]$ . If not, determine the value of the definite integral.**

$$f(x)=\frac{1}{6x}$$