### 2019-05-08 DATA605/MSilva_Assign15.Rmd
More work on Assigment 15

---
title: "DATA 605 Week 15"
author: "Mike Silva"
date: "May 6, 2019"
output: html_document
---

## Question 1

Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

Given the set of (X, Y) points we need to find the best fit line $\hat{Y}= a + bX$ such that the sum of square errors in Y ($\sum{(Y-\hat{Y})^2}$) is minimized.

$$SSE=\sum_{i=1}^{n}{(Y_i-\hat{Y})^2} = \sum_{i=1}^{n}{(Y_i-a - bX_i)^2}$$

SSE will be minimized when $\frac{d{SSE}}{da} = 0$ and $\frac{d{SSE}}{db} = 0$.

$$\frac{d{SSE}}{da} = \sum_{i=1}^{n}{-2(Y_i-a - bX_i)}=2\bigg(na+b\sum_{i=1}^{n}{X_i}-\sum_{i=1}^{n}{Y_i}\bigg)=0$$

Dividing by 2 and solving for a

$$a = \bar{Y}-b\bar{X}$$

$$\frac{d{SSE}}{db} = \sum_{i=1}^{n}{-2X_i(Y_i-a - bX_i)}=\sum_{i=1}^{n}{-2(X_iY_i-aX_i-bX_i^2)}=0$$

Substituting in a gives

$$\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y}-bX_i\bar{X}-bX_i^2)}=0 \rightarrow \sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}-b\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}=0$$

Solving for b gives

$$b=\frac{\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}}{\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}}$$

```{r}
X <- c(5.6, 6.3, 7, 7.7, 8.4)
Y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

X_bar <- mean(X)
Y_bar <- mean(Y)

b <- sum(X * Y - X * Y_bar) / sum(X ^ 2 - X * X_bar)

a <- Y_bar - b * X_bar
```

The least squares line is $\hat{Y}$ = `r a` + `r b` $X$

Double checking my work
```{r}
lm(Y ~ X)
```

## Question 2

Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.

$$f(x,y)=24x-6xy^2-8y^3$$

## Question 3

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81  21x + 17y units of the "house" brand and 40 + 11x  23y units of the "name" brand.

**Step 1.** Find the revenue function R ( x, y ).

**Step 2.** What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

## Question 4

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x,y)=\frac{1}{6}x^2+\frac{1}{6}y^2+7x+25y+700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

## Question 5

Evaluate the double integral on the given region.

$\int{\int{(e^{8x+3y})dA}}$; $R: 2 \leq x \leq 4$ and $2 \leq y \leq 4$

Write your answer in exact form without decimals.