### 2019-05-24 DATA605/MSilva_Final.Rmd
More work on the final

---
title: "CUNY SPS DATA 605 Final"
author: "by Mike Silva"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

**NOTE:** The instructions/questions are in **bold**. My responses and comments are in *italics*.

```{r, message=FALSE, warning=FALSE}
library(randomForest)
library(caret)
library(MASS)
library(Matrix)
library(matlib)
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

## Problem 1

**Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6. Then generate a random variable Y that has 10,000 random normal numbers with a mean of $\mu=\sigma=\frac{N+1}{2}$.**

```{r, echo=TRUE, eval=TRUE, comment=NA}
set.seed(12345)
N <- 10
n <- 10000
mu <- sigma <- (N + 1)/2
df <- data.frame(X = runif(n, min=1, max=N), 
                 Y = rnorm(n, mean=mu, sd=sigma))
summary(df$X)
summary(df$Y)
```

### Probability

**Calculate as a minimum the below probabilities a through c. Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable. Interpret the meaning of all probabilities.**

```{r echo=TRUE, eval=TRUE}
x <- median(df$X)
y <- as.numeric(quantile(df$Y)["25%"])
```

*Small letter "x" is `r x` and small letter "y" is `r y`.*

a. $P(X>x | X>y)$

```{r echo=TRUE, eval=TRUE}
P_a_and_b <- df %>%
  filter(X > x,
         X > y) %>%
  nrow() / n

P_b <- df %>%
  filter(X > y) %>%
  nrow() / n

problem_1a <- P_a_and_b / P_b
```

*The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` given that it is greater than `r y` is `r problem_1a`*

b. $P(X>x, Y>y)$

```{r echo=TRUE, eval=TRUE}
problem_1b <- df %>%
  filter(X > x,
         Y > y) %>%
  nrow() / n
```

*The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` and a random normally distributed number with a mean and standard deviation of `r mu` being greater than `r y` is `r problem_1b`*

c. $P(X<x | X>y)$

```{r echo=TRUE, eval=TRUE}
P_a_and_b <- df %>%
  filter(X < x,
         X > y) %>%
  nrow() / n

P_b <- df %>%
  filter(X > y) %>%
  nrow() / n

problem_1c <- P_a_and_b / P_b
```

*The probablity of a random number uniformly ranging from 1 to `r N` being less than `r x` given that it is greater than `r y` is `r problem_1c`*

**Investigate whether P(X>x and Y>y)=P(X>x)P(Y>y) by building a table and evaluating the marginal and joint probabilities.**

```{r echo=TRUE, eval=TRUE}
# Create Joint Probabilities
temp <- df %>%
  mutate(A = ifelse(X > x, " X greater than x", " X not greater than x"),
         B = ifelse(Y > y, " Y greater than y", " Y not greater than y")) %>%
  group_by(A, B) %>%
  summarise(count = n()) %>%
  mutate(probability = count / n)

# Create Marginal Probabilities
temp <- temp %>%
  ungroup() %>%
  group_by(A) %>%
  summarise(count = sum(count),
            probability = sum(probability)) %>%
  mutate(B = "Total") %>%
  bind_rows(temp)

temp <- temp %>%
  ungroup() %>%
  group_by(B) %>%
  summarise(count = sum(count),
            probability = sum(probability)) %>%
  mutate(A = "Total") %>%
  bind_rows(temp)

# Create Table
temp %>%
  select(-count) %>%
  spread(A, probability) %>%
  rename(" " = B) %>%
  kable() %>%
  kable_styling()
```

*P(X>x and Y>y) is `r temp[temp$A == " X greater than x" & temp$B == " Y greater than y",]$probability`. P(X>x)P(Y>y) is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability` $\times$ `r temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability` which is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability * temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability`.  They are not the same.*

**Check to see if independence holds by using Fisher's Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?**

```{r, echo=TRUE, eval=TRUE, comment=NA}
count_data <- temp %>%
  filter(A != "Total",
         B != "Total") %>%
  select(-probability) %>%
  spread(A, count) %>%
  as.data.frame()

row.names(count_data) <- count_data$B

count_data <- count_data %>%
  select(-B) %>%
  as.matrix()

fisher.test(count_data)
```

```{r, echo=TRUE, eval=TRUE, comment=NA}
chisq.test(count_data)
```

*Fisher's Exact Test is for is used when you have small cell sizes (less than 5).  The Chi Square Test is used when the cell sizes are large.  It would be appropriate in this case.*

## Problem 2

**You are to register for Kaggle.com (free) and compete in the House Prices: Advanced Regression Techniques competition. https://www.kaggle.com/c/house-prices-advanced-regression-techniques. I want you to do the following.**

```{r, echo=TRUE, eval=TRUE}
kaggle <- read.csv("Kaggle_House_Prices/train.csv") %>%
  # Removing outliers per http://jse.amstat.org/v19n3/decock.pdf
  filter(GrLivArea < 4000) %>% 
  # Filling in missing values with zeros where it makes sense
  mutate(BedroomAbvGr = replace_na(BedroomAbvGr, 0),
         BsmtFullBath = replace_na(BsmtFullBath, 0),
         BsmtHalfBath = replace_na(BsmtHalfBath, 0),
         BsmtUnfSF = replace_na(BsmtUnfSF, 0),
         EnclosedPorch = replace_na(EnclosedPorch, 0),
         Fireplaces = replace_na(Fireplaces, 0),
         GarageArea = replace_na(GarageArea, 0),
         GarageCars = replace_na(GarageCars, 0),
         HalfBath = replace_na(HalfBath, 0),
         KitchenAbvGr = replace_na(KitchenAbvGr, 0),
         LotFrontage = replace_na(LotFrontage, 0),
         OpenPorchSF = replace_na(OpenPorchSF, 0),
         PoolArea = replace_na(PoolArea, 0),
         ScreenPorch = replace_na(ScreenPorch, 0),
         TotRmsAbvGrd = replace_na(TotRmsAbvGrd, 0),
         WoodDeckSF = replace_na(WoodDeckSF, 0)) 
```

### Descriptive and Inferential Statistics

**Provide univariate descriptive statistics and appropriate plots for the training data set.**

```{r stats_and_plots, echo=FALSE, eval=TRUE, results='asis', warning=FALSE}
for (variable in  names(kaggle)){
  cat(paste("<h4>", variable, "</h4>"))

  d <- kaggle[,names(kaggle) == variable]

  if(is.factor(d)){
    p <- ggplot(kaggle, aes_string(variable)) + 
      geom_bar() + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

    d_table <- kaggle %>%
      group_by(.dots = variable) %>%
      summarise(count = n()) %>%
      kable() %>%
      kable_styling()
  } else {
    p <- ggplot(kaggle, aes_string(variable)) + 
      geom_density()
    q <- quantile(d, na.rm = TRUE)
    d_table <- data.frame(Min = min(d, na.rm = TRUE),
                          First = q[2],
                          Median = median(d, na.rm = TRUE), 
                          Mean = mean(d, na.rm = TRUE),
                          Third = q[4],
                          Max = max(d, na.rm = TRUE)) %>%
      rename("First Quartile" = First,
             "Third Quartile" = Third) %>%
      gather(Measure, Value) %>%
      kable() %>%
      kable_styling()
  }
  print(p)
  print(d_table)
}
```

**Provide a scatterplot matrix for at least two of the independent variables and the dependent variable.**

```{r, echo=TRUE}
kaggle %>%
  select(TotalBsmtSF, GrLivArea, YearBuilt, SalePrice) %>%
  pairs()
```

**Derive a correlation matrix for any three quantitative variables in the dataset.**

```{r, echo=TRUE}
correlation_matrix <- kaggle %>%
  select(TotalBsmtSF, GrLivArea, YearBuilt) %>%
  cor() %>%
  as.matrix()

correlation_matrix %>%
  kable() %>%
  kable_styling()
```

**Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.**

```{r, results="asis"}
zero_vars <- c()

variables <- kaggle %>%
  select(-SalePrice) %>%
  names()

cat("<table><thead><tr><th>Variable</th><th>Correlation Equal to Zero</th></tr></thread><tbody>")

for(variable in variables){
  d <- kaggle[,names(kaggle) == variable]
  if(is.numeric(d)){
    results <- cor.test(kaggle$SalePrice, d, conf.level = 0.8)
    if(0 > results$conf.int[1] & results$conf.int[2] > 0){
      hypothesis_test_results <- "Yes"
      zero_vars <- c(zero_vars, variable)
    } else {
      hypothesis_test_results <- "No"
    }
    cat(paste("<tr><td>",variable,"</td><td>", hypothesis_test_results, "</td>"))
  }
}
cat("</tbody></table>")
```

**Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?**

*There are `r length(zero_vars)` variables that have no correlation with the sale price.  I would be worried be woried about family-wise error because I just ran `r length(variable)` hypothesis tests.  The family-wise error rate would be: ${FWER} = 1 - (1 - .2)^`r length(zero_vars)` = 1 - `r .8 ^ length(zero_vars)` = `r 1 - (.8^ length(zero_vars))`$.  This means there is a really high probability of committing a Type I error.*

### Linear Algebra and Correlation

**Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.)**

```{r}
precision_matrix <- inv(correlation_matrix)

precision_matrix %>%
  kable() %>%
  kable_styling()
```

**Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix.**

*This will generate the identity matrix.*

```{r}
correlation_matrix %*% precision_matrix %>%
  round() %>%
  kable() %>%
  kable_styling()
```

*This will too.*

```{r}
precision_matrix %*% correlation_matrix %>%
  round() %>%
  kable() %>%
  kable_styling()
```

**Conduct LU decomposition on the matrix.**

```{r}
lu_decomposition <- Matrix::expand(lu(correlation_matrix))
```

*The LU decomposition should yield the correlation matrix after multiplying the two components.  In other words this:*

```{r}
A <- lu_decomposition$L %*% lu_decomposition$U %>%
  as.matrix() 
colnames(A) <- colnames(correlation_matrix)
rownames(A) <- rownames(correlation_matrix)

A %>%
  kable() %>%
  kable_styling()
```

*Should match this:*

```{r}
correlation_matrix %>%
  kable() %>%
  kable_styling()
```

*Which it does.*

### Calculus-Based Probability & Statistics

**Many times, it makes sense to fit a closed form distribution to data. Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary. Then load the MASS package and run fitdistr to fit an exponential probability density function. (See https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html). Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)).**

*I will be using the Square Footage of the Unfinished Basement for this part of the exam.  I know it is right skewed because the mean (`r mean(kaggle$BsmtUnfSF)`) is larger than the median (`r median(kaggle$BsmtUnfSF)`) as shown in the figure below:*

```{r, echo=FALSE}
ggplot(kaggle, aes(BsmtUnfSF)) + geom_density() + 
  geom_vline(xintercept = mean(kaggle$BsmtUnfSF), color = "red")+ 
  geom_vline(xintercept = median(kaggle$BsmtUnfSF), color = "blue") +
  annotate("text", x = 700, y = 0.0005, label = "Mean", color="red") + 
  annotate("text", x = 300, y = 0.0005, label = "Median", color="blue")
```

*The minimum (`r min(kaggle$BsmtUnfSF)`) is not smaller than zero so we don't need to shift the data.  Remember that the maximum value is `r max(kaggle$BsmtUnfSF)`.  I have already loaded the MASS package.*

```{r}
lambda <- fitdistr(kaggle$BsmtUnfSF, densfun = "exponential")$estimate
```

*The optimal value for $\lambda$ is `r lambda`.*

```{r}
samples <- rexp(10000, rate = lambda)
```

**Plot a histogram and compare it with a histogram of your original variable.**

```{r, echo=FALSE}
histogram_data <- data.frame(Value = samples, Data = c("Sample"))
histogram_data <- data.frame(Value = kaggle$BsmtUnfSF, Data = c("Original")) %>%
  rbind(histogram_data) %>%
  mutate(Data = as.factor(Data))

ggplot(histogram_data, aes(Value, fill = Data)) + geom_histogram(bins = 50, alpha = 0.5)
```

*The original data is not quite exponential, but it's not a bad approximation.  The exponential distribution data has a longer tail than the basement square footage data.*

**Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF).**

*The PDF would be $f(x;\lambda) = \lambda e^{-\lambda x}$ where $x \geq 0$ and otherwise zero.*

*The CDF would be $F(x;\lambda) = 1 - e^{-\lambda x}$.  lambda is `r lambda`. To find the 5th percentile we need to solve for x in:*

$$0.05 = 1 - e^{-`r lambda` x}$$

$$-0.95 = - e^{-`r lambda` x}$$

$$-ln(0.95) = `r lambda` x$$

$$x = \frac{-ln(0.95)}{`r lambda`} = `r -log(0.95)/lambda`$$
*For the 95th percentile we need to solve for x in:*

$$0.95 = 1 - e^{-`r lambda` x}$$

$$-0.05 = - e^{-`r lambda` x}$$

$$-ln(0.05) = `r lambda` x$$

$$x = \frac{-ln(0.05)}{`r lambda`} = `r -log(0.05)/lambda`$$

*So the 5th and 95th percentiles are approximately `r round(-log(0.95)/lambda,0)` and `r round(-log(0.05)/lambda,0)`, respectively.*

**Also generate a 95% confidence interval from the empirical data, assuming normality.**

```{r, comment=NA}
mu <- mean(kaggle$BsmtUnfSF)
s <- sd(kaggle$BsmtUnfSF)
n <- nrow(kaggle)
error <- qnorm(0.975) * s / sqrt(n)

ci <- c(mu - error, mu + error)
names(ci) <- c("5%", "95%")
ci
```

*The 95% confidence interval (assuming normality) is `r round(ci[1])` to `r round(ci[2])`. Although what I think was meant was assume the data is normally distributed.  Calculate the 5th and 95th percentile.  That is found using $x = \mu + Z \sigma$.  The Z is -1.645 for the 5th percentile, and 1.645 for the 95th.  So the percentiles would be `r round(mu + (-1.645 * s), 0)` and `r round(mu + (1.645 * s), 0)`.*

**Finally, provide the empirical 5th percentile and 95th percentile of the data. Discuss.**

```{r, comment=NA}
quantile(kaggle$BsmtUnfSF, c(0.05, 0.95))
```

*The actual 5th percentile is `r quantile(kaggle$BsmtUnfSF, c(0.05))` and the 95th is `r quantile(kaggle$BsmtUnfSF, c(0.95))`.  So the findings are summarized in the following table:*

| Method | 5% | 95% |
| ------ |:--:|:---:|
| Exponential CDF | `r round(-log(0.95)/lambda,0)` | `r round(-log(0.05)/lambda,0)` |
| Normal 95% CI | `r round(ci[1])`      | `r round(ci[2])` |
| Normal Percentiles | `r round(mu + (-1.645 * s), 0)`  | `r round(mu + (1.645 * s), 0)` | 
| Emperical Percentiles | `r round(quantile(kaggle$BsmtUnfSF, c(0.05)),0)`       | `r round(quantile(kaggle$BsmtUnfSF, c(0.95)),0)` |

*If we model the data as exponentially distributed the 5th percentile is `r round(-log(0.95)/lambda,0)`.  If we model it as normally distributed the 5th is at `r round(mu + (-1.645 * s), 0)` which in the context of square footage does not make any sense. The actual 5th percentile is `r round(quantile(kaggle$BsmtUnfSF, c(0.05)),0)`. The difference is explained to the assumed shape/distibution of the underlying data.*

*Looking at the 95th percentile we have `r round(-log(0.5)/lambda,0)` if the data are exponentially distributed, `r round(mu + (1.645 * s), 0)` if it is normally distributed and `r round(quantile(kaggle$BsmtUnfSF, c(0.95)),0)` in reality.  Again the difference is due to the assumed shape.*

*I have left out the 95% CI from the discussion because the confidence interval is a way of estimating the mean of the population.  We would know if we took 100 estimates that the actual mean falls within the confidence intervals 95% of the time.  Within this context it is meaningless.*

### Modeling

**Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score.**

*Since there are so many variables I will be using a random forest to detect which variables are the most important and will be using them to build the model.  First I will drop out the variables that are missing a lot of data or otherwise meaningless.*

```{r rf, cache=TRUE, message=FALSE, warning=FALSE}
set.seed(12345)

rf_data <- kaggle %>%
  # Dropping variables with a lot of missing values
  select(-Alley, -MasVnrType, -MasVnrArea, -BsmtQual, -BsmtCond, -BsmtExposure, -BsmtFinType1, -BsmtFinType2, -Electrical, -FireplaceQu, -GarageType, -GarageYrBlt, -GarageFinish, -GarageQual, -GarageCond, -PoolQC, -Fence, -MiscFeature) %>%
  # Dropping the ID because it is meaningless
  select(-Id)

ctrl <- trainControl(method = "cv", number = 5)
rf_fit <- train(SalePrice ~ ., 
                data = rf_data, 
                method = "rf", 
                ntree = 1000, 
                trControl = ctrl, 
                tuneGrid = data.frame(mtry = 10), 
                importance = TRUE, 
                verbose = TRUE)
```

*Now we can examine the variables to in order of importance:*

```{r rf_importance}
imp_var <- importance(rf_fit$finalModel) 
variables <- rownames(imp_var)

imp_var <- imp_var %>%
  as.data.frame()

names(imp_var) <- c("IncMSE", "IncNodePurity")
imp_var$Variable <- variables

imp_var %>%
  select(Variable, IncMSE, IncNodePurity) %>%
  arrange(-IncMSE) %>%
  kable() %>%
  kable_styling()
```

*The takeaways from the preceeding table is that size seems to consistently rise to the top as a variable to include in the model.  I don't want to have multiple variables that are measuring the same thing (i.e. Total SF of Living Area, Total SF of Basement, etc.).  I will use this table to select the variables that I think should be included in the linear regression model.*

```{r model, comment=NA}
fit <- lm(SalePrice ~ GrLivArea + GarageArea + LotArea + YearRemodAdd + Fireplaces + YearBuilt + OverallQual, kaggle)
summary(fit)
```

*These seven variables explain 80% of the variability in the sale price.  I will use this model to make the Kaggle submission.*

```{r predict, comment=NA}
test <- read.csv("Kaggle_House_Prices/test.csv")
y_hat <- predict(fit, test)
submission <- data.frame(Id = test$Id, SalePrice = y_hat)
summary(submission)
```

*It appears there is an NA in the submission.  In the interest of getting my final turned in on time I will be using the median to plug that hole.*

```{r}
fill <- median(submission$SalePrice, na.rm = TRUE)
submission <- submission %>%
  mutate(SalePrice = replace_na(SalePrice, fill))
write.csv(submission, "submission.csv", quote = FALSE, row.names =  FALSE)
```

*My Kaggle score is* **0.47768** *and my username is* **mikesilva***.  I am ranked 4520 on the leaderboard.  I will try to improve this score when I have time.*

*This analysis is available on  [GitHub](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA605/MSilva_Final.Rmd).*