### 2019-08-30 DATA621/Week 1 Textbook Exercises.Rmd
Initial commit of Week 1 Textbook Exercises

---
title: "DATA 612 Week #1 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## LMR 1.1

The dataset *teengam* concers a study of teenage gambline in Britain.  Make a numerical and graphical summary of the data, commenting on any features that you find interesting.  Limit the output you present to a quantity that a busy reader would find sufficeint to get a basic understanding of the data.

## LMR 1.3

The dataset *prostate* is from a study on 97 men with prostate conacer who were due to receive a radical prostatectomy.  Make a numerical and graphical summary of the data as in the first question.

## LMR 1.4

The dataset *sat* comes from a study entitled "Getting What You Pay For: The Debate Overty Equity in Public School Expenditures."  Make a numerical and graphical summary of the data as in the first question.

## LMR 1.5

The dataset *divusa* contains data on divorces in the United States from 1920 to 1996.  Make a numerical and graphical summary of the data as in the first question.