### 2019-10-04 DATA621/Week 06 Textbook Exercises.Rmd
Updated Week 06 Textbook Exercises

---
title: "DATA 621 Week #6 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    code_folding: "hide"
---

```{r knitr_init, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(faraway)
library(ggthemes)
## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

## MARR 6.3

## MARR 6.5

## LMR 7.3

Using the `divusa` data:

a) Fit a regression model with `divorce` as the response and `unemployed`, `femlab`, `marriage`, `birth` and `military` as predictors. Compute the condition numbers and interpret their meanings.

```{r, null_prefix=TRUE}
divusa <- read.delim("data/divusa.txt")
fit <- lm(divorce ~ unemployed + femlab + marriage + birth + military, data = divusa)
summary(fit)
x <- model.matrix(fit)[,-1]
e <- eigen(t(x) %*% x)
sqrt(e$val[1]/e$val)
```

b) For the same model, compute the VIFs.  Is there evidence that collinearity causes some predictors not to be significant? Explain.

```{r, null_prefix=TRUE}
vif(x)
```

c) Does the removal of insignificant predictors from the model reduce the collinearity? Investigate.

```{r, null_prefix=TRUE}
fit <- lm(divorce ~ femlab + marriage + birth, data = divusa)
summary(fit)
x <- model.matrix(fit)[,-1]
e <- eigen(t(x) %*% x)
sqrt(e$val[1]/e$val)
vif(x)
```

## LMR 7.8