### 2019-10-31 DATA621/HW3.Rmd
Initial commit of DATA 621 HW3

---
title: "DATA 621 Homework #3"
subtitle: "From Work by Critical Thinking Group 3"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
library(Amelia)
library(pROC)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv('data/crime-training-data_modified.csv')
evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
```

## Data Exploration

### Are There Missing Values?
First we look at the data to see if any variables have missing data:

```{r}
missmap(df, main = "Missing vs Observed Values")
```

It looks like we have a complete data set.  No need to impute values.

### Splitting the Data
Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
```{r}
set.seed(42)
train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Exploratory Data Analysis
Now, let's look at our training data. By looking at a correlation matrix, we can see which variables may be too correlated to be included together in a model as predictor variables. This will help us later during the model selection process.

```{r}
train %>% 
  cor(.) %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```

Next we will look at each potential predictor and how it is distributed across the target variable.

The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e. blue) and a negative target variable (areas with crime rates below the median, i.e. red). What we are looking for is variables that show way to split data into two groups.

```{r}
for (var in names(train)){
  if(var != "target"){
    plot_df <- train
    plot_df$x <- plot_df[,var]
    p <- ggplot(plot_df, aes(x, color = factor(target))) +
      geom_density() +
      theme_light() +
      ggtitle(var) +
      scale_color_brewer(palette = "Set1") +
      theme(legend.position = "none",
            axis.title.y = element_blank(),
            axis.title.x = element_blank())
    print(p)
  }
}
```

Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.

## Basic Model Building

We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.

```{r}
baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
summary(baseline)
test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
                  deviance=baseline$deviance, 
                  r2 = 1 - baseline$deviance/baseline$null.deviance,
                  aic=baseline$aic)
cm
```

Our baseline model is ok with an F1 score of `r cm$byClass[7]`.

Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:

```{r}
fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
summary(fullmodel)
test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
results <- rbind(results,tibble(model = "fullmodel",
                                predictors = 12,F1 = cm$byClass[7],
                                deviance=fullmodel$deviance, 
                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
                                aic=fullmodel$aic))
cm
```

The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.

After some backward elimination of non-significant predictor variables, we arrive at the following model:

```{r}
model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
              family = binomial(link = "logit"), 
              train)
summary(model1)
test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
results <- rbind(results,tibble(model = "model1",
                                predictors = 6,F1 = cm$byClass[7],
                                deviance=model1$deviance, 
                                r2 = 1-model1$deviance/model1$null.deviance,
                                aic=model1$aic))
cm
```

## Alternative Models

### Feature Engineering
In this portion below we explored using the difference between the class distributions as a feature itself. So instead of the raw feature itself the we created density features that are essientially the differences in probability of the positve and negative classes at that certain point for the feature.

```{r}
library(fuzzyjoin)
density_diff <- function(data,eval_data,var,target,new_column){
  data[[paste0(var,"temp")]] <- data[[var]]
  eval_data[[paste0(var,"temp")]] <- eval_data[[var]]

  #standardize variable between 0 and 1
  data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))
  eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))

  ## Calculate density estimates
  g1 <- ggplot(data, aes(x=data[[var]], group=target, colour=target)) +
    geom_density(data = data) + xlim(min(data[[var]]), max(data[[var]]))
  gg1 <- ggplot_build(g1)
  # construct the dataset
  x <- gg1$data[[1]]$x[gg1$data[[1]]$group == 1]
  y1 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 1]
  y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
  df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), side=(y1<y2), ydiff = y2-y1)
  ##creating the second graph object
  g3 <- ggplot(df2) +
    geom_line(aes(x = x, y = ydiff, colour = side)) +
    geom_area(aes(x = x, y = ydiff, fill = side, alpha = 0.4)) +
    guides(alpha = FALSE, fill = FALSE)

  data$join <- data[[var]]
  df2$join <- df2$x
  temp <- difference_left_join(data,df2, by="join", max_dist =.001)
  means <- aggregate(ydiff ~ temp$join.x, temp, mean)

  colnames(means) <- c("std_var","density_diff")

  new_data <- merge(means,data, by.x ="std_var", by.y =var)
  new_data$std_var <- NULL
  new_data$join <- NULL
  new_data[new_column] <- new_data$density_diff
  new_data$density_diff <- NULL
  ###################same thing with eval data ############################
  eval_data$join <- eval_data[[var]]
  df2$join <- df2$x
  temp <- difference_left_join(eval_data,df2, by="join", max_dist =.001)
  means <- aggregate(ydiff ~ temp$join.x, temp, mean)

  ##new stuff
  colnames(means) <- c("std_var","density_diff")
  #data["key"] <- (var - min(var))/(max(var) - min(var))

  eval_data2 <- merge(means,eval_data, by.x ="std_var", by.y =var)
  eval_data2$std_var <- NULL
  eval_data2$join <- NULL
  eval_data2[new_column] <- eval_data2$density_diff
  eval_data2$density_diff <- NULL

  eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
  eval_data2[[paste0(var,"temp")]] <- NULL

  new_data[[var]] <- new_data[[paste0(var,"temp")]]
  new_data[[paste0(var,"temp")]] <- NULL

  mylist <- list(g1,g3,new_data,eval_data2)
  names(mylist)<- c("dist","dist_diff","new_data_training","new_data_eval")
  return(mylist)
}
```

### Creating Some New Variables

```{r}
str(train)
str(test)
## try it out 
train <- density_diff(train,test,"nox",train$target,"nox_density")$new_data_training
train <- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
train <- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
train <- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
train <- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
train <- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
train <- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
train <- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
train <- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training

test<- density_diff(train,test,"nox",train$target,"nox_density")$new_data_eval
test<- density_diff(train,test,"age",train$target,"age_density")$new_data_eval 
test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_eval
test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_eval
test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_eval
test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_eval
test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_eval
test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_eval
test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_eval

str(train)
str(test)
```

### Example of the new meta feature
The first graph below is the difference between the distributions and the second graph is the new derived predictive feature
```{r}
density_diff(train,test,"age",train$target,"age_density")$dist
density_diff(train,test,"age",train$target,"age_density")$dist_diff
```

With our new density variables, we can run another model:

```{r}
density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
summary(density_models)
test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
test$density_models_yhat <- predict(density_models, test)
cm <- confusionMatrix(factor(test$density_models), factor(test$target))
results <- rbind(results,tibble(model = "density models",
                                predictors = 9,
                                F1 = cm$byClass[7],
                                deviance = density_models$deviance,
                                r2 = 1 - density_models$deviance / density_models$null.deviance,
                                aic = density_models$aic))
cm
```

## Model Selection

Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.

```{r}
kable(results)
```

Looking the three models together, `model1` looks like the best one. Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.

### pROC Output

ROC curves can give us another look at which model might be better suited for prediction.

#### Baseline
```{r}
par(pty = "s")
roc(test[["target"]], test[["baseline"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Full Model
```{r}
par(pty = "s")
roc(test[["target"]], test[["fullmodel"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Model1
```{r}
par(pty = "s")
roc(test[["target"]], test[["model1"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Density Model 
```{r}
par(pty = "s")
roc(test[["target"]], test[["density_models_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```