### 2020-02-08 DATA624/HW2.Rmd
Initial commit of Homework 2

---
title: "DATA 624 Homework 2"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(fpp2)
```

## Question 3.1

For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance.

* `usnetelec`
* `usgdp`
* `mcopper`
* `enplanements`

## Question 3.2

Why is a Box-Cox transformation unhelpful for the `cangas` data?

## Question 3.3

What Box-Cox transformation would you select for your `retail` data?

## Question 3.8

For your `retail` time series:

Split the data into two parts using
```{r}
myts.train <- window(myts, end=c(2010,12))
myts.test <- window(myts, start=2011)
```

Check that your data have been split appropriately by producing the following plot.

```{r}
autoplot(myts) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")
```

Calculate forecasts using snaive applied to myts.train.

```{r}
fc <- snaive(myts.train)
```

Compare the accuracy of your forecasts against the actual values stored in myts.test.

```{r}
accuracy(fc,myts.test)
```

Check the residuals.

```{r}
checkresiduals(fc)
```

Do the residuals appear to be uncorrelated and normally distributed?

How sensitive are the accuracy measures to the training/test split?