### 2020-02-18 DATA624/HW3.Rmd
Initial commit

---
title: "DATA 624 Homework 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(fpp2)
library(forcats)
```

## Question 6.2

The `plastics` data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

* Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?

```{r}
autoplot(plastics)
```

***The `plastics` data has has an increasing trend and a cycle of 1 year***

* Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r}
decompose_plastics <- decompose(plastics, type="multiplicative")

decompose_plastics %>% 
  autoplot() +
  ggtitle("Multiplicative Decomposition of Plastic Product A Sales")
```

```{r, eval=FALSE, echo=FALSE}
trend_cycle_component <- ma(ma(plastics, order = 12, centre = TRUE), order = 2)
detrended_plastics <- plastics / trend_cycle_component
seasonal_component <- tapply(detrended_plastics, cycle(detrended_plastics), mean)

decompose(plastics)
```

* Do the results support the graphical interpretation from part a?

**Yes**

* Compute and plot the seasonally adjusted data.

```{r}
seasonally_adjusted_plastics <- plastics - decompose_plastics$seasonal
autoplot(seasonally_adjusted_plastics)
```

* Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

* Does it make any difference if the outlier is near the end rather than in the middle of the time series?

## Question 6.3

Recall your retail time series data (from Exercise 3 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?