### 2020-03-03 DATA624/Project 1-B Power.Rmd
Initial commit

---
title: "DATA 624 Project 1 Part B - Power"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
```

## Directions

Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.  Your assignment is to model these data and a monthly forecast for 2014.  The data is given in a single file.  The variable ‘KWH’ is power consumption in Kilowatt hours, the rest is straight forward.    Add this to your existing files above.