### 2020-03-18 DATA624/HW6.Rmd
More work on HW 6

---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(fpp2)
library(kableExtra)
library(gridExtra)
library(readxl)
library(tseries)
options(scipen=999)
```

## Question 8.1

Figure 8.31 shows the ACFs for 36 random numbers, 360 random numbers and 1,000 random numbers.

<img src ="https://otexts.com/fpp2/fpp_files/figure-html/wnacfplus-1.png" alt="Left: ACF for a white noise series of 36 numbers. Middle: ACF for a white noise series of 360 numbers. Right: ACF for a white noise series of 1,000 numbers.">

a. Explain the differences among these figures. Do they all indicate that the data are white noise?

**The ACF bands become narrower as the number of numbers increase.   There is no discernable pattern in the bars of the ACF charts and the bars are within the bands, which indicates the data are white noise.**

b. Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?

**It's related to the law of large numbers.  As the number of observations increase, the number of large outliers from the mean decreases.  We are more certain that a large observation is really an outlier with more data.**

## Question 8.2

A classic example of a non-stationary series is the daily closing IBM stock price series (data set `ibmclose`). Use R to plot the daily closing prices for IBM stock and the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.

```{r}
ggtsplot <- function(ts, title){
  # A ggplot2 version of tsdisplay(df)
  # Args:
  #    ts (Time-Series): The time series we want to plot
  #    title (str): The title of the graph
  grid.arrange( 
    autoplot(ts) +
      ggtitle(title) +
      theme(axis.title = element_blank()),
    grid.arrange(
      ggAcf(ts) + 
        ggtitle(element_blank()), 
      ggPacf(ts) + 
        ggtitle(element_blank()), ncol = 2), nrow = 2)
}

ggtsplot(ibmclose, "IBM Closing Stock Price")
```

**Time series with trends or with seasonality are not stationary.  It is clear from the top plot that there is a trend in IBM's stock price.  The ACF will drop to zero quickly for a stationary time series.  This ACF does not have that property.  The gradual decrease is a tell that it is not stationary.  The first PACF is almost one which is another sign that the data is not stationary.**

## Question 8.3

For the following series, find an appropriate Box-Cox transformation and order of differencing in order to obtain stationary data.

a. `usnetelec`

```{r}
autoplot(usnetelec, main = "US Net Electricity Generation") +
  theme(axis.title = element_blank())
```

**The trend looks nearly linear.  It will probably only need to be differenced to be stationary.  I will use the `ndiffs` function to determine the order of differencing.***

```{r}
ggtsplot(diff(usnetelec, ndiffs(usnetelec)), paste0("Differenced US Net Electricity Generation (lag=", ndiffs(usnetelec), ")"))
```

**It needed to be differenced by `r ndiffs(usnetelec)`, or a first order difference.  Now to check if the data is stationary.**

```{r}
is_stationary <- function(data){
  results <- kpss.test(data)
  if (results$p.value > 0.05){
    "Data IS Stationary"
  } else {
    "Data IS NOT Stationary"
  }
}

cat(is_stationary(diff(usnetelec, ndiffs(usnetelec))))
```

**The p-value of 0.36 indicates the data are stationary.**

b. `usgdp`

```{r}
autoplot(usgdp, main = "Quarterly US Gross Domestic Product") +
  theme(axis.title = element_blank())
```

**Since this series is dollars we probably need to log transform it.  It also looks like it needs to be differenced.**

```{r}
usgdp_lambda <- BoxCox.lambda(usgdp)
bc_usgdp <- BoxCox(usgdp, lambda = usgdp_lambda)
ggtsplot(diff(bc_usgdp, ndiffs(bc_usgdp)), paste0("Differenced Quarterly US GDP (lag=", ndiffs(bc_usgdp), ", lambda=",round(usgdp_lambda,2),")"))
```

**The Box Cox transformation with a lambda of `r round(usgdp_lambda,2)` and a differencing of `r ndiffs(bc_usgdp)`.**

```{r}
cat(is_stationary(diff(bc_usgdp, ndiffs(bc_usgdp))))
```

c. `mcopper`

```{r}
autoplot(mcopper, main = "Monthly Grade A Copper Prices") +
  theme(axis.title = element_blank())
```

**This has the spike in copper prices that happened late in the the 2000's.**

```{r}
mcopper_lambda <- BoxCox.lambda(mcopper)
bc_mcopper <- BoxCox(mcopper, lambda = mcopper_lambda)
ggtsplot(diff(bc_mcopper, ndiffs(bc_mcopper)), paste0("Differenced Monthly Grade A Copper Prices (lag=", ndiffs(bc_mcopper), ", lambda=",round(mcopper_lambda,2),")"))
```

**The Box Cox transformation with a lambda of `r round(mcopper_lambda,2)` and a differencing of `r ndiffs(bc_mcopper)`.**

```{r}
cat(is_stationary(diff(bc_mcopper, ndiffs(bc_mcopper))))
```

d. `enplanements`

```{r}
autoplot(enplanements, main = "Monthly US Domestic Enplanements") +
  theme(axis.title = element_blank())
```

**The variance is not constant so this series needs a Box-Cox transformation and differencing to be stationary.**

```{r}
enplanements_lambda <- BoxCox.lambda(enplanements)
bc_enplanements <- BoxCox(enplanements, lambda = enplanements_lambda)
ggtsplot(diff(bc_enplanements, ndiffs(bc_enplanements)), paste0("Differenced Monthly US Domestic Enplanements (lag=", ndiffs(bc_enplanements), ", lambda=",round(enplanements_lambda,2),")"))
```

**The Box Cox transformation with a lambda of `r round(enplanements_lambda,2)` and a differencing of `r ndiffs(bc_enplanements)` (1st order).**

```{r}
cat(is_stationary(diff(bc_enplanements, ndiffs(bc_enplanements))))
```

e. `visitors`

```{r}
autoplot(visitors, main = "Monthly Australian Overseas Visitors") +
  theme(axis.title = element_blank())
```

**This series also exhibits an increasing variance and trend so Box-Cox and differencing will be needed.**

```{r}
visitors_lambda <- BoxCox.lambda(visitors)
bc_visitors <- BoxCox(visitors, lambda = visitors_lambda)
ggtsplot(diff(bc_visitors, ndiffs(bc_visitors)), paste0("Differenced Monthly Australian Overseas Visitors (lag=", ndiffs(bc_visitors), ", lambda=",round(visitors_lambda,2),")"))
```

**The Box Cox transformation had a lambda of `r round(visitors_lambda,2)` and a differencing of `r ndiffs(bc_visitors)` (1st order).**

```{r}
cat(is_stationary(diff(bc_visitors, ndiffs(bc_visitors))))
```

## Question 8.5

For your retail data (from Exercise 3 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.

```{r}
retaildata <- read_excel("retail.xlsx", skip = 1)
retail <- ts(retaildata[, "A3349873A"], frequency = 12, start = c(1982, 4))
retail_lambda <- BoxCox.lambda(retail)
bc_retail <- BoxCox(retail, lambda = retail_lambda)
ggtsplot(diff(bc_retail, ndiffs(bc_retail)), paste0("Differenced Monthly Retail Sales (lag=", ndiffs(bc_retail), ", lambda=",round(retail_lambda,2),")"))
```

**The data is differenced with a first order difference and adjusted with a Box Cox transformation with a lambda of `r round(retail_lambda,2)`.  The result:**

```{r}
cat(is_stationary(diff(bc_retail, ndiffs(bc_retail))))
```

## Question 8.6

Use R to simulate and plot some data from simple ARIMA models.

a. Use the following R code to generate data from an AR(1) model with $ϕ_1 = 0.6$ and $σ_2 = 1$.  The process starts with $y_1=0$.

```{r}
y <- ts(numeric(100))
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.6*y[i-1] + e[i]
```

b. Produce a time plot for the series. How does the plot change as you change $ϕ_1$?

c. Write your own code to generate data from an MA(1) model with $θ_1=0.6$ and $σ_2=1$.

d. Produce a time plot for the series. How does the plot change as you change $θ_1$?

e. Generate data from an ARMA(1,1) model with $ϕ_1=0.6$, $θ_1=0.6$ and $σ_2=1$.

f. Generate data from an AR(2) model with $ϕ_1=−0.8$, $ϕ_2=0.3$ and $σ_2=1$. (Note that these parameters will give a non-stationary series.)

g. Graph the latter two series and compare them.

## Question 8.7

Consider `wmurders`, the number of women murdered each year (per 100,000 standard population) in the United States.

a. By studying appropriate graphs of the series in R, find an appropriate ARIMA(p,d,q) model for these data.

b. Should you include a constant in the model? Explain.

c. Write this model in terms of the backshift operator.

d. Fit the model using R and examine the residuals. Is the model satisfactory?

e. Forecast three times ahead. Check your forecasts by hand to make sure that you know how they have been calculated.

f. Create a plot of the series with forecasts and prediction intervals for the next three periods shown.

g. Does auto.arima() give the same model you have chosen? If not, which model do you think is better?