### 2020-03-20 DATA624/Project 1-B Power.Rmd
Work on Project 1-B

---
title: "DATA 624 Project 1 Part B - Power"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(readxl)
library(lubridate)
library(kableExtra)
library(fpp2)
library(forcats)
library(gridExtra)
options(scipen=999)
```

## Introduction

I am to forecast how much power will be consumed by residential customers in kilowatt hours for the next 12 months.  Data was provided for this project spaning January 1998 until December 2013.

## Data Exploration

I will begin by reading in the data and ploting the series.

```{r}
KWH <- read_excel("ResidentialCustomerForecastLoad-624.xlsx") %>%
  select(KWH) %>%
  ts(start = decimal_date(date("1998-01-01")), frequency = 12)

autoplot(KWH, main = "Residential Power Usage (KWH)") +
  theme(axis.title = element_blank())
```

```{r}
summary(KWH)
```

We have outliers and missing data.  We will clean them up with `tsclean` from the forcasts package, and replot the data with the ACF and PACF:

```{r}
KWH <- tsclean(KWH)

ggtsplot <- function(ts, title){
  # A ggplot2 version of tsdisplay(df)
  # Args:
  #    ts (Time-Series): The time series we want to plot
  #    title (str): The title of the graph
  grid.arrange( 
    autoplot(ts) +
      ggtitle(title) +
      theme(axis.title = element_blank()),
    grid.arrange(
      ggAcf(ts) + 
        ggtitle(element_blank()), 
      ggPacf(ts) + 
        ggtitle(element_blank()), ncol = 2), nrow = 2)
}

ggtsplot(KWH, "Cleaned Residential Power Usage (KWH)")
```

This looks much better.  There is a seasonal component that looks to be additive.  I will need to use a model that captures seasonality.

## Model Approach

Since the data is highly seasonal, I we will need to use an ARIMA or Holt-Winters model.  Let's fit an models using `auto.arima()` and `hw()` and visualize their projections.

```{r}
h <- 12
lambda <- BoxCox.lambda(KWH)
autoplot(KWH) +
  autolayer(hw(KWH, h=h), series="Holt-Winters") +
  autolayer(hw(KWH, h=h, lambda = lambda), series="Holt-Winters (Box-Cox Adjusted)") +
  autolayer(forecast(auto.arima(KWH), h=h), series="ARIMA") +
  facet_grid(.~series) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "none", axis.title = element_blank())
```

### Residual Analysis

All models do a good job picking up the seasonality.  Let's look at the residuals to see which model did the best job.

#### Holt-Winters Model

```{r}
hw_fit <- hw(KWH, h=h)
checkresiduals(hw_fit)
```

#### Adjusted Holt-Winters Model

```{r}
adj_hw_fit <- hw(KWH, h=h, lambda = lambda)
checkresiduals(adj_hw_fit)
```

#### ARIMA Model

```{r}
arima_fit <- auto.arima(KWH)
#autoplot(forecast(arima_fit, h=h))
checkresiduals(arima_fit)
```