### 2020-03-28 DATA624/Project 1-C Waterflow.Rmd

---
title: "DATA 624 Project 1 Part C - Waterflow"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(readxl)
library(fpp2)
library(forcats)
library(gridExtra)
library(openxlsx)
library(tseries)
library(scales)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
```

## Introduction

For this project I have data on water flows from two pipes.  The data are captured with timestamps.  I am to forecast a week forward.  In the instructions for this project I am given the following hint: For multiple recordings within an hour, take the mean.  I will be on the lookout for data under the hour.

## Data Exploration

I will begin by plotting the datasets:

```{r}
pipe_1 <- read_excel("Waterflow_Pipe1.xlsx") %>%
  mutate(`Date Time` = as.POSIXct(`Date Time` * (60 * 60 * 24), origin = "1899-12-30", tz = "GMT"))

pipe_2 <- read_excel("Waterflow_Pipe2.xlsx") %>%
  mutate(`Date Time` = as.POSIXct(`Date Time` * (60 * 60 * 24), origin = "1899-12-30", tz = "GMT"))

pipes <- pipe_1 %>%
  mutate(Pipe = "Pipe 1") %>%
  rbind(mutate(pipe_2, Pipe = "Pipe 2"))

ggplot(pipes, aes(`Date Time`, WaterFlow, color = Pipe)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  facet_wrap(. ~ Pipe) +
  theme(legend.position = "none", axis.title = element_blank())
```

The two pipes are quite different in terms of the data they offer.  Pipe 1 has a median waterflow of about `r round(median(pipe_1$WaterFlow))`, while pipe 2 has a median of `r round(median(pipe_2$WaterFlow))`.  Pipe 1 also has what looks like over a week of data, while pipe 2 has over a month.  Both datasets have 1000 observations.  This suggests that pipe one has more fine grain observations (i.e. multiple observations in a hour or minute) while pipe 2's data captures a longer period of time.  That would explain the difference in median and the length of time covered by the respective data sets.  This also is hinted at in the instructions.

### Closer Look at Pipe 1

I will look at the head and tail of pipe 1's data to get a sense of what's included:

```{r}
head(pipe_1) %>% kable() %>% kable_styling()
```

There is multiple observations per minute which confirms my suspicions.  I will need to create the averages per minute as directed.

### Closer Look at Pipe 2

I will also take a look at pipe 2's data to see if there is anything I need to be aware of:

```{r}
head(pipe_2) %>% kable() %>% kable_styling()
```

Hmmm.  Some of the timestamps are just a bit off from the hour mark.  I will need to clean that up.

## Data Cleanup

### Pipe 1

I will clean up the pipe 1 data by creating the averages per minute.

```{r}
pipe_1 <- pipe_1 %>%
  mutate(Date = as.Date(`Date Time`),
         Time = paste0(format(`Date Time`, "%H"), ":00:00")) %>%
  group_by(Date, Time) %>%
  summarise(WaterFlow = mean(WaterFlow)) %>%
  ungroup() %>%
  mutate(`Date Time` = as.POSIXct(paste(as.character(Date), Time)), format = "%Y-%m-%d %H:%M:%OS") %>%
  select(`Date Time`, WaterFlow)
```

After doing this I have `r nrow(pipe_1)` observations.

#### Sidebar: Is there any pattern in the usage by hour?

I want to pause right here to see if there is any pattern in the the water flow by hour.

```{r}
pipe_1 %>%
  mutate(Time = format(`Date Time`, "%H")) %>%
  ggplot(aes(Time, WaterFlow)) +
  geom_boxplot() +
  ggtitle("Pipe #1 Water Flow by Hour") +
  theme(axis.title = element_blank())
```

There is not a discernable pattern in the water flow by hour.  With that exploration done I will convert the data to a time series object and plot it:

```{r}
pipe_1_ts <- ts(pipe_1$WaterFlow, start = c(2015, 10, 23), frequency = 24)

is_stationary <- function(ts) {
  results <- kpss.test(ts)
  if (results$p.value > 0.05) {
    "data IS stationary"
  } else {
    "data is NOT stationary"
  }
}

ggtsplot <- function(ts, title) {
  # A ggplot2 version of tsdisplay()
  # Args:
  #    ts (Time-Series): The time series we want to plot
  #    title (str): The title of the graph
  grid.arrange(
    autoplot(ts) +
      scale_y_continuous(labels = comma) +
      ggtitle(paste0(title, " (", is_stationary(ts), ")")) +
      theme(axis.title = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank()),
    grid.arrange(
      ggAcf(ts) + ggtitle(element_blank()),
      ggPacf(ts) + ggtitle(element_blank()), ncol = 2)
    , nrow = 2)
}

ggtsplot(pipe_1_ts, "Pipe #1 Waterflow")
```

This time series falls within the ACF and the data is stationary.  With stationary data I am done with this data set and can focus on pipe #2.

### Pipe 2

I will clean up pipe 2's `Date Time` field by rounding it to the nearest hour.

```{r}
pipe_2 <- pipe_2 %>%
  mutate(Date = as.Date(`Date Time`),
         Time = format(round(`Date Time`, units = "hours"), format = "%H:%M"),
         `Date Time` = as.POSIXct(paste(as.character(Date), Time)), format = "%Y-%m-%d %H:%M:%OS") %>%
  select(`Date Time`, WaterFlow)
```

#### Sidebar: Is there any pattern in the usage by hour?

I again want to see if there is a pattern in this pipe's usage by hour:

```{r}
pipe_2 %>%
  mutate(Time = format(`Date Time`, "%H")) %>%
  ggplot(aes(Time, WaterFlow)) +
  geom_boxplot() +
  ggtitle("Pipe #2 Water Flow by Hour") +
  theme(axis.title = element_blank())
```

Again there is not a discernable pattern in the water flow by hour.  Now I will convert the data to a time series object and plot it:

```{r}
pipe_2_ts <- ts(pipe_2$WaterFlow, start = c(2015, 10, 23, 1), frequency = 24)
ggtsplot(pipe_2_ts, "Pipe #2 Water Flow")
```

The data is stationary so we can begin forecasting.

## Model Creation

### A Note on Process

I will develop a variety of models using different methods and will validate their performance using cross validation. The model that minimizes the error will be the model of choice.

### Pipe 1 Candidates

#### STL + ETS

I will begin with a STL decomposition based model.

```{r}
h <- 24 * 7
pipe_1_stl_ets_fit <- pipe_1_ts %>%
  stlf(h = h, s.window = 24, robust = TRUE, method = "ets")
checkresiduals(pipe_1_stl_ets_fit)
```

This model appears to have preformed fairly well.  The residuals are normal and there's only one minor ACF spike.  I'll see how this preforms in the cross validation stage.

#### STL + ARIMA

Next I will create an ARIMA model on the STL decomposed time series.

```{r}
pipe_1_stl_arima_fit <- pipe_1_ts %>%
  stlf(h = h, s.window = 24, robust = TRUE, method = "arima")
checkresiduals(pipe_1_stl_arima_fit)
```

That's strange.  It is a ARIMA(0,0,0) model.  That is a white-noise model.  Let's see if this is an artifact of the STL decomposition or if we get a similar result training the ARIMA model.

#### ARIMA

Last of all I will try an ARIMA model using the `auto.arima` function to define the (p,d,q) parameters.

```{r}
pipe_1_arima_fit <- auto.arima(pipe_1_ts)
checkresiduals(pipe_1_arima_fit)
```

Again the best model was an ARIMA(0,0,0) model!  This suggests there is no reliable way to forecast the waterflow.

### Pipe 2 Candidates

#### STL + ARIMA

I will first check with a STL decomposition and ARIMA to see if it results in a white noise model.

```{r}
pipe_2_stl_arima_fit <- pipe_2_ts %>%
  stlf(h = h, s.window = 24, robust = TRUE, method = "arima")
checkresiduals(pipe_2_stl_arima_fit)
```

It did.  Let's see what kind of model `auto.arima` comes up with.

#### ARIMA

```{r}
pipe_2_arima_fit <- auto.arima(pipe_2_ts)
checkresiduals(pipe_2_arima_fit)
```

We have another white noise model.  The water flow from pipe 2 cannot be forecasted reliably.

## Summary

White noise models were recommended when modeling the water flow of both pipes using the `auto.arima` function.  This suggests that there is not a reliable way to model the waterflow.