### 2020-04-04 DATA624/HW7.Rmd
Initial commit

---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(gridExtra)
options(scipen=999)
```

## Question 6.2

Developing a model to predict permeability (See Sect. 1.4) could save significant resources for a pharmaceutical company, while at the same time more rapidly identifying molecules that have a sufficient permeability to become a drug:

(A) Start `R` and use these commands to load the data:
```{r}
library(AppliedPredictiveModelings)
data(permeability)
```

## Question 6.3