### 2020-04-29 DATA624/HW10.Rmd
Initial commit

---
title: "DATA 624 Homework 10"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 4
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(dplyr)
library(kableExtra)
library(tidyr)
library(igraph)
library(arules)
library(arulesViz)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set(output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if (is.na(comment)) comment <- ""
  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
                     x, perl = TRUE)
  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
  if (can_null && do_null) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr("\\]", x)[1] - 1
    # Two cases: start or newline
    re <- paste0("^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub(re, rep,  x)
    re <- paste0("\\\n.{", align_index, "}\\]")
    rep <- paste0("\n", comment)
    x <- gsub(re, rep,  x)
  }

  default_output_hook(x, options)

})

knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
```

## Introduction

The assignment is to use R to mine the data for association rules. It should report support, confidence and lift and your top 10 rules by lift.

## Market Basket Analysis

```{r}
transactions <- read.transactions("GroceryDataSet.csv", sep=",")
```

```{r}
itemFrequencyPlot(transactions, topN=10, type="absolute", main="Top 10 Items")
```

In order to complete this market basket analysis, I will use the Apriori algorithm.

```{r}
apriori(transactions, parameter=list(supp=0.001, conf=0.5) , control=list(verbose=FALSE)) %>%
  DATAFRAME() %>%
  arrange(desc(lift)) %>%
  top_n(10) %>%
  kable() %>%
  kable_styling()
```

## Cluster Analysis

If I am given the choice, I will use a network graph to preform the cluster analysis.

```{r}
temp <- read.csv("GroceryDataSet.csv", header = FALSE) %>%
  mutate(shoper_id = row_number()) %>%
  pivot_longer(-shoper_id) %>%
  filter(value != "") %>%
  select(-name)

edges <- temp %>%
  rename(to = value) %>%
  merge(temp) %>%
  rename(from = value) %>%
  filter(to != from) %>%
  select(-shoper_id)

G <- graph_from_data_frame(edges)
```

```{r}
newman_girvan_clusters <- cluster_edge_betweenness(G)
dendPlot(newman_girvan_clusters, mode="hclust")
```

```{r}
plot(newman_girvan_clusters, G) 
```