### 2020-04-29 DATA624/Project 2.Rmd

---
title: "DATA 624 Project 2"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidyr)
library(kableExtra)
library(caret)
library(scales)
library(readxl)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
```

## Introduction

We are to develop a model that predicts the pH level of a beverage.

## Data Exploration

```{r}
df <- read_excel("StudentData.xlsx")
test <- read_excel("StudentEvaluation.xlsx")
```

### Missing Values

```{r}

```