### 2020-05-07 DATA624/Project 2.Rmd
More work on Project 2

---
title: "DATA 624 Project 2"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidyr)
library(kableExtra)
library(corrplot)
library(caret)
library(scales)
library(readxl)
library(naniar)
library(xgboost)
library(Matrix)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
seed <- 42
```

## Introduction

We are to develop a model that predicts the pH level of a beverage.

## Data Exploration

```{r read_data}
df <- read_excel("StudentData.xlsx") %>%
  data.frame()

eval_df <- read_excel("StudentEvaluation.xlsx") %>%
  data.frame()
```

```{r summary_df}
summary(df)
```

### Missing Values

```{r missing_vals_df}
df %>%
  select(-PH) %>%
  vis_miss()
```

```{r df_drop_na_ph}
df <- df %>% 
  drop_na(PH)
```

```{r missing_vals_eval_df}
eval_df %>%
  select(-PH) %>%
  vis_miss()
```

```{r knn_impute}
set.seed(seed)
impute_model <- df %>%
  select(-PH) %>%
  preProcess(., method = "knnImpute")

df_imputed <- predict(impute_model, df) %>%
  na.omit()

eval_df_imputed <- predict(impute_model, eval_df)
```

## Near-Zero Variance

```{r near_zero_variance}
nearZeroVar(df_imputed, saveMetrics = TRUE) %>%
  data.frame() %>%
  rownames_to_column() %>%
  filter(nzv == TRUE)
```

## Correlations

```{r corrplot}
temp <- df_imputed %>%
  select(-Brand.Code) %>%
  na.omit()

p_mat <- cor.mtest(temp, conf.level = 0.95)$p

df_imputed %>%
  select(-Brand.Code) %>%
  na.omit() %>%
  cor() %>%
  corrplot(p.mat = p_mat, sig.level = 0.05, insig = "blank", order = "hclust", type = "upper")
```

## Train/Test Split

```{r train_test_split}
set.seed(seed)
in_train <- createDataPartition(df_imputed$PH, p = 0.8, list = FALSE, times = 1)
train_df_imputed <- df_imputed[in_train,]
test_df_imputed <- df_imputed[-in_train,]
train_df <- df[in_train,]
test_df <- df[-in_train,]
```

## Random Forest

### Ranger

```{r ranger_rf}
set.seed(seed)
tc <- trainControl(method = "cv", number = 10)
rf_fit <- train(PH ~ ., data = train_df_imputed, method = "ranger", importance = "permutation", trControl = tc)
rf_fit

varImp(rf_fit)

rf_predictions <- predict(rf_fit, test_df_imputed)
postResample(pred = rf_predictions, obs = test_df_imputed$PH)
```

### RF

```{r rf, cache=TRUE}
set.seed(seed)
rf_fit2 <- train(PH ~ ., data = train_df_imputed, method = "rf", trControl = tc)
rf_fit2

varImp(rf_fit2)

rf2_predictions <- predict(rf_fit2, test_df_imputed)
postResample(pred = rf2_predictions, obs = test_df_imputed$PH)
```

### Brand Code Random Forests

```{r brand_ranger_rf}
for (brand_code in unique(train_df_imputed$Brand.Code)){
  print(paste("Brand Code", brand_code))
  temp_df <- train_df_imputed %>%
    filter(Brand.Code == brand_code) %>%
    select(-Brand.Code)
  set.seed(seed)
  temp_rf <- train(PH ~ ., data = temp_df, method = "ranger", importance = "permutation", trControl = tc)
  print(temp_rf)
  print(varImp(temp_rf))
  temp_test <- test_df_imputed %>%
    filter(Brand.Code == brand_code) %>%
    select(-Brand.Code) 
  temp_predictions <- predict(temp_rf, temp_test)
  print(postResample(pred = temp_predictions, obs = temp_test$PH))
}
```

## XGBOOST

```{r xgboost}
#converting datasets to matrices
#options(na.action="na.pass")
training2 <- train_df_imputed %>% drop_na(Brand.Code)
testing2 <- test_df_imputed %>% drop_na(Brand.Code)

trainingmx <- model.matrix(~.+0,data=training2[,names(training2) != c("PH")])
testingmx <- model.matrix(~.+0,data=testing2[,names(testing2) != c("PH")])

trainingdmx <- xgb.DMatrix(data = trainingmx, label=training2$PH)
testingdmx <- xgb.DMatrix(data = testingmx, label=testing2$PH)

#default parameters
params <- list(booster = "gbtree", objective = "reg:linear", eta=0.3, gamma=0, max_depth=6, min_child_weight=1, subsample=1, colsample_bytree=1)

#determine the best nround parameter (It controls the maximum number of iterations. For classification, it is similar to the number of trees to grow.)
xgbcv <- xgb.cv( params = params, data = trainingdmx, nrounds = 300, nfold = 5, showsd = T, stratified = T, print_every_n = 10, early_stop_rounds = 20, maximize = F) # best at 250 iterations

xgb_model1 <- xgb.train (params = params, data = trainingdmx, nrounds = 250, watchlist = list(val=testingdmx,train=trainingdmx), print_every_n = 10, early_stop_round = 10, maximize = F)

xgbpred <- predict(xgb_model1,testingdmx)

mat <- xgb.importance (feature_names = colnames(trainingmx),model = xgb_model1)

xgb.plot.importance (importance_matrix = mat)
```