### 2020-05-10 DATA624/Project 2 Group Work.Rmd
More work on Group 2

---
title: "DATA 624 Final Project"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

# R Libraries

```{r, echo = T, results = 'hide'}
library(caret)
library(tidyverse)
library(corrplot)
library(kableExtra)
library(mlbench)
library(dplyr)
library(readxl)
library(VIM)
library(psych)
library(xgboost)
library(Matrix)
library(writexl)
seed <- 12345
```

# Exploratory Data Analysis and PreProcessing

## Data Loading

```{r}
train <- read_excel("StudentData.xlsx")
test <- read_excel("StudentEvaluation.xlsx")
```

## Data Statistics

```{r}
dim(train)
dim(test)
```

#### Complete observations

```{r}
nrow(train[complete.cases(train),])
```

#### All predictors are numeric except for "Brand Code, 1st Variable)

```{r}
summary(train)
summary(test)
```

## Handling missing values (Train and Test sets)

```{r, warning=FALSE}
# Response
train <- train %>% drop_na(PH)

#Predictors
train <- kNN(train, imp_var=FALSE)
```

## Correlation Analysis

```{r}
# Analyze Correlations with the response variable

names <- colnames(train[,-26])
pairs.panels(train[, c("PH", names[1:8])])
pairs.panels(train[, c("PH", names[9:17])])
pairs.panels(train[, c("PH", names[18:26])])
pairs.panels(train[, c("PH", names[27:32])])
```

Top correlated features to PH:
- Mnf Flow (-0.45)
- Bowl Setpoint (0.35)
- Filler Level (0.32)
- Usage Cont (-0.32)
- Pressure Setpoint (-0.31)
- Hyd Pressure3 (-0.24)
- Pressure Vacuum (0.22)
- Hyd Pressure2 (-0.20)

```{r}
train2 <- train %>% dplyr::select(-'Brand Code')

mydata.cor = cor(train2, method = c("spearman"))
corrplot(mydata.cor,cl.cex = 0.7,tl.cex = .7,diag = TRUE)
```

## Correlated Predictors

```{r}
corr <- cor(train[,-c(1,26)], use='complete.obs')
topcorr <- findCorrelation(corr) #top correlated predictors that could be removed to improve modeling
colnames(train[,topcorr])
corrplot(cor(train[,topcorr], use='complete.obs'))
```

## Near Zero Variance Predictors

```{r}
nzv <- nearZeroVar(train)
colnames(train[,nzv])
```

"Hyd Pressure1" should be removed from the dataset as it is constant across observations

## Data Distribution & Variability

```{r}
boxplot(train[, c(names[2:6])])
boxplot(train[, c(names[7:11])])
boxplot(train[, c(names[12:15])])
boxplot(train[, c(names[16:20])])
boxplot(train[, c(names[21:26])])
boxplot(train[, c(names[27:32])])

############ Look at some Decision Trees ##################
source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")
 colnames(train)<- make.names(colnames(train), unique=TRUE)
 colnames(test)<- make.names(colnames(test), unique=TRUE)
train <- as.data.frame(train)
test <- as.data.frame(test)

test <- RF_with_Nulls(train,test,"PH",.5,5,10,.01,5,1)
```

## Data Split

```{r}
set.seed(seed)
train_index <- createDataPartition(train$PH, p = .7, list = FALSE, times = 1)
training <- train[train_index,]
testing <- train[-train_index,]

# Preparing test set for validation
Xtest <- testing[,-grep("PH", colnames(testing))]
```

# Modeling & Evaluation

## Linear Regression

```{r}
set.seed(seed)
lm <- lm(PH~.,data = training)
summary(lm)
```

## Bagged Tree

```{r, warning=FALSE}
set.seed(seed)
bagControl = bagControl(fit = ctreeBag$fit, predict = ctreeBag$pred, aggregate = ctreeBag$aggregate)
bag_model <- train(PH ~., 
                    data = training, method="bag", bagControl = bagControl,
                   center = TRUE,
                   scale = TRUE,
                   trControl = trainControl("cv", number = 5),
                   tuneLength = 25)
bag_model

bag_pred <- predict(bag_model, newdata = Xtest)
postResample(obs = testing$PH, pred=bag_pred)

varImp(bag_model)
```

## XGBoost (Extreme Gradient Boosting)

```{r}
# Converting datasets to matrices
training2 <- training %>% drop_na(`Brand.Code`)
testing2 <- testing %>% drop_na(`Brand.Code`)
trainingmx<-model.matrix(~.+0,data=training2[,names(training2) != c("PH")])
testingmx<-model.matrix(~.+0,data=testing2[,names(testing2) != c("PH")])
trainingdmx <- xgb.DMatrix(data = trainingmx, label=training2$PH) 
testingdmx <- xgb.DMatrix(data = testingmx, label=testing2$PH) 
# Default parameters
params <- list(booster = "gbtree", objective = "reg:linear", eta=0.3, gamma=0, max_depth=6, min_child_weight=1, subsample=1, colsample_bytree=1)
# Determine the best nround parameter (It controls the maximum number of iterations. For classification, it is similar to the number of trees to grow.)
xgbcv <- xgb.cv( params = params, data = trainingdmx, nrounds = 300, nfold = 5, showsd = T, stratified = T, print_every_n = 10, early_stop_rounds = 20, maximize = F) 
# Best at 260 iterations:
```

```{r}
set.seed(seed)
xgb_model1 <- xgb.train (params = params, data = trainingdmx, nrounds = 260, watchlist = list(val=testingdmx,train=trainingdmx), print_every_n = 10, early_stop_round = 10, maximize = F)
# Best result at iteration 211 for training and validation sets:
```

```{r}
mat <- xgb.importance (feature_names = colnames(trainingmx),model = xgb_model1)
xgb.plot.importance (importance_matrix = mat) 
```

## SVM

```{r}
ctrl = trainControl(method='cv', number = 10)
set.seed(seed)
svmRad <- train(PH ~.,
                data=training,
                method = "svmRadial",
                preProc = c("center", "scale"),
                tuneLength = 14,
                trControl = ctrl)
svmRad
svm_pred <- predict(svmRad, newdata = Xtest)
postResample(obs = testing$PH, pred=svm_pred)
varImp(svmRad)
```

## Cubist

```{r}
set.seed(seed)
cubist <- train(PH ~., 
                data = training,
                method='cubist')
cubist  #display model performance
varImp(cubist) #display variable importance
cubist_pred <- predict(cubist, newdata=Xtest) # generate preds
postResample(obs=testing$PH, pred=cubist_pred) # evaluate model over test set
```

## Random Forest

```{r}
ctrl = trainControl(method='cv', number = 10, allowParallel = TRUE)
set.seed(seed)
rforest <- train(PH ~., 
                 data = training,
                 method = "ranger", 
                 importance = "permutation",
                 tuneLength = 10,
                 trControl = ctrl
                 )
rforest #display model performance
varImp(rforest) #display variable importance
rf_pred <- predict(rforest, newdata = Xtest) # generate preds
postResample(obs = testing$PH, pred=rf_pred) # Evaluate model over test set
```

#### Training Random Forest over Individual Brand Codes

```{r brand_ranger_rf}
for (brand_code in unique(training$Brand.Code)){
  print(paste("Brand Code", brand_code))

  temp_df <- training %>%
    filter(Brand.Code == brand_code) %>%
    select(-Brand.Code)
  set.seed(seed)
  temp_rf <- train(PH ~ ., data = temp_df, method = "ranger", importance = "permutation", trControl = ctrl)
  print(temp_rf)
  print(varImp(temp_rf))
  temp_test <- testing %>%
    filter(Brand.Code == brand_code) %>%
    select(-Brand.Code)
  temp_predictions <- predict(temp_rf, temp_test)
  print(postResample(pred = temp_predictions, obs = temp_test$PH))
}
```

# Predicting New Data w/Random Forest

```{r, warning=FALSE}
pfile <- read_excel("StudentEvaluation.xlsx")
#Preparing the dataset we will ultimately predict PH on
test <- pfile[,-grep("PH", colnames(pfile))]
test <- kNN(test, imp_var=FALSE)
colnames(test)<- make.names(colnames(test), unique=TRUE)

ctrl = trainControl(method='cv', number = 10)
set.seed(seed)
rf_model <- train(PH ~., 
                 data = train,
                 method = "ranger", 
                 importance = "permutation",
                 tuneLength = 10,
                 trControl = ctrl
                 )

final_rf_pred <- predict(rf_model, newdata=as.data.frame(test))
pfile$PH <- final_rf_pred # applying predictions to unimputed dataset

write_xlsx(pfile, "Predictions_file.xlsx") # write to excel
```