### 2020-09-15 DATA622/M02-glm.rmd
Initial commit

---
title: "Chapter 02: Our first model,Grand Old Logistic Regression"
output:
html_document: M02-glm.html
df_print: paged
pdf_document: default
---
```{r Startup}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

We agree to maintain all the data in
/home/data/ML.Data directory in 12.42.205.8 and 9. 
If you wish to run these in your local then create a directory. 
I created c:\Users\rkannan\rk\data 
and copied these csv files.
We assume data will be in csv format.

#Load input data heart.csv and ecoli.csv
```{r MainCode}
unixpath<-"/home/data/ML.Data/heart.csv"
unixecolipath<-"/home/data/ML.Data/ecoli.csv"
path<-"c://Users/rkannan/rk/data/heart.csv"
ecolipath<-"c://Users/rkannan/rk/data/ecoli.csv"
heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)
ecoli<-read.csv(ecolipath,head=T,sep=',',stringsAsFactors=F)
```
Let us write a one-line utility function to read in a CSV as above
We assume, comma separated and includes a header record.
```{r}
loadData<-function(csvfile) { read.csv(csvfile,head=T,sep=',',stringsAsFactors=F) }
ecoli_1<-ecoli
ecoli_2<-loadData(ecolipath)
TBL=table(ecoli_1==ecoli_2)
TBL
```
## Review the data, quick and easy EDA
View a few records (head/tail), dimension, number of rows and records
and the variable names.
Everything is an object and mostly a list. We can index into a list to 
access the elements in the list. R index runs from 1 to the length.
```{r}
head(heart)
```
### unix head and tail commands
R commands *head* and *tail* are like the Unix head and tail shell commands
These commands output a few records from the beginning or at the end of the file.
```{r}
tail(heart)
```
### What is the shape of this dataset
How many  observations does it have (each row is an observation)
How many variables are available to us -- number of columns

```{r}
dim(heart)
```
*dim* prints the row X column -- equivalent to *shape in python*
```{r}
nrow(heart)
```
we can get just the number of rows in a data.frame
or just the number of columns
```{r}
ncol(heart)
```
### Manipulating column names 
Recall **heart<-read.csv(path,*header=TRUE*,sep=',',stringsAsFactors=FALSE)**
so R would load the data into a data.frame with column names
Here are some ways to manipulate header
```{r}
names(heart)
names(heart)[3]
length(names(heart))
```
let us change the name of the first column

```{r}
names(heart)[[1]]<-'age'
```
Always trust but verify that our indexing worked...
```{r}
names(heart)
```
# Are the covariates correlated?
Numerically review the pairwise correlation amongst independent variables
```{r}
(corheart<-cor(heart[,1:13]))
```
# Are there any constant feature vectors?
Let us find out if there are any constant variables.
Constant variables cannot possibly contribute to predicting the target.
For example, given this simple birds dataset, and that I saw a bird with two legs,
can you identify which bird that might be?
```{r}
birds<-data.frame(legs=rep(2,8),name=c('crow','pigeon','owl','eagle','sparrow',
'duck', 'woodpecker', 'humming bird'),stringsAsFactors=F)
birds
```
Let us write a utility function we can call on any list...
we can also do length(unique(x)) < 1
```{r}
isConstant<-function(x) length(names(table(x)))<2
apply(heart,2,isConstant)
```
Let us plot,correlations using pairs function
```{r}
library(plotly)
p <- pairs(heart)
```
## What is our class/label variable?
Is the problem at hand *binary classification* or *multi-class classification!*?
Let us assume, our customers want us to predict **target** given the rest.
```{r}
classLabels<-table(heart$target) # this is the actual class distribution both test/train partitions must be close to this.
print(classLabels)
names(classLabels)
```
# Class Imbalance
Let us try to understand the *distribution of the target variable.* Let us make sure
the distribution is balanced. 
### When is a class imbalanced?
If we find 1(minority) in 10 in the extreme case
or anything less than 1 (minority) in 6, we will consider our dataset to be imbalanced.
These are not hard rules but my intuition is as follows
if we are going to create training data with 70/30 split then a random 
sample that contains none or very few of the minority class becomes likely.
Our efforts to train a model is not going to be very fruitful. The model
may not know how to detect minority class or overwhelmed by majority class signature.

```{r}
length(names(classLabels))
ifelse(length(names(classLabels))==2,"binary classification", "multi-class classification")

```
CONSIDER WRITING SMALL ONE LINERS isBinaryClassification(classVector)
and one for, class imbalanced, isImbalancedData(classVector)
IMHO, only one[or two] instance of a certain class -- better to remove that class initially.

## First Classification run
Let us run a simple generalized linear model and try to
understand each variable and how important they are in 
determining the target variable.

we will **NOT** split the data as this is a trial run
and this run is to develop a rough draft script for
any classification exercise.
```{r}
glm_model<-glm(target~.,data=heart, family='binomial')
glm_model
summary_glm_model<-summary(glm_model)
coef_summary_glm_model<-coef(summary_glm_model)
dim(coef_summary_glm_model)
coef_summary_glm_model[[1]]
row.names(coef_summary_glm_model)
```
We will reject any variable with a p-value greater than 0.05.

# p-value
There is much debate regarding p-value. Nothing prevents us from saying
we will reject if p-value is greater than 0.07 after we run the model.

While I respect that viewpoint, I continue to use
p-value as I use 0.05 always. Sometimes, I do experiment with variables
that are indicated as not-significant (p-value > 0.05 for that variable).
For the overall model significance I consistently use 0.05. I do not change that
to accept unreliable model.
I interpret p-value to be the probability of rejecting H0 (the hypothesis) when it is
actually, valid/true.

The p value is the probability of observing a sample value as extreme as, or more extreme than, the value actually observed, given that the null hypothesis is true.
p value represents the risk of rejecting a true null hypothesis. It is the probability of a Type I error if the null hypothesis is rejected.

# Type I and Type II error
When a Type I error (a) is committed, a true null hypothesis is rejected; the innocent person is unjustly convicted. The value is called the level of significance and is the probability of rejecting the true null. With
a Type II error (b), one fails to reject a false null hypothesis; the result is an unjust acquittal, with the guilty person going free. In our system of justice, it is more important to reduce the probability of convicting
the innocent than that of acquitting the guilty. Similarly, hypothesis testing places a greater emphasis
on Type I errors than on Type II errors.

# Ethical Practice
Statistical tools cannot enforce ethical science or human behavior.
So when running  a ML experiment, write ahead of time what your 
acceptance criteria is prior to running the experiment. Include the result
in your reporting regardless of the results.

# File Drawer Problem
Include all results from all the experiments, regardless of the results whether it is in agreement with your expectation or NOT. Eliminate the so called File Drawer problem -- of not including unfavorable results --.
```{r}
coef_summary_glm_model
coef_summary_glm_model[,4]<0.05
row_names<-row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,])
row_names
summary_glm_model$aic
summary_glm_model$null.deviance
summary_glm_model$deviance
ifelse(summary_glm_model$deviance<summary_glm_model$null.deviance,"model has improved","model has not helped")
```
### p-values
Now let us rerun the model with the signifcant variables, that is, with p-value <0.05
```{r}
formulastr<-paste('target~',paste(row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,]),collapse='+'),sep='')
formulastr
paste(row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,]),collapse='+')
model2<-glm(formulastr,data=heart,family='binomial')
summ_model2<-summary(model2)
coef_summ_model2<-coef(summ_model2)
coef_summ_model2
summ_model2$aic
summ_model2$null.deviance
summ_model2$deviance
summary_glm_model$deviance
ifelse(summary_glm_model$deviance<summary_glm_model$null.deviance,"model has improved","model has not helped")
```
So we removed predictors based on p-value and our deviance of the
revised model is higher by about 6%.
So let us compute the VIF factors for the model predictors and 
determine if we removed some variables that should not have been removed.
```{r}
if(!require(car))install.packages('car')
library(car)
vif_model<-vif(glm_model)
```
VIF close to 1 indicates the variables are not correlated
Above 4 is highly correlated.
```{r}
vif_model
vif_model[vif_model>4]
names(vif_model)
nl<-names(vif_model[vif_model<4])
(newformulastr<-paste('target~',paste(nl,collapse='+')))
```
We learned from the entire dataset. Let us get the target variable into a vector.
```{r}
Y<-heart[[14]]
head(Y)
```
Now let us predict using our model. The model has already seen the data.
We do not expect our model, to reproduce at 100%. But we do expect that our model performs better than random guess. If our model under-performs random guess, then that would be an indication of the model's inability to learn at all.
Here in such cases that model is deficient and is a case of under-fitting.
```{r}
newmodel<-glm(formulastr,data=heart,family='binomial')
summary(newmodel)
summnewmodel<-summary(newmodel)
(summnewmodel$aic)
(summnewmodel$deviance)
(summnewmodel$null.deviance)
(p_values<-coef(summnewmodel)[,4])
```
Now let us compute the performance of the classifier.As indicated above,
a random choice would have scored 0.5. If our model is any good it must perform
better than 0.5.
```{r}
table(p_values<0.05)
predYprob<-predict(newmodel,heart[,1:13],type='response')
predY<-ifelse(predYprob<0.5,0,1)
table(predY)
table(heart[[14]])
table(heart[[14]],predY)
library(caret)
require(caret)
cfm<-caret::confusionMatrix(table(heart[[14]],predY))
cfm
```
Accuracy is 84.49%.
But this is from data the model has already seen.

# Underfitting 
A Model with an accuracy of 70% or higher is performant, and therefore, is not underfitting. We can surmise the model is capable of learning. It is better than
random guess which would yield 50% accuracy.

Business is not interested in predicting what has already happened. Business need help
when there is uncertainty and not enough information is available. How will this model perform when a new patient or a new customer arrives. 
*How can we estimate model performance over never seen before data? That is the capability,
**generalize**, we are seeking. Is Life not the response to un-controlled, un-anticipated, unpredictable events?*

# Never seen before data -- HOLD-OUT
We need to devise a process to find out how our model performs when new data is
presented. To do that, we will split our dataset into two disjoint partitions: a training set **70%** and a test set **30%**.
We will hold out some of the data given to us as test set.
We will train the model on the training set. 
Measure its performance on the training set.
Make sure it is at or near 0.85 accuracy, affirming it still is capable of learning with **70%** of the data.
Then we predict the lable for the observations in the test set. 
Measure the performance on the test set.

# Over-fitting, learning from noise
We normally expect performance over test set to be less than training set. If the performance over test set decreases by 20% or more, our model is over-fitting. Two well known causes of over-fitting, (1) *model complexity* and (2) *over-training*. The model has captured both noise and signal and the noise is reflected in the predictions, resulting in sub-optimal **generalization**.

*I have quoted many numerical values as above, but they are my discipline definitely not a rule. They are my choice. You can set your own limits, please do so before running the model and do not change.*

# Data Partition Repeatability/Reproducibility
Let us split our data into "random" test/train disjoint partitions. Always set the seed.
We will train our model on the training set. We will predict using training set to
ensure the model is NOT under-fitting and then predict using the test set, some data
model is not trained on. The test set represents new data. We will evaluate the model's ability to generalize.
We can determine if our model is *over-fitting* influenced by noise in the training set.
```{r}
set.seed(43)
tstidx<-sample(1:nrow(heart),0.30*nrow(heart),replace=F)
trdata<-heart[-tstidx,]
tsdata<-heart[tstidx,]
```
Now let us run the model.
```{r}
glm.trmodel<-glm(formulastr,data=trdata,family='binomial')
summary(glm.trmodel)
predtr<-predict(glm.trmodel,trdata[,1:13],type='response')
head(predtr)

predtrclass<-ifelse(predtr<0.5,0,1)
table(trdata[[14]])
table(predtrclass)

length(predtrclass)==length(trdata[[14]])
(trcfm<-caret::confusionMatrix(table(trdata[[14]],predtrclass)))
```
Accuracy on training data is 0.8404 not bad..model is capable of learning.
Now let us predict the class for never seen before data.
That is our 'held out' test dataset. This is what matters to the business.
```{r}
predts<-predict(glm.trmodel,tsdata[,1:13],type='response')
predtsclass<-ifelse(predts<0.5,0,1)
table(predtsclass)
table(tsdata[[14]])
table(tsdata[[14]],predtsclass)
tscfm<-caret::confusionMatrix(table(tsdata[[14]],predtsclass))
tscfm
(precision <- tscfm$byClass['Pos Pred Value'])    #https://stackoverflow.com/questions/13548092
(recall <- tscfm$byClass['Sensitivity'])
(f_measure <- 2 * ((precision * recall) / (precision + recall))) #geometric mean instead of arithmatic mean
```
Accuracy on the test set (held out data or never seen before data) 
is 0.8111, and has fallen down from accuracy obtained during training phase 0.84. The drop  is about 4% drop in performance and therefore we conclude Model is not over-fitting.
WHAT IF, our performance had fallen quite a bit, say more than 25%, from training to testing phase. That would be a classic case of overfitting. That limit 25% is my discipline. It is not a rule. In consultation with your sponsor you set, these limits and follow it with discipline.

* There are many advanced techniques to mitigate overfitting: 
    +   LASSO, 
    +   RIDGE, and 
    +   Cross Validation.

Here, We will remove some variables and see if it has any bearing on our results.
```{r}
tst.model2<-glm(target~cp+ca+thal+oldpeak,data=trdata,family='binomial')
summary(tst.model2)
tr.tst.pred<-predict(tst.model2,trdata[,1:13],family='binomial')
tr.pred.class<-ifelse(tr.tst.pred<0.5,0,1)
tr.pred.table<-table(trdata[[14]],tr.pred.class)
tr.pred.table
(tr.pred.cfm<-confusionMatrix(tr.pred.table))
(precision <- tr.pred.cfm$byClass['Pos Pred Value'])    #https://stackoverflow.com/questions/13548092
(recall <- tr.pred.cfm$byClass['Sensitivity'])
(f_measure <- 2 * ((precision * recall) / (precision + recall)))
```
Our accuracy for the training phase is `r tr.pred.cfm$overall['Accuracy']`. 
Now let us repeat it over test to find out if there is over-fitting.
```{r}
tst.pred2<-predict(tst.model2,tsdata[,1:13],type='response')
tst.pred2.class<-ifelse(tst.pred2<0.5,0,1)
tst.pred2.table<-table(tsdata[[14]],tst.pred2.class)
tst.pred2.table
(tst.pred2.cfm<-confusionMatrix(tst.pred2.table))
(accuracy<-tst.pred2.cfm$overall['Accuracy'])
(precision <- tst.pred2.cfm$byClass['Pos Pred Value'])    #https://stackoverflow.com/questions/13548092
(recall <- tst.pred2.cfm$byClass['Sensitivity'])
(f_measure <- 2 * ((precision * recall) / (precision + recall)))
```
Accuracy is `r accuracy`. Accuracy on training data was `r tr.pred.cfm$overall['Accuracy']`.
There is no over-fitting. So we move on.
Now let us visualize our performance using ROC plots.
We need ROCR or pROC package and there are other packages.
I use pROC.
```{r echo=FALSE}
graphics.off() 
par("mar") 
par(mar=c(1,1,1,1))
if(!require(pROC))install.packages('pROC')
library(pROC)
par(pty="s") 
glmROC <- roc(tsdata[[14]]~ tst.pred2.class,plot=TRUE,print.auc=TRUE,col="green",lwd =4,legacy.axes=TRUE,main="ROC Curves")
```
Now let us compute AUC and plot Receiver Operating Curve (ROC) using ROCR 
package.
```{r}
getMetrics<-function(actual_class,predicted_response)
{
X=list()
if ( require(ROCR) ) {
auc_1=prediction(predicted_response,actual_class)
prf=performance(auc_1, measure="tpr",x.measure="fpr")
slot_fp=slot(auc_1,"fp")
slot_tp=slot(auc_1,"tp")

fpr=unlist(slot_fp)/unlist(slot(auc_1,"n.neg"))
tpr=unlist(slot_tp)/unlist(slot(auc_1,"n.pos"))

auc<-performance(auc_1,"auc")
AUC<-auc@y.values[[1]]
X=list(fpr=fpr,tpr=tpr,auc=AUC)
}
X
}
```
time to test our utility function...
```{r}
L<-getMetrics(tsdata[[14]],tst.pred2)
plot(L$fpr,L$tpr,main=" ROC Plot tpr vs fpr")
print(paste("AUC=",L$auc,sep=''))
```

We would prefer 90 or above...but AUC=0.89 and accuracy = 0.8 is acceptable.
the recall for this learner is 0.8718 and precision is 0.7727 and their geometric mean, aka F-Score is 0.8193.

## Logistic Regression is a goto algo!!
Definitely, GLM is doing better than a random guess!!! 
Logistic Regression is an excellent baseline algorithm,
My opinion, one must always must perform logistic regression in any ML exercise as a base line.
To show some respect, to grand old **Ockham and his razor**, do NOT use, ANY exotic hyper-parameters, at this stage,
Use just plain vanilla models, as much as possible using objective criteria.
Try to refine programmatically and numerically without requiring human intervention.
That is smart!

*Most importantly there is significant mathematics, statistics and probability behind these algorithms. In the word document, we take a deeper look.*
This module will serve as a template, for all our clinical modules in this course we will provide.

* We will include
   +   R script, 
   +   RMD Script, 
   +   HTML (from the RMD) and 
   +   an essay in word format.

Next Module we will take up Naive Bayes, another parametric model.