### 2020-09-20 DATA622/M03-nb-script.rmd
Initial commit of Module 3

---
title: "NaiveBayes"
output:  
html_document:
df_print: paged
pdf_document: default
---
Let us run a classification task using NaiveBayes.

```{r Startup}

```
##NaiveBayes
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Let us train NaiveBayes model using standard implementation 
and compare it with our own computation. TO make our computation simpler
we will include only categorical variables in the dataset.
```{r MainCode}
library(e1071)
library(caret)
file<-"heart.csv"
```

As practitioners we must always be wary of Overfitting and underfitting.

We consider when the performance of a classifier falls by 50% from 
training to test dataset to be due to overfitting.

We consider when the performance of a classifier is below 50% to be
an indication of a weak classifier unable to learn. Also known as
underfitting.

Let us create some utility functions.

# usage isOverfitting(tr.cfm,tst.cfm)
```{r}
isOverfitting<-function(trcfm,tstcfm)
{
  # if the accuracy from training is much more than accuracy testing phase
  # then we conclude Model is overfitting 
  # unable to generalize
  tracc=trcfm$overall['Accuracy']
  tstacc=tstcfm$overall['Accuracy']
  tracc > 2*tstacc 
}
```
# usage isUnderfitting(tr.cfm)
```{r}
isUnderfitting<-function(trcfm)
{
  # if the accuracy from training is much more than accuracy testing phase
  # then we conclude Model is overfitting 
  # unable to generalize
  tracc=trcfm$overall['Accuracy']
  0.5 >  tracc
}
```
# Load data and EDA
Let us load the data.
And subset it to include only categorical predictors.
```{r}
file<-'c:/Users/rkannan/rk/data/heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)
head(heart)
catheart<-heart[,c(2,3,6,7,9,11,12,13,14)]
```
#Preliminary EDA

```{r}
head(catheart)
dim(catheart)
table(catheart$target)
sum(table(catheart$target))
```
# Repeatable and Reproducible Experiments
Set a seed for repeatable/reproducible random number sequence
and split the data into training/test sets.

```{r}
set.seed(43)
trdidx<-sample(1:nrow(catheart),0.7*nrow(catheart),replace=F)
trcatheart<-catheart[trdidx,]
tstcatheart<-catheart[-trdidx,]
```
# Check for class imbalance
Here we check visually to see if there are labels which occur in the single digit.
If a class exists in just once, if we wish to learn then we cannot test the algorithm for that class and if we include it in the test then we cannot train on it.
There are many statistical techniques to correct for class imbalance, left to another follow up workshop.
```{r}
table(catheart$target)
```
# Training and testing sets must be similar to the entire dataset
Are the training and test set distributed approximately similar to
given dataset. This is important so that statistics that applies 
to training and test sets, are applicable to the entire dataset.
## Otherwise, we cannot generalize.
```{r}
dim(trcatheart)
table(trcatheart$target)
dim(tstcatheart)
table(tstcatheart$target)
```
Train NB model, standard implementation using training set

```{r}
nbtr.model<-naiveBayes(target~.,data=trcatheart)
```
Performance over the training set

```{r}
nbtr.trpred<-predict(nbtr.model,trcatheart[,-c(9)],type='raw')
nbtr.trclass<-unlist(apply(round(nbtr.trpred),1,which.max))-1
nbtr.trtbl<-table(trcatheart[[9]], nbtr.trclass)
tr.cfm<-caret::confusionMatrix(nbtr.trtbl)
tr.cfm
```
Performance over held out data, the test set

```{r}
nbtr.tspred<-predict(nbtr.model,tstcatheart[,-c(9)],type='raw')
#roc.nbtr.tspred<-unlist(apply(nbtr.tspred,1,which.max)) # this returns the position
#roc.nbtr.tspred<-apply(nbtr.tspred,1,FUN=function(v)v[which.max(v)]) # this returns the value
#https://stackoverflow.com/questions/47883541/how-can-i-implement-roc-curve-analysis-for-naive-bayes-classification-algorithm
roc.nbtr.tspred<-nbtr.tspred[,2]
nbtr.tsclass<-unlist(apply(round(nbtr.tspred),1,which.max))-1
nbtr.tstbl<-table(tstcatheart[[9]], nbtr.tsclass)
tst.cfm<-caret::confusionMatrix(nbtr.tstbl)
tst.cfm
```
#Underfitting / Overfitting
```{r}
ifelse(isUnderfitting(tr.cfm),"model is deficient",paste("There is no underfitting", "model is an effective learner@ [",tr.cfm$overall['Accuracy'],"] accuracy"))
ifelse(isOverfitting(tr.cfm,tst.cfm),"model is overfitting -- too complex",
       paste(" There is no overfitting, model is an effective learner@ [",tr.cfm$overall['Accuracy'],"] training accuracy vs testing accuracy=[",tst.cfm$overall['Accuracy'],']'))
```
#Conditionally Independent Class Probabilities
Let us now compute NaiveBayes probabilities for each class
directly from data.
Given any predictor and a value
we compute the class conditional probabilities
p(c1|x) and p(c2|x) using 
p(c1|X)=p(x1|c1)p(x2|c1)p(x3|c1)..p(xn|c1)p(c1)
and
p(c2|X)=p(x1|c2)p(x2|c2)p(x3|c2)..p(xn|c2)p(c2)

Here p(c1)=frequency of c1 / total frequency
p(x1|c1)=frequency of x1 given c1/frequency of c1

we assign the class with the higher probability

```{r}
trcathearttarget_0<-trcatheart[trcatheart$target==0,]

trcathearttarget_1<-trcatheart[trcatheart$target==1,]

tstcathearttarget_0<-tstcatheart[tstcatheart$heart==0,]
tstcathearttarget_1<-tstcatheart[tstcatheart$heart==1,]
```
We compute class frequencies as follows, 
the number of rows in each class is the frequency of that class.
# class c1 proportion
```{r}
c0freq<-nrow(trcathearttarget_0)
c1freq<-nrow(trcathearttarget_1)
p0<-c0freq/(c0freq+c1freq)
p1<-c1freq/(c0freq+c1freq)
c(p0,p1)
```
We count, the number of observations in which the attribute has the given value
#We return the two proportions for a binary classifier.
```{r}

classProbabilities<-
  function(dsetc1,dsetc2,prop,val,c1prob,c2prob)
  {

    propdsetc1=dsetc1[dsetc1[prop]==val,]
    propdsetc2=dsetc2[dsetc2[prop]==val,]
    c(nrow(propdsetc1)/nrow(dsetc1),nrow(propdsetc2)/nrow(dsetc1))
  }
```
Let us put our utility functions to test, calculate conditional probability
```{r}
trcathearttarget_0[1,]
```
sex cp fbs restecg exang slope ca thal target
276   1  0   0       1     0     2  2    3      0

Let us now compute the probabilities for each of these attribute/value pairs.

```{r}
classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',1,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',0,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',0,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',1,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',2,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',3,p0,p1)
```

```{r}
xx<-c(
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',3,p0,p1),
  c(p0,p1))
nr<-9
c(p0,p1)
mx<-matrix(xx,nr,ncol=2,byrow=T)
#mx
cp<-unlist(apply(mx,2,cumprod))
#cp[nr,]
pred_class<-which.max(cp[nr,])-1
trcathearttarget_0[1,]$target
ifelse(trcathearttarget_0[1,]$target==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))
```

```{r}
tstcatheart[7,]
```
sex cp fbs restecg exang slope ca thal target
26   0  1   0       1     0     2  2    2      1
```{r}
tstcathearttarget_0<-tstcatheart[tstcatheart$heart==0,]
tstcathearttarget_1<-tstcatheart[tstcatheart$heart==1,]

xx<-c(
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',2,p0,p1),
  c(p0,p1))

mx<-matrix(xx,nrow=9,ncol=2,byrow=T)

unlist(apply(mx,2,cumprod))
cp<-unlist(apply(mx,2,cumprod))
which.max(cp[9,])-1
pred_class<-which.max(cp[nr,])-1
pred_class
tstcatheart[7,]$target
ifelse(tstcatheart[7,]$target==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))
```
#NOTE THAT we are using the distribution of training set 
#to evaluate the test set probabilities
# otherwise we could not claim to be generalizing
Let us do one more 23rd observation from the test set.seed
```{r}
tstcatheart[23,]

xx<-c(
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',2,p0,p1),
  c(p0,p1))

mx<-matrix(xx,nrow=9,ncol=2,byrow=T)

unlist(apply(mx,2,cumprod))
cp<-unlist(apply(mx,2,cumprod))
which.max(cp[9,])-1
pred_class<-which.max(cp[nr,])-1
pred_class
tstcatheart[23,]$target
ifelse(tstcatheart[23,]$target==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))

```
```{r}
evaluate_naiveBayes<-
function(nv,vv,targetclass,trdsetcl1,trdsetcl2,cl1prob,cl2prob)
{
nvl<-length(nv)
xx<-c()
for (idx in 1:nvl)
{
attr<-nv[idx]
value<-vv[[idx]]
#print(c(attr,value))
#xx<-c(xx,classProbabilities(trdsetcl1,trdsetcl2,nv[idx],vv[[idx]],cl1prob,cl2prob))
xx<-c(xx,classProbabilities(trdsetcl1,trdsetcl2,attr,value,cl1prob,cl2prob))
}
xx<-unlist(c(xx,  c(p0,p1)))
nvl<-nvl+1
mx<-matrix(xx,nrow=nvl,ncol=2,byrow=T)

unlist(apply(mx,2,cumprod))
cp<-unlist(apply(mx,2,cumprod))

pred_class<-which.max(cp[nvl,])-1
pred_class
targetclass
metric<-ifelse(targetclass==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))
c(targetclass,pred_class,metric)
}

```
Now we can iterate over the entire test set.
```{r}
tstcount<-nrow(tstcatheart)
tstcol<-ncol(tstcatheart)
featuremx<-tstcatheart[,-c(which(names(tstcatheart)=='target'))]
nv<-names(featuremx)
tstclass<-tstcatheart$target
tstpred<-c()
for(r in 1:tstcount)
{
  tstpred<-c(tstpred,
             evaluate_naiveBayes(nv,featuremx[r,],
                            tstclass[r],trcathearttarget_0,
                            trcathearttarget_1,p0,p1))
}
tstresults<-matrix(tstpred,nrow=tstcount,ncol=3,byrow=T,dimnames=
                     list(1:tstcount,c("GT",'Pred','metric')))
tstresults
table(tstresults[,3])
table(tstresults[,1],tstresults[,2])
```
# Final observation 
Note that we estimated the class probabilities using the training set
however, we compute the metric using the test set.
# Is that not what we want?
Statistics and probabilities is one big hat from which we can pull many rabbits,
and if you are NOT cautious you can end up with a 3 legged rabbit.
This classifier assumes that features are uncorrelated and yet we achieved significant
inferences.

If you recall, the e1071:naiveBayes achieved 
```{r}
nbtr.tstbl
```
in comparison to our naive implementation.
```{r}
table(tstresults[,1],tstresults[,2])
```
Let us plot the ROCR curve and determine AUC for the e1071 standard
implementation.
```{r}
getMetrics<-function(actual_class,predicted_response)
{
  test.set=data.frame(target=actual_class,prediction=predicted_response)
  TPrates = c()
TNrates = c()
thresholds = seq(0, 1, by = 0.05)
total_positive=nrow(test.set[test.set$target==0 ,])
total_negative=nrow(test.set[test.set$target==1,])
for (threshold in thresholds) {
  TPrateForThisThreshold = nrow(test.set[test.set$target == 0 & test.set$prediction <= threshold,])/total_positive
  TNrateForThisThreshold = nrow(test.set[test.set$target == 1 & test.set$prediction > threshold,])/total_negative
  TPrates = c(TPrates, TPrateForThisThreshold)
  TNrates = c(TNrates, TNrateForThisThreshold)
}
if(!require(pracma))install.packages('pracma')
library(pracma)
AUC <- trapz(TPrates,TNrates)
X=list(fpr=1-TNrates,tpr=TPrates,auc=AUC)
X
#X=list()
#if ( require(ROCR) ) {
#auc_1=prediction(predicted_response,actual_class)
#prf=performance(auc_1, measure="tpr",x.measure="fpr")
#slot_fp=slot(auc_1,"fp")
#slot_tp=slot(auc_1,"tp")

#fpr=unlist(slot_fp)/unlist(slot(auc_1,"n.neg"))
#tpr=unlist(slot_tp)/unlist(slot(auc_1,"n.pos"))

#auc<-performance(auc_1,"auc")
#AUC<-auc@y.values[[1]]
#X=list(fpr=fpr,tpr=tpr,auc=AUC)
#}
#X
}
```
time to test our utility function...
```{r}
L<-getMetrics(tstcatheart[[9]],roc.nbtr.tspred)
plot(L$fpr, L$tpr, type="l",main=" NB ROC Plot tpr vs fpr")
#plot(L$fpr,L$tpr,main=" NB ROC Plot tpr vs fpr")
print(paste(" NB AUC=",L$auc,sep=''))
```