### 2020-09-30 DATA622/M04-lda-heart.RMD
Initial commit of M04 files

---
title: "lda-heart"
author: "Raman Kannan"
date: "8/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Discriminant Analysis

LDA attempts to identify a dimension in which the separation between classes is maximized. It can be implemented using linear algebra or the ISLR formulation. In the accompanying chapter we illustrated using iris to show that LDA can handle multi-class problems. Here, we will customize it for heart dataset, a binary classification problem.

```{r }
heart<-read.csv('heart.csv',sep=',',head=T,
                stringsAsFactors=F)
unlist(lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

nvpairsdf3<-do.call('rbind',lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

categoricalFeatures<-function(dset,ncol=7)
{
  df4<-as.data.frame(do.call('rbind',lapply(names(dset),FUN=function(x,data=dset){c(cname=x,uvalfreq=length(unique(data[[x]])))})))
  df4$uvalfreq=as.numeric(df4$uvalfreq)
  dset[,df4$uvalfreq<=ncol]
}

X<-categoricalFeatures(heart)

table(heart$target)
dim(X)

heart0<-X[X$target==0,]
heart1<-X[X$target==1,]
names(heart0)
```

## Let us compute class means for each feature

data.frames and vector manipulation -- made for R --

```{r }
heart0mean<-apply(heart0[,1:(ncol(X)-1)],2,mean)
heart1mean<-apply(heart1[,1:(ncol(X)-1)],2,mean)
```

Let us compute the mean for the entire dataset
```{r}
allclassmean<-apply(X[,1:(ncol(X)-1)],2,mean)
```
And the covariances
```{r}
S1<-cov(heart0[,1:(ncol(X)-1)])
S2<-cov(heart1[,1:(ncol(X)-1)])
```
Class frequencies
```{r}
N1<-nrow(heart0)
N2<-nrow(heart1)
```
Let us compute pooled covariance
```{r}
SWpooled<-((N1-1)*S1+(N2-1)*S2)/(N1+N2-2)
```
For LDA we need scatter between (SB) for each class.
```{r}
SB1<-N1*(heart0mean-allclassmean)*(heart0mean-allclassmean)
SB2<-N2*(heart1mean-allclassmean)*(heart1mean-allclassmean)
SB<-SB1+SB2
```
Let us predict using linear algebra
```{r}
invSWpooled<-solve(SWpooled)
invSWSBpooled<-invSWpooled*SB
SWSBpooledeigendecomp<-eigen(invSWSBpooled)
lambdas<-SWSBpooledeigendecomp$vectors[,which.max(SWSBpooledeigendecomp$values)]
as.matrix(X[,1:(ncol(X)-1)])%*% lambdas->heart.projections.pooled

```

Convert probabilities into class projections...

```{r}
rounded.heart.projections.pooled<-round(heart.projections.pooled)
heart.lda.pooled<-cbind(X,classp=heart.projections.pooled,spectral=rounded.heart.projections.pooled)
heart.lda.pooled$predicted <-ifelse(heart.lda.pooled$spectral < -1,1,0)
heart.lda.confMat.pooled<-table(  heart.lda.pooled$target,heart.lda.pooled$predicted)
heart.lda.confMat.pooled
```
Not exactly impressive...
Let us calculate the metrics 
```{r}
print(paste("Accuracy is::",
            as.numeric(sum(diag(heart.lda.confMat.pooled))/sum(heart.lda.confMat.pooled)),sep= " "))
print(paste("Error is::",
            1-as.numeric(sum(diag(heart.lda.confMat.pooled))/sum(heart.lda.confMat.pooled)),
            sep= " "))
plot(heart.lda.pooled$classp,col=as.factor(heart.lda.pooled$predicted),pch=16)
```
Let us now reimplement ISLR chapter 4 - implemented by Professor Dalal (Columbia.edu)
for the heart dataset.
```{r}
default_pos<-X[X[,9]==0,]
default_neg<-X[X[,9]==1,]
dim(default_pos)
mu_pos<-apply(default_pos[,1:8],2,mean)
mu_neg<-apply(default_neg[,1:8],2,mean)
sigma_pos<-cov(default_pos[,1:8])
sigma_neg<-cov(default_neg[,1:8])
n_pos<-dim(default_pos)[1]
n_neg<-dim(default_neg)[1]
sigma_all<-((n_pos-1)*sigma_pos+(n_neg-1)*sigma_neg)/(n_pos+n_neg-2) #pooled Cov matrix-
mu<-cbind(mu_pos,mu_neg)
pi.vec <- rep(0,2) # why not just pi?
pi.vec[1] <- sum((heart$target==0))/nrow(heart)
pi.vec[2] <- sum((heart$target==1))/nrow(heart)

my.lda <- function(pi.vec,mu,Sigma,x){
x.dims <- dim(x)
n <- x.dims[1]
Sigma.inv <- solve(Sigma) #Find inverse of Sigma
out.prod <- rep(0,n) #all items initiated to be negative
# equation 4.19 from ISLR Sixth Printing Page 157
discrim.pos <- apply(x,1,function(y) y %*% Sigma.inv %*% mu[,1]
- 0.5*t(mu[,1]) %*% Sigma.inv %*% mu[,1] + log(pi.vec[1])) # target == 0
discrim.neg <- apply(x,1,function(y) y %*% Sigma.inv %*% mu[,2]
- 0.5*t(mu[,2]) %*% Sigma.inv %*% mu[,2] + log(pi.vec[2]))# target == 1 
out.prod[discrim.neg >= discrim.pos] <- 1
out.prod.df<-data.frame(posp=discrim.pos,negp=discrim.neg,prod=out.prod)
return(out.prod.df)
}
```
Let us run LDA on heart dataset using Sid's implementation.

```{r}
heart_default_all<-my.lda(pi.vec,mu,sigma_all,X[,1:8])
hda<- heart_default_all
hda$posprob<-hda$posp/(hda$posp+hda$negp)
hda$negprob<-hda$negp/(hda$posp+hda$negp)
#produce confusion matrices
table(X$target,heart_default_all$prod)

cfmx<-caret::confusionMatrix(table(X$target,heart_default_all$prod))
```
Now let us run LDA using standard implementation in the MASS package
```{r}
require(MASS)
lda_model<-lda(target~.,data=X,type='response')
lda_pred<-predict(lda_model,X[,1:8])
table(X$target,lda_pred$class)
cfmx_mass<-caret::confusionMatrix(table(X$target,lda_pred$class))

dfcomp<-data.frame(sid=cfmx$byClass,mass=cfmx_mass$byClass)
dfcomp
```
```{r}
require(ROCR)
pred <- prediction(lda_pred$posterior[,2], X$target)
AUC<-performance(pred,"auc")
perf <- performance(pred,"tpr","fpr")
plot(perf,colorize=TRUE)
text(0.8,0.2,paste("AUC=",round(AUC@y.values[[1]],4),sep=''))

```