### 2020-09-30 DATA622/M05-kNN.rmd
Initial commit

---
title: "kNN Implementation"
author: "Raman Kannan"
date: "8/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

We will continue with Instance based methods. These algorithms do NOT compute
probability distribution.

kNN works as follows

compute the distance between the given instance and all the known instance.
Sort by distance and assign the most frequent class of the k nearest neighbors.

```{r}
euclideanDist <- function(a, b){
  d = 0
  for(i in c(1:(length(a)) ))
  {
    d = d + (a[[i]]-b[[i]])^2
  }
  d = sqrt(d)
  return(d)
}
```
Let us write another function to iterate over the test set 
and over the training set for each observation in the test set.
```{r}
knn_predict2 <- function(test_data, train_data, k_value, labelcol){
  print(k_value)
  print(labelcol)
  pred <- c()  #empty pred vector 
  #LOOP-1
  for(i in c(1:nrow(test_data))){   #looping over each record of test data
    eu_dist =c()          #eu_dist & eu_char empty  vector
    eu_char = c()
    good = 0              #good & bad variable initialization with 0 value
    bad = 0

    #LOOP-2-looping over train data 
    for(j in c(1:nrow(train_data))){

      #adding euclidean distance b/w test data point and train data to eu_dist vector
      eu_dist <- c(eu_dist, euclideanDist(test_data[i,-c(labelcol)], train_data[j,-c(labelcol)]))

      #adding class variable of training data in eu_char
      eu_char <- c(eu_char, as.character(train_data[j,][[labelcol]]))
      #print(i,j,as.character(train_data[j,][[labelcol]])))
    }

    eu <- data.frame(eu_char, eu_dist) #eu dataframe created with eu_char & eu_dist columns

    eu <- eu[order(eu$eu_dist),]       #sorting eu dataframe to gettop K neighbors
   # print(k_value)
    eu <- eu[1:k_value,]               #eu dataframe with top K neighbors

    head(eu[1:k_value,])
    tbl.sm.df<-table(eu$eu_char)
    cl_label<-  names(tbl.sm.df)[[as.integer(which.max(tbl.sm.df))]]

    pred <- c(pred, cl_label)
  }
  return(pred) #return pred vector
}

```
Let us also write a utility function for accuracy.
```{r}
accuracy <- function(test_data,labelcol,predcol){
  correct = 0
  for(i in c(1:nrow(test_data))){
    if(test_data[i,labelcol] == test_data[i,predcol]){ 
      correct = correct+1
    }
  }
  accu = (correct/nrow(test_data)) * 100  
  return(accu)
}
```
Now let us load and prepare training/test set from the heart dataset.

We will call the predict function we wrote above.
```{r}
#load data
file<-'heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)
head(heart)
catheart<-heart[,c(2,3,6,7,9,11,12,13,14)]

knn.df<-catheart
labelcol <- 9 # for iris it is the fifth col 
predictioncol<-labelcol+1
# create train/test partitions
set.seed(43)
n<-nrow(knn.df)
knn.df<- knn.df[sample(n),] # randomize

train.df <- knn.df[1:as.integer(0.7*n),]
test.df <- knn.df[as.integer(0.7*n +1):n,]

K<-3 # number of neighbors to determine the class

predictions <- knn_predict2(test.df, train.df, K,labelcol) #calling knn_predict()

```
Computing ROCR and AUC is some what non-trivial as kNN do not compute probabilities
and results in unreliable ROCR plots.
```{r}
test.df[,predictioncol] <- predictions #
print(paste('accuracy=',accuracy(test.df,labelcol,predictioncol)))
```
Now let us label the test set using standard kNN implementation from class
package.
```{r}
library(class)
library(ROCR)

train.df <- knn.df[1:as.integer(0.7*n),]
test.df <- knn.df[as.integer(0.7*n +1):n,]

cl<-factor(train.df[,9])
knnPred<-knn(train.df[,-c(labelcol)],test.df[,-c(labelcol)], cl, k = 3, prob=TRUE)
prob<-attr(knnPred,"prob")
kp<-prediction(prob,test.df[,labelcol])
AUC<-performance(kp,"auc")
#perf <- performance(pred,"tpr","fpr")
#plot(perf,colorize=TRUE)
pred_knn<-performance(kp,"tpr","fpr")
```
```{r echo=FALSE}
plot(pred_knn,avg="threshold",colorize=T,lwd=3,main="ROCR knn")
text(0.8,0.2,paste("AUC=",round(AUC@y.values[[1]],4),sep=''))
```
```{r echo=FALSE}
TBL<-table(test.df[,labelcol],knnPred)
print(paste("accuracy=",sum(diag(TBL))/sum(TBL)))
```