### 2020-10-15 DATA622/M06-tree.RMD
Initial commit

---
title: "M06-DTree.RMD"
author: "Raman Kannan"
date: "8/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
heart<-read.csv('heart.csv',sep=',',head=T,
                stringsAsFactors=F)
unlist(lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

nvpairsdf3<-do.call('rbind',lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

categoricalFeatures<-function(dset,ncol=7)
{
  df4<-as.data.frame(do.call('rbind',lapply(names(dset),FUN=function(x,data=dset){c(cname=x,uvalfreq=length(unique(data[[x]])))})))
  df4$uvalfreq=as.numeric(df4$uvalfreq)
  dset[,df4$uvalfreq<=ncol]
}

X<-categoricalFeatures(heart)

table(heart$target)
dim(X)

```

```{r, echo=FALSE}
library(rpart)
library(rpart.plot)
library(ROCR)

isModelReliable<-function(cfmx)
{	cfmx$overall[['AccuracyPValue']]<0.05;
 }

isUnderfitting<-function(cfmx)
{	cfmx$overall[['Accuracy']]<0.60;
 }

isOverfitting<-function(tr,tst)
{	tst$overall[['Accuracy']]/tr$overall[['Accuracy']]<0.75;
 }

 getSensitivity<-function(cfmx) cfmx$byClass[['Sensitivity']]

getAccuracy<-function(cfmx) cfmx$overall[['Accuracy']]

getSpecificity<-function(cfmx) cfmx$byClass[['Specificity']]

getMetrics<-function(actual_class,predicted_response)
{
X=list()
if ( require(ROCR) ) {
auc_1=prediction(predicted_response,actual_class)
prf=performance(auc_1, measure="tpr",x.measure="fpr")
slot_fp=slot(auc_1,"fp")
slot_tp=slot(auc_1,"tp")

fpr=unlist(slot_fp)/unlist(slot(auc_1,"n.neg"))
tpr=unlist(slot_tp)/unlist(slot(auc_1,"n.pos"))

auc<-performance(auc_1,"auc")
AUC<-auc@y.values[[1]]
X=list(prf=prf,fpr=fpr,tpr=tpr,auc=AUC)
}
X
}
X$target=as.factor(X$target)

rpart.model<-rpart(target~.,data=X,minsplit=3)
rpart.plot(rpart.model)

```

That is a nice chart but several questions come to mind....
What are the numbers in each node?
What do they represent?
There are three numbers in the root node 1,0.54,100%

Here 1 is the class.
54% of the observations belong to class 1 and we can confirm this by 
```{r} table(X$target)/nrow(X)```
The third number 100% says, all observations are accounted for in this node, obviously so for the
root node.

the test at this node is cp < 1.

Let us take the cp<1 branch to the left and here we find 0,0.27,47%.
This node is class 0 node.And, 27% of the observations will end up being class 1.
The node id is always the majority class. And these are easily confirmed 
`r table(X[X$cp<1,]$target)/sum(table(X[X$cp<1,]$target)) ` and this node represents
`r nrow(X[X$cp<1,])/nrow(X)` of the dataset.
```{r}
table(X[X$cp<1,]$target)/sum(table(X[X$cp<1,]$target)) #class probabilities 
nrow(X[X$cp<1,])/nrow(X) # proportion of the dataset
#or we can follow the next node
table(X[X$cp<1&X$ca>=1,]$target)/sum(table(X[X$cp<1&X$ca>=1,]$target))
# and confirm the majority class for this node is 0 and proportion of class 1 to be

```

Let us go to the next node the test here is ca >= 0
If ca < 1 we should get `r table(X[X$cp<1&X$ca<1,]$target)/sum(table(X[X$cp<1&X$ca<1,]$target)) `
The majority class here is 1. And the proportion of the dataset in this node is
`r nrow(X[X$cp<1&X$ca<1,])/nrow(X) `
```{r}
table(X[X$cp<1&X$ca<1,]$target)/sum(table(X[X$cp<1&X$ca<1,]$target))
nrow(X[X$cp<1&X$ca<1,])/nrow(X)

# if we follow ca>0
table(X[X$cp>0,]$target)/nrow(X[X$cp>0,]) # we get 79% for class 1
nrow(X[X$cp>0,])/nrow(X) # for 52% of the dataset
```
Confirm the numbers are in agreement for 3 more nodes..
Node ID is always the majority class.
The second number is the probability for class 1.
and the third number is % of observations out of the whole dataset represented by the node.

##Why did the tree start with the feature 'cp'?
##How did it find that is the right strategy?

Let us reduce the number of features and run a small model to understand 
these things. With this large a dataset we may never get to the point.

```{r}
library(sqldf)

s1<-sqldf('select count(*) from X where cp < 1')
head(s1)

nrow(X[X$cp<1,])/nrow(X)

table(X[X$cp<1,]$target)

table(X[X$cp<1,]$target)/sum(table(X[X$cp<1,]$target))
```
There are 143 with cp<1 out of ```{r} nrow(X) ``` let us start with a smaller 3 feature dataset ...
```{r}
smh<-X[,c(2:4,9)]
smh$target<-as.factor(smh$target)
smh_rpart<-rpart(target~.,data=smh,minsplit=2)
rpart.plot(smh_rpart,main=' 3 feature dataset to illustrate gini_index')

```

```{r}

smhx<-sqldf('select * from smh where cp <1 and fbs =1')

smhx

gini<-function(dset,cname,label)
{
cnames<-unique(dset[[label]])
vprop<- as.matrix(as.numeric(table(dset[[cname]])/sum(table(dset[[cname]]))))
 T1<-table(dset[[cname]],as.vector(dset[[label]]))
 T1mx<-as.matrix(T1)
 T11<-as.matrix(t(apply(T1mx,1,FUN=function(v){v/sum(v)})))
 pT1<-as.data.frame(cbind(T1mx,vprop,T11))
nclass<-length(unique(dset[[label]]))
pT1x<-pT1[,(nclass+1+1):(1+2*nclass)]
 pT2<-cbind(pT1,t(apply(pT1[,(nclass+1+1):(1+2*nclass)],1,FUN=function(v) v*v)))
pT3<-cbind(pT2,apply(pT2,1,FUN=function(v)(v[nclass+1]*(1-sum(v[(1+2*nclass+1):(1+3*nclass)])))))
pT3.df<-as.data.frame(pT3)
colnames<-c(paste('f_',cnames,sep=''),'valprop',paste('p_',cnames,sep=''),
paste('sqp_',cnames,sep=''),'gini')
names(pT3.df)<-colnames
pT3.df$feature=cname
pT3.df$label=label
pT3.df$feature_value=row.names(T1)
pT3.df
 }

compute_gini<-function(dset,labelCol)

{

ginimx<-do.call('rbind',lapply(names(dset)[names(dset)!=labelCol],FUN=function(x,df=dset,label=labelCol)gini(df,x,label)))

ginimx
}

ginimx<-compute_gini(smh,'target')

ginimx

ginif<-sqldf('select feature,sum(gini) as tgini from ginimx group by feature')

ginif[which.min(ginif$tgini),1]

```
Does it work recursively?

let us check on the full dataset for the partition when cp>0

Recall the tree we rendered initially.

```{r}
#catheart<-categoricalFeatures(heart)
bigkahuna<-compute_gini(X[X$cp>0,],'target')
ginit<- sqldf('select feature, sum(gini) tgini from bigkahuna group by feature')
ginit[which.min(ginit$tgini),c('feature')]
```

So we now have a systematic method to partition a given dataset
by the feature that most isolate the class using GINI index.