### 2020-10-16 DATA622/M07-svm.rmd
Initial commit

---
title: "M07-SVM"
author: "Raman Kannan"
date: "9/28/2020"
output: 
html_document: M07-svm.html
df_print: paged
pdf_document: default
---

```{r Startup}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

#Load input data heart.csv and ecoli.csv
```{r MainCode}
unixpath<-"/home/data/ML.Data/heart.csv"
unixecolipath<-"/home/data/ML.Data/ecoli.csv"
path<-"c://Users/rk215/Data/heart.csv"
ecolipath<-"c://Users/rk215/Data/ecoli.csv"
heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)
ecoli<-read.csv(ecolipath,head=T,sep=',',stringsAsFactors=F)

```

## Review the data, quick and easy EDA
View a few records (head/tail), dimension, number of rows and records
and the variable names.
Everything is an object and mostly a list. We can index into a list to 
access the elements in the list. R index runs from 1 to the length.
```{r}
head(heart)
```
### unix head and tail commands
R commands *head* and *tail* are like the Unix head and tail shell commands
These commands output a few records from the beginning or at the end of the file.
```{r}
tail(heart)
```

let us change the name of the first column

```{r}
names(heart)[[1]]<-'age'
```
Always trust but verify that our indexing worked...
```{r}
names(heart)
```
We have already performed EDA when we ran glm. 
```{r}
library(plotly)
library(e1071)
library(caret)
library(pROC)

```

## Let us run
Let us run SVM from e1071 package and try to
understand each variable and how important they are in 
determining the target variable.

we will **NOT** split the data as this is a trial run
and this run is to develop a rough draft script for
any classification exercise.
```{r}
heart$target<-factor(heart$target)
svm_model<-svm(target~.,data=heart, probability=TRUE)
svm_model

```

```{r}
svm_predict<-predict(svm_model,heart[,-c(14)],probability=TRUE)
probs<-attr(svm_predict,"prob")[,"1"]
predclass<-ifelse(probs>0.5,1,0)
table(heart$target==predclass)
(cfmx<-caret::confusionMatrix(table(heart$target,predclass)))
```
### Our model can learn ....
As evidenced by the F1 ```{cfmx$byClass[['F1']]}``` and the Balanced Accuracy ```{cfmx$byClass[['Balanced Accuracy']]}``` Now we should compute testing error -so called- generalization error...
To determine the generalization error, we split the data into training and testing dataset.
We dont tune as a matter of principle so we do not need the validation set.
Then we train on the training set and predict the target for the testing set.
That will provide us a window into the generalization capability for the learner.
This capability to generalize is all that matters for the business. So let us do it.
```{r}
set.seed(43)
tstidx<-sample(1:nrow(heart),0.30*nrow(heart),replace=F)
trdata<-heart[-tstidx,]
tsdata<-heart[tstidx,]

#svm(target~.,data=heart, probability=TRUE)

svm_train_model<-svm(target~.,data=trdata,probability=TRUE)
svm_train_predict<-predict(svm_train_model,trdata[,-c(14)],probability=TRUE)
train_probs<-attr(svm_train_predict,"prob")[,"1"]
train_predclass<-ifelse(train_probs>0.5,1,0)
tr_tbl<-table(trdata$target,train_predclass)
(tr_cfmx<-caret::confusionMatrix(tr_tbl))

cdf<-data.frame(all=cfmx$byClass,tr=tr_cfmx$byClass)
cdf
```
The performance measures are very similar therefore we can conclude the data partition has no impact on 
the distribution or the model performance and the model is stable. Now we can evaluate the performance on the test set.
```{r}
svm_tst_predict<-predict(svm_train_model,tsdata[,-c(14)],probability=TRUE)
tst_probs<-attr(svm_tst_predict,"prob")[,"1"]
tst_predclass<-ifelse(tst_probs>0.5,1,0)
tst_tbl<-table(tsdata$target,tst_predclass)
(tst_cfmx<-caret::confusionMatrix(tst_tbl))

cdf<-cbind(cdf,ts=tst_cfmx$byClass)
cdf
```
# Let us now run the SVM on the attributed we identified with VIF
```{r}
Y<-heart[[14]]
vif_col<-c(2,3,6,7,9,11,12,13,14)
vif_heart<-heart[,vif_col]
vif_trdata<-trdata[,vif_col]
vif_tsdata<-tsdata[,vif_col]
head(vif_heart)
head(Y)

svm_vif_tr_model<-svm(target~.,data=vif_trdata,probability=TRUE)
svm_vif_tr_predict<-predict(svm_vif_tr_model,vif_trdata[,-c(14)],probability=TRUE)
vif_train_probs<-attr(svm_vif_tr_predict,"prob")[,"1"]
vif_train_predclass<-ifelse(vif_train_probs>0.5,1,0)
vif_tr_tbl<-table(vif_trdata$target,vif_train_predclass)
(vif_tr_cfmx<-caret::confusionMatrix(vif_tr_tbl))
(cdf<-cbind(cdf,vif_tr=vif_tr_cfmx$byClass)) # this represents our model's ability to learn -- training error
# is interpreted as underfitting when < 0.75 I use 0.85 
# remedy is to consider more complex model 
```
**Why is the performance with VIF predictors slightly better?**  
Next we will predict the target for our test data.
And the performance (or error) calculated on the test data represents generalization error for this model.
```{r}
svm_vif_tst_predict<-predict(svm_vif_tr_model,vif_tsdata[,-c(9)], probability=TRUE)
svm_vif_tst_predict_probs=attr(svm_vif_tst_predict, "prob")[,"1"]
svm_roc_obj<-roc(vif_tsdata[,c(9)],svm_vif_tst_predict_probs)
svm_auc<-auc(svm_roc_obj)

```
Now let us plot ROC Curve
```{r}
plot.new()
plot.window(xlim=c(1,0), ylim=c(0,1), xaxs='i', yaxs='i')
axis(1, las=1)
axis(2, las=1)
abline(1,-1, col="black", lwd=0.5)
box()
title(main="ROC", xlab="Specificity", ylab="Sensitivity")
lines(svm_roc_obj, col="black", las=2)
text(x=0.2,y=0.2,paste("SVM_AUC=",round(svm_auc,6),sep=''))
```
Area Under the Curve (AUC) for ROC Curve is ```{svm_auc}```.
Given the AUC, SVM has outperformed all the classifiers we have evaluated so far.

SVM by far has some attractive properties.
Adding new observations to the left or right of the computed support vectors
does not change the support vectors and our inferences are stable as new
data arrives.
SVM needs to carry the vectors nothing more -- not the dataset as we did in
the case of kNN.
SVMs are most explainable determined by directly by the data.

The decision boundary for a SVM is linear (recall it is straight line)
SVMs are most known for its customizability using kernel trick.
Here let us look at an example from ESL book which demonstrates
the power of SVM when the boundary is not linear.
Download this data, 
https://web.stanford.edu/~hastie/ElemStatLearn/datasets/ESL.mixture.rda
from the ESL book site.

```{r}
load(url("https://web.stanford.edu/~hastie/ElemStatLearn/datasets/ESL.mixture.rda"))
plot(ESL.mixture$x,col=ESL.mixture$y+3)
data_svm<-data.frame(y=factor(ESL.mixture$y),ESL.mixture$x)

 data.svm<-svm(factor(y)~.,data=data_svm,scale=FALSE,kernel="radial",cost=15)
 plot(data.svm,data_svm)

```

Clearly this is not a linear boundary but SVM does this even though the data is not
separable. The performance of this model may be computed just as we have been
doing and that is left as an exercise, not for credit though.

For Software Engineers, kernel technique, just passing in 
method as a parameter, is not new.