### 2020-11-11 DATA622/Test1.rmd
More work on Test 1

---
title: "Test 1"
author: "Mike Silva"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Use the dataset you used for HW-1 (Blue/Black)
# 
# (A) Run Bagging (ipred package)   
# 
#   -- sample with replacement
# 
#   -- estimate metrics for a model
# 
#   -- repeat as many times as specied and report the average
# 
# (B) Run LOOCV (jacknife) for the same dataset
# 
# --- iterate over all points
# 
#      -- keep one observation as test
# 
#     -- train using the rest of the observations
# 
#     -- determine test metrics
# 
#     -- aggregate the test metrics
# 
# end of loop
# 
# find the average of the test metric(s)
# 
# Compare (A), (B) above with the results you obtained in HW-1  and write 3 sentences explaining the
# 
# observed difference.
# 
# You can implement it in java/spark, python or R.
# 
# Submit RMD or PDF and the python/R scripts.
# 
# --
# 
# i would recommend that you implement it from scratch but you may use any package. It is not a requirement.
# 
# bagging is available in ipred package in R.
```
This test is an extension of the work done in Homework 1.  I will be extending it with bagging and leave one out cross-validation (LOOCV), and adding in the results into the evaluation of the algorithms.

## Preliminaries

### Load Packages

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(kableExtra)
library(caret)
library(e1071)
library(class)
library(ROCR)
library(stringr)
```

### Read in the data

```{r}
df <- read.csv("https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA622/HW1.csv", stringsAsFactors = TRUE)
```

## Data Prep

Since we will be using KNN we will need to normalize `X`.  In order to get a fair comparison, we will used the same processed data to see how the algorithms preform.

```{r}
normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}

adjusted_df <- df
adjusted_df$X <- normalize(adjusted_df$X)
```

We also need to transform `Y` into dummy variables.

```{r}
adjusted_df <- adjusted_df %>%
  mutate(Y_a = ifelse(Y == "a", 1, 0),
         Y_b = ifelse(Y == "b", 1, 0),
         Y_c = ifelse(Y == "c", 1, 0),
         Y_d = ifelse(Y == "d", 1, 0),
         Y_e = ifelse(Y == "e", 1, 0)) %>%
  select(-Y)
```

In order to evaluate the ability of the of the algorithms to generalize, the data needs to be divided into training and test sets.  Since there is only `r nrow(df)` observations, I will be reserving 20% for the evaluation.

```{r}
set.seed(123)
in_training_set <- createDataPartition(adjusted_df$label, p = 0.8, list = FALSE, times = 1)
training_df <- adjusted_df[in_training_set,]
test_df  <- adjusted_df[-in_training_set,]
# These subsets will help with training and evaluation
training_df_without_label <- training_df %>% select(-label)
test_df_without_label  <- test_df %>% select(-label)
training_df_label <- training_df$label
test_df_label <- test_df$label
```

## Helper Functions

I will use the following helper functions.  This first one takes a logistic regression model and returns predictions for the `data` data.frame

```{r}
get_lr_yhat <- function(lr_model, data, label_col_name, threshold = 0.5){
  data_levels <- levels(data[[label_col_name]])
  cols_to_keep <- label_col_name != names(data)
  data <- data[,cols_to_keep]
  lr_yhat <- predict(lr_model, data, type = "response")
  lr_yhat <- as.factor(ifelse(lr_yhat <= threshold, data_levels[1], data_levels[2]))
  return(lr_yhat)
}
```

The second will take the the ground truth labels and predicted labels and create metrics for evaluation.

```{r}
evaluate_algo <- function(ground_truth, yhat, algo){
  cm <- confusionMatrix(table(ground_truth, yhat))
  cm_table <- cm$table
  tpr <- cm_table[[1]] / (cm_table[[1]] + cm_table[[4]])
  fnr <- 1 - tpr
  fpr <- cm_table[[3]] / (cm_table[[3]] + cm_table[[4]])
  tnr <- 1 - fpr
  accuracy <- cm$overall[[1]]
  for_auc <- prediction(c(yhat), ground_truth)
  auc <- performance(for_auc, "auc")
  auc <- auc@y.values[[1]]
  return(data.frame(Algo = algo, AUC = auc, ACCURACY = accuracy, TPR = tpr, FPR = fpr, TNR = tnr, FNR = fnr))
}
```

Now for the extensions for the test.  This third function will be used to create a bootstrap sample.  I will also make the assumption that we will use 100 boostrap samples in evaluating the performance of bagging.

```{r}
get_bootstrap_sample <- function(data, bootstrap_proportion = 0.6, sample_with_replacement = TRUE){
  n_bootstrap_observations <- round(nrow(data) * bootstrap_proportion, 0)
  return(data[sample(nrow(data), n_bootstrap_observations, replace = sample_with_replacement),])
}

n_bags <- 100
```

I will also need a way to aggregate/average the metrics

```{r}
average_metrics <- function(metrics, algo){
  avg_metrics <- metrics %>%
  select(-Algo) %>%
  colMeans(na.rm = TRUE) %>%
  t() %>%
  data.frame("Algo" = c(algo), .)
  return(avg_metrics)
}
```

## Baseline Models

First I will train the models following what was previously done in HW1 and collect the metrics on both the training and test data sets

### Logistic Regression

```{r, comment=NA}
lr_model <- glm(label ~ ., data = training_df, family = "binomial")
summary(lr_model)
```
#### Capacity to Learn

**The sample data are too small (`r nrow(training_df)`) to have any confidence in the Logistic Regression model, so I will be throwing this one out.  Bagging and LOOCV will not resolve this issue.**

```{r}
capacity_to_learn <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
```

#### Capacity to Generalize

**Again the sample data are too small to use to make meaningful inferences using Logistic Regression.  Bagging and LOOCV will not help and will be excluded from the metrics.**

```{r}
capacity_to_generalize <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
```

### Naive Bayes

I will repeat the same process for Naive Bayes.

#### Capacity to Learn

```{r}
nb_model <- naiveBayes(label ~ ., data = training_df)
training_nb_yhat <- predict(nb_model, training_df_without_label)
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(training_df_label, training_nb_yhat, "NB"))
```

#### Capacity to Generalize

```{r}
nb_yhat <- predict(nb_model, test_df_without_label)
capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, nb_yhat, "NB"))
```

### KNN

Finally I will repeat the process for KNN where k=3 and k=5

#### Capacity to Learn

```{r}
training_knn3_yhat <- knn(training_df_without_label, training_df_without_label, training_df_label, k = 3)
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(training_df_label, training_knn3_yhat, "KNN3"))

training_knn5_yhat <- knn(training_df_without_label, training_df_without_label, training_df_label, k = 5)
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(training_df_label, training_knn5_yhat, "KNN5"))
```

#### Capacity to Generalize

```{r}
knn3_yhat <- knn(training_df_without_label, test_df_without_label, training_df_label, k = 3)
capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, knn3_yhat, "KNN3"))

knn5_yhat <- knn(training_df_without_label, test_df_without_label, training_df_label, k = 5)
capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, knn5_yhat, "KNN5"))
```

## Models With Bagging

Now to try bagging the Naive Bayes, and KNN models.  I will use the models trained on the bootstrap samples to predict the test data and collect metrics on the capacity to learn and generalize, which will be aggregated.

```{r}
the_bags <- list()
for (i in 1:n_bags){
  bag_df <- get_bootstrap_sample(training_df)
  bag_df_without_label <- bag_df %>% select(-label)
  bag_df_label <- bag_df$label
  the_bags[[i]] <- bag_df

  # Train the Naive Bayes Model
  bag_nb_model <- naiveBayes(label ~ ., data = bag_df)
  training_bag_nb_yhat <- predict(bag_nb_model, bag_df_without_label)
  ## Evaluate the capacity to learn
  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_nb_yhat, paste("NB Bag", i))
  if(exists("nb_bag_capacity_to_learn")){
    nb_bag_capacity_to_learn <- rbind(nb_bag_capacity_to_learn, bag_capacity_to_learn)
  } else {
    nb_bag_capacity_to_learn <- bag_capacity_to_learn
  }
  ## Evaluate the capacity to generalize
  bag_nb_test_yhat <- predict(bag_nb_model, test_df_without_label)
  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_nb_test_yhat, paste("NB Bag", i))
  if(exists("nb_bag_capacity_to_generalize")){
    nb_bag_capacity_to_generalize <- rbind(nb_bag_capacity_to_generalize, bag_capacity_to_generalize)
  } else {
    nb_bag_capacity_to_generalize <- bag_capacity_to_generalize
  }

  # Train the KNN3 Model
  ## Evaluate the capacity to learn
  training_bag_knn3_yhat <- knn(bag_df_without_label, bag_df_without_label, bag_df_label, k = 3)
  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_knn3_yhat, paste("KNN3 Bag", i))
  if(exists("knn3_bag_capacity_to_learn")){
    knn3_bag_capacity_to_learn <- rbind(knn3_bag_capacity_to_learn, bag_capacity_to_learn)
  } else {
    knn3_bag_capacity_to_learn <- bag_capacity_to_learn
  }
  ## Evaluate the capacity to generalize
  bag_knn3_test_yhat <- knn(bag_df_without_label, test_df_without_label, bag_df_label, k = 3)
  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_knn3_test_yhat, paste("KNN3 with Bagging", i))
  if(exists("knn3_bag_capacity_to_generalize")){
    knn3_bag_capacity_to_generalize <- rbind(knn3_bag_capacity_to_generalize, bag_capacity_to_generalize)
  } else {
    knn3_bag_capacity_to_generalize <- bag_capacity_to_generalize
  }

  # Train the KNN5 Model
  ## Evaluate the capacity to learn
  training_bag_knn5_yhat <- knn(bag_df_without_label, bag_df_without_label, bag_df_label, k = 5)
  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_knn5_yhat, paste("KNN5 with Bagging", i))
  if(exists("knn5_bag_capacity_to_learn")){
    knn5_bag_capacity_to_learn <- rbind(knn5_bag_capacity_to_learn, bag_capacity_to_learn)
  } else {
    knn5_bag_capacity_to_learn <- bag_capacity_to_learn
  }
  ## Evaluate the capacity to generalize
  bag_knn5_test_yhat <- knn(bag_df_without_label, test_df_without_label, bag_df_label, k = 5)
  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_knn5_test_yhat, paste("KNN5 with Bagging", i))
  if(exists("knn5_bag_capacity_to_generalize")){
    knn5_bag_capacity_to_generalize <- rbind(knn5_bag_capacity_to_generalize, bag_capacity_to_generalize)
  } else {
    knn5_bag_capacity_to_generalize <- bag_capacity_to_generalize
  }

}
```

Now to average the metrics:

```{r}
# Naive Bayes
capacity_to_learn <- rbind(capacity_to_learn, average_metrics(nb_bag_capacity_to_learn, "NB with Bagging"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(nb_bag_capacity_to_generalize, "NB with Bagging"))

# KNN3
capacity_to_learn <- rbind(capacity_to_learn, average_metrics(knn3_bag_capacity_to_learn, "KNN3 with Bagging"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn3_bag_capacity_to_generalize, "KNN3 with Bagging"))

# KNN5
capacity_to_learn <- rbind(capacity_to_learn, average_metrics(knn5_bag_capacity_to_learn, "KNN5 with Bagging"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn5_bag_capacity_to_generalize, "KNN5 with Bagging"))
```

## Models With LOOCV

Now to examine how the models preform using LOOCV

```{r}
training_loocv_nb_yhat <- c()
training_loocv_knn3_yhat <- c()
training_loocv_knn5_yhat <- c()
loocv_test_label <- c()
loocv_dfs <- list()

for (i in 1:nrow(training_df)){
  # Leave One Out
  loocv_test <- training_df[i,]
  loocv_test_without_label <- loocv_test %>% select(-label)
  loocv_test_label <- c(loocv_test_label, loocv_test$label)
  loocv_training_df <- training_df[-c(i),]
  loocv_training_df_without_label <- loocv_training_df %>% select(-label)
  loocv_training_df_label <- loocv_training_df$label
  loocv_dfs[[i]] <- loocv_training_df

  # Train the Naive Bayes Model
  loocv_nb_model <- naiveBayes(label ~ ., data = loocv_training_df)
  ## Evaluate the capacity to learn
  ### Note: This will be done once we have gone over the whole training data set
  training_loocv_nb_yhat <- c(training_loocv_nb_yhat, predict(loocv_nb_model, loocv_test_without_label))
  ## Evaluate the capacity to generalize
  loocv_nb_test_yhat <- predict(loocv_nb_model, test_df_without_label)
  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_nb_test_yhat, paste("NB with LOOCV", i))
  if(exists("nb_loocv_capacity_to_generalize")){
    nb_loocv_capacity_to_generalize <- rbind(nb_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
  } else {
    nb_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
  }

  # Train the KNN3 Model
  ## Evaluate the capacity to learn
  ### Note: This will be done once we have gone over the whole training data set
  training_loocv_knn3_yhat <- c(training_loocv_knn3_yhat, knn(loocv_training_df_without_label, loocv_test_without_label, loocv_training_df_label, k = 3))
  ## Evaluate the capacity to generalize
  loocv_knn3_test_yhat <- knn(loocv_training_df_without_label, test_df_without_label, loocv_training_df_label, k = 3)
  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_knn3_test_yhat, paste("KNN3 with LOOCV", i))
  if(exists("knn3_loocv_capacity_to_generalize")){
    knn3_loocv_capacity_to_generalize <- rbind(knn3_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
  } else {
    knn3_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
  }

  # Train the KNN5 Model
  ## Evaluate the capacity to learn
  ### Note: This will be done once we have gone over the whole training data set
  training_loocv_knn5_yhat <- c(training_loocv_knn5_yhat, knn(loocv_training_df_without_label, loocv_test_without_label, loocv_training_df_label, k = 5))
  ## Evaluate the capacity to generalize
  loocv_knn5_test_yhat <- knn(loocv_training_df_without_label, test_df_without_label, loocv_training_df_label, k = 5)
  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_knn5_test_yhat, paste("KNN5 with LOOCV", i))
  if(exists("knn5_loocv_capacity_to_generalize")){
    knn5_loocv_capacity_to_generalize <- rbind(knn5_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
  } else {
    knn5_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
  }

}
```

Now to average the metrics:

```{r}
# Naive Bayes
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_nb_yhat, "NB with LOOCV"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(nb_loocv_capacity_to_generalize, "NB with LOOCV"))

# KNN3
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_knn3_yhat, "KNN3 with LOOCV"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn3_loocv_capacity_to_generalize, "KNN3 with LOOCV"))

# KNN5
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_knn5_yhat, "KNN5 with LOOCV"))
capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn5_loocv_capacity_to_generalize, "KNN5 with LOOCV"))
```

## Summary and Analysis

We have evaluated `r nrow(na.omit(capacity_to_learn))` algorithm's capacity to learn and generalize.  The following tables summarize their performance:

**Table 1. Capacity to Learn**
```{r, echo=FALSE}
capacity_to_learn %>% arrange(Algo) %>% kable() %>%  kable_styling()
```

**Table 2. Capacity to Generalize**
```{r, echo=FALSE}
capacity_to_generalize %>% arrange(Algo) %>% kable() %>% kable_styling()
```

### Discussion

There are `r nrow(training_df)` rows in the training set.  When doing LOOCV the model is trained on `r nrow(training_df) - 1` rows of data.  This subsample is going to be very similar to the full training set.  The bagging on the other hand will be significantly different.  There are `r nrow(bag_df)` rows of data and since it is sampled with replacement a same observation could be selected multiple times.  Consequently the underlying data is different.  We can see this by looking at the share of records that are labeled "BLACK."

```{r}
get_share_black <- function(df){
  df %>% filter(label == "BLACK") %>% nrow() / nrow(df)
}
```

The share of the training data set that is labeled black is `r get_share_black(training_df)`. Let's compare this to the shares in the bags.

```{r, comment=NA}
share_black <- c()
for(i in 1:n_bags){
  share_black <- c(share_black, get_share_black(the_bags[[i]]))
}

summary(share_black)
```

You can see that the share of observations that are black range from `r round(min(share_black) * 100, 0)`% to `r round(max(share_black) * 100, 0)`%.  This is very different than what we observe with the LOOCV data set:

```{r, comment=NA}
share_black <- c()
for(i in 1:nrow(training_df)){
  share_black <- c(share_black, get_share_black(loocv_dfs[[i]]))
}

summary(share_black)
```

This is a lot closer to the full training set.  Consequently a model trained on the bootstrap sample is likely to have different results if it is a weak learner, than the full training set.  Some of the models will do a better job at generalizing than others due to the composition of the data.  This can be seen by looking at a metric like the accuracy of the model.  The following images represent the mean accuracy on the test dataset as a dashed vertical line.  For the bagging and LOOCV runs the distribution of the accuracy is also shown:

```{r}
vline_df <- capacity_to_generalize %>% filter(str_detect(Algo, "NB"))
ggplot() + geom_vline(data=vline_df, aes(xintercept=ACCURACY, color=Algo), linetype="dashed") + geom_density(aes(x=ACCURACY, color="NB with Bagging"), data=nb_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="NB with LOOCV"), data=nb_loocv_capacity_to_generalize) + ggtitle("Naive Bayes Capacity to Generalize") + scale_color_brewer(palette = "Set1")
```

```{r}
vline_df <- capacity_to_generalize %>% filter(str_detect(Algo, "KNN3"))
ggplot() + geom_vline(data=vline_df, aes(xintercept=ACCURACY, color=Algo), linetype="dashed") + geom_density(aes(x=ACCURACY, color="KNN3 with Bagging"), data=knn3_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="KNN3 with LOOCV"), data=knn3_loocv_capacity_to_generalize) + ggtitle("KNN3 Capacity to Generalize") + scale_color_brewer(palette = "Set1")
```

```{r}
vline_df <- capacity_to_generalize %>% filter(str_detect(Algo, "KNN5"))
ggplot() + geom_vline(data=vline_df, aes(xintercept=ACCURACY, color=Algo), linetype="dashed") + geom_density(aes(x=ACCURACY, color="KNN5 with Bagging"), data=knn5_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="KNN5 with LOOCV"), data=knn5_loocv_capacity_to_generalize) + ggtitle("KNN5 Capacity to Generalize") + scale_color_brewer(palette = "Set1")
```

Now I hasten to note that all of these figures hare very specific to the initial split in the data which is controlled by the random number seed.  So the statements I will make at this point are specific to this run.  If I would have used another random number seed the results could be vastly different.

When I completed HW-1 I noted the the KNN family of models preformed the best.  KNN3 with bagging out preformed KNN3 on the whole training data.  As previously discussed it is due to the variation in the bootstrap samples.  When it comes to the ability to generalize KNN5 with LOOCV out preformed KNN5 and KNN3 with bagging out preformed KNN3.  The accuracy increases above the "baseline" algorithm.  NB improved the ability to generalize through both bagging and LOOCV.  This suggests that the performance of these weak learners is highly dependent on the data, but when averaged together the group outperforms the single model.

Again these results are highly dependent on the underling data, and additional runs should be done in order to ensure the conclusions are real and not soley based on idiosyncrasies in the data.  The idea is inline with what has been taught in this class.