### 2020-11-13 DATA698/08 - Report Visualizations.rmd
More work on project

---
title: "08 - Report Viz"
author: "Mike Silva"
date: "Fall 2020"
output: html_document
---

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

yolov3_results <- read.csv("report/yolov3_results.csv") %>%
  rowwise() %>%
  mutate(Epoch = as.numeric(strsplit(Epoch, "/")[[1]][1])+1)

yolov5_results <- read.csv("report/yolov5_results.csv") %>%
  rowwise() %>%
  mutate(Epoch = as.numeric(strsplit(Epoch, "/")[[1]][1])+1)

```

```{r}
yolo_results <- pivot_longer(yolov3_results, GIoU:Val.classification, names_to = "Measure")
yolo_results <- pivot_longer(yolov5_results, GIoU:Val.classification, names_to = "Measure") %>%
  rbind(yolo_results)
```

```{r}
viz_data <- yolov3_results %>%
  mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
ggplot(viz_data, aes(x=Epoch, y=mAP.0.5, color=Model, group=Model)) + geom_line() + scale_color_brewer(palette = "Set1")
```

```{r}
viz_data <- yolov5_results %>%
  mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
ggplot(viz_data, aes(x=Epoch, y=mAP.0.5, color=Model, group=Model)) + geom_line() + scale_color_brewer(palette = "Set1")
```

```{r}
for(m in unique(yolo_results$Measure)){
  viz_data <- yolo_results %>%
    filter(Measure == m) %>%
    mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
    p <- ggplot(viz_data, aes(x=Epoch, y=value, color=Model, group=Model)) + 
      geom_line() +
      theme(axis.title.y = element_blank()) +
      ggtitle(m)
    print(p)
}
```