### 2020-11-25 DATA698/08 - Report Visualizations.rmd
Updated visualization generating script

---
title: "08 - Report Viz"
author: "Mike Silva"
date: "Fall 2020"
output: html_document
---

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

yolov3_results <- read.csv("report/yolov3_results.csv") %>%
  rowwise() %>%
  mutate(Epoch = as.numeric(strsplit(Epoch, "/")[[1]][1])+1)

yolov5_results <- read.csv("report/yolov5_results.csv") %>%
  rowwise() %>%
  mutate(Epoch = as.numeric(strsplit(Epoch, "/")[[1]][1])+1)
```

## Images by Month Chart

```{r}
months_df <- read.csv("report/all_month_counts.csv") %>%
  mutate(Images = ifelse(Images == 0, NA, Images)) %>%
  mutate(Plant = ifelse(Plant.Type == "Poison Ivy", "A", "ERROR")) %>%
  mutate(Plant = ifelse(Plant.Type == "Box Elder", "B", Plant)) %>%
  mutate(Plant = ifelse(Plant.Type == "Bramble", "C", Plant)) %>%
  mutate(Plant = ifelse(Plant.Type == "Virginia Creeper", "D", Plant)) %>%
  mutate(Plant = as.factor(Plant)) %>%
  mutate(Plant = recode(Plant, A = "Poison Ivy", B = "Box Elder", C = "Bramble", D = "Virginia Creeper")) %>%
  select(-Plant.Type)

ggplot(months_df, aes(x=Month, y=Images, color=Plant, fill=Plant)) +
  geom_col() +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") + 
  facet_wrap(~Plant) +
  theme(legend.position = "none") +
  #xlim(1, 12)
  scale_x_discrete(name ="Month",  limits=c("1","2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
```

## Model Performance Charts

```{r}
viz_data <- yolov3_results %>% select(model, n_images, Epoch, mAP.0.5)
viz_data <- yolov5_results %>% select(model, n_images, Epoch, mAP.0.5) %>% rbind(viz_data) 
viz_data$model <- as.factor(viz_data$model)
viz_data$n_images <- as.factor(viz_data$n_images)

ggplot(viz_data) + 
  geom_line(aes(x=Epoch, y=mAP.0.5, color=n_images, group=n_images)) + 
  facet_wrap(~model) + 
  scale_color_brewer(palette = "Set1") +
  labs(color = "Number of Images") +
  ylab("mAP@0.5") +
  theme(legend.position = "bottom")
```

## YOLOv3

```{r}
viz_data <- yolov3_results %>% select(n_images, Epoch, obj) %>% mutate(set = "Training") %>% rename(measure = obj)
viz_data <- yolov3_results %>% select(n_images, Epoch, Val.objectness) %>% mutate(set = "Test") %>% rename(measure = Val.objectness) %>% rbind(viz_data) 
viz_data$n_images <- as.factor(viz_data$n_images)

ggplot(viz_data) + 
  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
  facet_wrap(~n_images) + 
  scale_color_brewer(palette = "Set1") +
  labs(color = "Data Set") +
  theme(legend.position = "bottom") +
  ylab("Object Detection Error")
```

```{r}
viz_data <- yolov3_results %>% select(n_images, Epoch, cls) %>% mutate(set = "Training") %>% rename(measure = cls)
viz_data <- yolov3_results %>% select(n_images, Epoch, Val.classification) %>% mutate(set = "Test") %>% rename(measure = Val.classification) %>% rbind(viz_data) 
viz_data$n_images <- as.factor(viz_data$n_images)

ggplot(viz_data) + 
  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
  facet_wrap(~n_images) + 
  scale_color_brewer(palette = "Set1") +
  labs(color = "Data Set") +
  theme(legend.position = "bottom") +
  ylab("Classification Error")
```

## YOLOv5

```{r}
viz_data <- yolov5_results %>% select(n_images, Epoch, obj) %>% mutate(set = "Training") %>% rename(measure = obj)
viz_data <- yolov5_results %>% select(n_images, Epoch, Val.objectness) %>% mutate(set = "Test") %>% rename(measure = Val.objectness) %>% rbind(viz_data) 
viz_data$n_images <- as.factor(viz_data$n_images)

ggplot(viz_data) + 
  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
  facet_wrap(~n_images) + 
  scale_color_brewer(palette = "Set1") +
  labs(color = "Data Set") +
  theme(legend.position = "bottom") +
  ylab("Object Detection Error")
```

```{r}
viz_data <- yolov5_results %>% select(n_images, Epoch, cls) %>% mutate(set = "Training") %>% rename(measure = cls)
viz_data <- yolov5_results %>% select(n_images, Epoch, Val.classification) %>% mutate(set = "Test") %>% rename(measure = Val.classification) %>% rbind(viz_data) 
viz_data$n_images <- as.factor(viz_data$n_images)

ggplot(viz_data) + 
  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
  facet_wrap(~n_images) + 
  scale_color_brewer(palette = "Set1") +
  labs(color = "Data Set") +
  theme(legend.position = "bottom") +
  ylab("Classification Error")
```

```{r}
yolo_results <- pivot_longer(yolov3_results, GIoU:Val.classification, names_to = "Measure")
yolo_results <- pivot_longer(yolov5_results, GIoU:Val.classification, names_to = "Measure") %>%
  rbind(yolo_results)

for(m in unique(yolo_results$Measure)){
  viz_data <- yolo_results %>%
    filter(Measure == m) %>%
    mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
    p <- ggplot(viz_data, aes(x=Epoch, y=value, color=Model, group=Model)) + 
      geom_line() +
      theme(axis.title.y = element_blank()) +
      ggtitle(m)
    print(p)
}
```

```{r}
ggplot(yolov3_results, aes(R, P, color=as.factor(paste(model, n_images)))) + geom_point()
```