### 2020-12-15 DATA622/HW2.rmd
Initial commit of HW 2

---
title: "HW2"
author: "Mike Silva"
output:
  pdf_document: default
  html_document: default
---
```{r, message=FALSE, warning=FALSE}
library(caret)
library(tictoc)
library(dplyr)
library(tidyr)
library(ROCR)
library(e1071)
library(kableExtra)
library(randomForest)
```

For this homework I will use Logistic Regression and SVM on the heart data set.  Instructions for the homework are in *italics*

```{r}
path <- "heart.csv"
df <- read.csv(path, stringsAsFactors = FALSE)
# Clean up the first column
names(df)[[1]] <- "age"
```

## PART-A

### Base Metrics

*Do a 80/20 split and determine the Accuracy, AUC and as many metrics as returned by the Caret
package (confusionMatrix) Call this the base_metric. Note down as best as you can development
(engineering) cost as well as computing cost(elapsed time).*

```{r}
set.seed(43)
in_train <- createDataPartition(df$target, p = 0.8, list = FALSE, times = 1)
train <- df[in_train, ]
test <- df[-in_train, ]
```

These are some general purpose utility functions to get the evaluations metrics

```{r}
get_cm_metrics <- function(pred, truth){
  cm <- confusionMatrix(table(pred, truth))
  metrics <- c(cm$overall["Accuracy"], cm$byClass)
  return(data.frame(as.list(metrics)))
}

get_auc <- function(pred, truth){
  temp <- prediction(pred, truth)
  prf <- performance(temp, "auc")
  return(prf@y.values[[1]])
}

get_elapsed <- function(tictoc){
  return(tictoc$toc - tictoc$tic)
}
```

#### Logistic Regression Classifier

It took me 20 minutes to develop this logistic regression pipeline with the evaluation metrics.  I also developed this logistic regression version of predict that will return the classes instead of the probabilities.

```{r}
lr_predict <- function(lr_model, df){
  lr_probs <- predict(lr_model, df, type="response")
  lr_yhat <- ifelse(lr_probs < 0.5, 0, 1)
  return(lr_yhat)
}

lr_engineering_cost <- 20
```

Now to train the baseline logistics regression classifier.

```{r}
tic("Logistic Regression Training")
glm_model <- glm(target~., data = df, family = "binomial")
summary_glm_model <- summary(glm_model)
coef_summary_glm_model <- coef(summary_glm_model)
# I only want statistically significant variables
lr_formula <- paste("target~", paste(row.names(coef_summary_glm_model[coef_summary_glm_model[, 4] < 0.05, ]), collapse = "+"), sep = "")
lr_model <- glm(lr_formula, data = df, family = "binomial")
lr_time <- toc()
lr_time <- get_elapsed(lr_time)
lr_yhat <- lr_predict(lr_model, test)
lr_auc <- get_auc(lr_yhat, test$target)
base_metric <- get_cm_metrics(lr_yhat, test$target) %>%
  mutate(AUC = lr_auc,
         model = "LR",
         engineering_cost = lr_engineering_cost,
         computing_cost = lr_time)
```

#### SVM Classifier

The SVM  classifier did not take as much engineering time since most of it was developed for the LR model.  Here's the SVM's helper functions

```{r}
svm_predict <- function(svm_model, df){
  svm_probs <- predict(svm_model, df)
  svm_yhat <- ifelse(svm_probs < 0.5, 0, 1)
  return(svm_yhat)
}

svm_engineering_cost <- 5
```

And now we train...

```{r}
tic("SVM Training")
svm_model <- svm(target~., data = df, probability = TRUE)
svm_time <- toc()
svm_time <- get_elapsed(svm_time)
svm_yhat <- svm_predict(svm_model, test)
svm_auc <- get_auc(svm_yhat, test$target)
base_metric <- get_cm_metrics(svm_yhat, test$target) %>%
  mutate(AUC = svm_auc,
         model = "SVM",
         engineering_cost = svm_engineering_cost,
         computing_cost = svm_time) %>%
  bind_rows(base_metric)
```

### K-Fold Cross Validated Metrics

*Start with the original dataset and set a seed (43). Then run a cross validation of 5 and 10 of the model
on the training set. Determine the same set of metrics and compare the cv_metrics with the
base_metric. Note down as best as you can development (engineering) cost as well as computing
cost(elapsed time).*

#### Logistic Regression

I have built this out to run both cross validations.  It took about 20 minutes to build which cost I split between the models.

```{r}
n <- nrow(df)
ks <- c(5, 10)
cv_lr_engineering_cost <- 20 / length(ks)

for (k in ks){
  set.seed(43)
  folds <- split(1:n, cut(1:n, quantile(1:n, probs = seq(0, 1, by = 1 / k))))
  ridx <- sample(1:nrow(df), nrow(df), replace = FALSE)

  tic(paste(k, "Fold CV LR Training"))

  k_fold_metrics <- do.call("rbind", lapply(folds, FUN = function(idx, data = df[ridx, ]) {
    cv_train <- data[-idx, ]
    cv_test <- data[idx, ]
    lr_model <- glm(lr_formula, data = cv_train, family = "binomial")
    lr_yhat <- lr_predict(lr_model, cv_test)
    lr_auc <- get_auc(lr_yhat, cv_test$target)
    temp <- get_cm_metrics(lr_yhat, cv_test$target) %>%
      mutate(AUC = lr_auc) 
    return(temp)
  }))

  cv_lr_time <- toc()
  cv_lr_time <- get_elapsed(cv_lr_time)

  k_fold_metrics <- k_fold_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = paste(k, "Fold LR"),
           engineering_cost = cv_lr_engineering_cost,
           computing_cost = cv_lr_time)
  if(exists("cv_metrics")){
    cv_metrics <- bind_rows(cv_metrics, k_fold_metrics)
  } else {
    cv_metrics <- k_fold_metrics
  }
}
```
#### SVM

Again the SVM benefited from the development of the workflow for the LR model.  I once again split the developmental costs between the models.

```{r}
cv_svm_engineering_cost <- 5 / length(ks)

for (k in ks){
  set.seed(43)
  folds <- split(1:n, cut(1:n, quantile(1:n, probs = seq(0, 1, by = 1 / k))))
  ridx <- sample(1:nrow(df), nrow(df), replace = FALSE)

  tic(paste(k, "Fold CV SVM Training"))

  k_fold_metrics <- do.call("rbind", lapply(folds, FUN = function(idx, data = df[ridx, ]) {
    cv_train <- data[-idx, ]
    cv_test <- data[idx, ]
    svm_model <- svm(target~., data = df, probability = TRUE)
    svm_yhat <- svm_predict(svm_model, cv_test)
    svm_auc <- get_auc(svm_yhat, cv_test$target)
    temp <- get_cm_metrics(svm_yhat, cv_test$target) %>%
      mutate(AUC = svm_auc) 
    return(temp)
  }))

  cv_svm_time <- toc()
  cv_svm_time <- get_elapsed(cv_svm_time)

  cv_metrics <- k_fold_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = paste(k, "Fold SVM"),
           engineering_cost = cv_svm_engineering_cost,
           computing_cost = cv_svm_time) %>%
    bind_rows(cv_metrics)
}
```

#### Comparison

The performance of the model increases with the number of folds.  This is said by looking at the AUC which is a good summary measure of the models performance.  The Base LR model has an AUC of about 0.81.  The K-5 CV LR model's AUC increases to 0.83 and K-10 CV LR increases to 0.84.  The same could be said with SVM where the AUC increases from a base of 0.88 to 0.913 and 0.914 for K-5 and 10 respectively.  Cross validation improves the model performance.  There is some computational cost.  The computational time on the SVM increased by a factor of 3 for K-5 CV and a factor of 6 (roughly) for K-10 CV.  I don't find the performance gain from the K-5 SMV to K-10 to be "worth it".  There is a bit of an uptick for logistic regression for a K-10 model in computational cost.  The performance metrics are summarized in the two tables below:

**Table 1. Base Metrics**

```{r, echo=FALSE}
base_metric %>% 
  kable() %>%
  kable_styling()
```

**Table 2. CV Metrics**

```{r, echo=FALSE}
cv_metrics %>% 
  kable() %>%
  kable_styling()
```

### Bootstrapped Metrics

*Start with the original dataset and set a seed (43) Then run a bootstrap of 200 resamples and compute
the same set of metrics and for each of the two classifiers build a three column table for each
experiment (base, bootstrap, cross-validated). Note down as best as you can development (engineering)
cost as well as computing cost(elapsed time).*

#### Bootstrapped Logistics Regression

First I will develop the bootstrap workflow for LR.  This is not the most efficient way of doing it but it works.

```{r}
set.seed(43)
n_runs <- 200
bs_lr_engineering_cost <- 10

tic("Bootstrap LR Training")

for(i in 1:n_runs){
  in_train <- sample(1:nrow(df), nrow(df), replace=T)
  train <- df[in_train, ]
  lr_model <- glm(lr_formula, data = train, family = "binomial")
  lr_yhat <- lr_predict(lr_model, df)
  lr_auc <- get_auc(lr_yhat, df$target)
  temp <- get_cm_metrics(lr_yhat, df$target) %>%
    mutate(AUC = lr_auc)
  if(exists("bs_lr_metrics")){
    bs_lr_metrics <- bind_rows(bs_lr_metrics, temp)
  } else {
    bs_lr_metrics <- temp
  }
}

bs_lr_time <- toc()
bs_lr_time <- get_elapsed(bs_lr_time)

bs_lr_metrics <- bs_lr_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = "Bootstrapped LR",
           engineering_cost = bs_lr_engineering_cost,
           computing_cost = bs_lr_time)
```

#### Bootstraped SVM

And here's the SVM which is basically the LR with some minor tweaks...

```{r}
set.seed(43)
bs_svm_engineering_cost <- 5

tic("Bootstrap SVM Training")

for(i in 1:n_runs){
  in_train <- sample(1:nrow(df), nrow(df), replace=T)
  train <- df[in_train, ]
  svm_model <- svm(target~., data = train, probability = TRUE)
  svm_yhat <- svm_predict(svm_model, df)
  svm_auc <- get_auc(svm_yhat, df$target)
  temp <- get_cm_metrics(svm_yhat, df$target) %>%
    mutate(AUC = svm_auc)
  if(exists("bs_svm_metrics")){
    bs_svm_metrics <- bind_rows(bs_svm_metrics, temp)
  } else {
    bs_svm_metrics <- temp
  }
}

bs_svm_time <- toc()
bs_svm_time <- get_elapsed(bs_svm_time)

bs_svm_metrics <- bs_svm_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = "Bootstrapped SVM",
           engineering_cost = bs_svm_engineering_cost,
           computing_cost = bs_svm_time)

bs_metrics <- bind_rows(bs_lr_metrics, bs_svm_metrics)
```

Here's how the bootstrapped models preformed:

**Table 3. Bootstrapped Metrics**

```{r, echo=FALSE}
bs_metrics %>% 
  kable() %>%
  kable_styling()
```

### Comparison of Experiments

Now to compare the base, CV and bootstrap results.  I will only be looking at the AUC.

**Table 4. Experiment Results**

```{r, echo=FALSE}
experiment_results <- base_metric %>% 
  select(model, AUC) %>%
  mutate(experiment = "Base")

experiment_results <- cv_metrics %>% 
  filter(model %in% c("10 Fold LR", "10 Fold SVM")) %>% 
  mutate(model = ifelse(model == "10 Fold LR", "LR", "SVM")) %>%
  select(model, AUC) %>%
  mutate(experiment = "10-Fold CV") %>%
  bind_rows(experiment_results)

experiment_results <- bs_metrics %>%
  mutate(model = ifelse(model == "Bootstrapped LR", "LR", "SVM")) %>%
  select(model, AUC) %>%
  mutate(experiment = "Bootstrapped") %>%
  bind_rows(experiment_results)

experiment_results %>%
  pivot_wider(id_cols = model, names_from = experiment, values_from = AUC) %>%
  select(model, Base, `10-Fold CV`, Bootstrapped) %>%
  kable() %>%
  kable_styling()
```

The Highest AUC is found in the cross validated models.  The bootstrapped models preformed a little better than the base models but that improvement came at a considerable cost:

**Table 5. Computational Cost**

```{r, echo=FALSE}
experiment_results <- base_metric %>% 
  select(model, computing_cost) %>%
  mutate(experiment = "Base")

experiment_results <- cv_metrics %>% 
  filter(model %in% c("10 Fold LR", "10 Fold SVM")) %>% 
  mutate(model = ifelse(model == "10 Fold LR", "LR", "SVM")) %>%
  select(model, computing_cost) %>%
  mutate(experiment = "10-Fold CV") %>%
  bind_rows(experiment_results)

experiment_results <- bs_metrics %>%
  mutate(model = ifelse(model == "Bootstrapped LR", "LR", "SVM")) %>%
  select(model, computing_cost) %>%
  mutate(experiment = "Bootstrapped") %>%
  bind_rows(experiment_results)

experiment_results %>%
  pivot_wider(id_cols = model, names_from = experiment, values_from = computing_cost) %>%
  select(model, Base, `10-Fold CV`, Bootstrapped) %>%
  kable() %>%
  kable_styling()
```

These results are not parallelized so there could be some decrease in the computational costs, but CV gives a lot of bang for the buck.

## PART-B

*For the same dataset, set seed (43) split 80/20.

Using randomForest grow three different forests varying the number of trees at least three times. Start
with seeding and fresh split for each forest. Note down as best as you can development (engineering)
cost as well as computing cost(elapsed time) for each run. And compare these results with the
experiment in Part A.*

```{r}
set.seed(43)
in_train <- createDataPartition(df$target, p = 0.8, list = FALSE, times = 1)
train <- df[in_train, ]
test <- df[-in_train, ]
train$target <- as.factor(train$target)
test$target <- as.factor(test$target)

rf_engineering_cost <- 15 / 3

for (n in c(100, 250, 1000)) {
  set.seed(43)
  tic(paste("Training RF ", n))
  rf_1000_model <- randomForest(target~., data = train, ntree = n)
  rf_time <- toc()
  rf_time <- get_elapsed(rf_time)
  rf_yhat <- predict(rf_1000_model, test)
  rf_auc <- get_auc(as.numeric(as.character(rf_yhat)), as.numeric(as.character(test$target)))
  temp <- get_cm_metrics(rf_yhat, test$target) %>%
    mutate(AUC = rf_auc,
         model = paste("RF", n),
         engineering_cost = rf_engineering_cost,
         computing_cost = rf_time)
  if(exists(("rf_metrics"))){
    rf_metrics <- bind_rows(rf_metrics, temp)
  } else {
    rf_metrics <- temp
  }
}
```

How did the random forests preform?  Table 6 below summarizes them.  The performance is not particularly strong.  The base SVM did a better job than the largest random forest.  Both bootstrapped models out preformed the random forests, and the 10 fold CV models did a lot better.

**Table 6. Random Forest Metrics**

```{r, echo=FALSE}
rf_metrics %>% 
  kable() %>%
  kable_styling()
```

## PART-C

*Include a summary of your findings. Which of the two methods bootstrap vs cv do you recommend to
your customer? And why? Be elaborate. Including computing costs, engineering costs and model
performance. Did you incorporate Pareto's maxim or the Razor and how did these two heuristics
influence your decision?*

In this exercise I developed two classifiers. One was a logistic regression model and the other a SVM.  I ran some experiments with these models to see if I could improve the model's performance.  The two experiments were to use cross validation and bootstrapping.  In looking at the accuracy of the model as measured using AUC, and if I had to recommend only one model I would recommend the 5-fold cross-validated SVM model.  It showed considerable improvement from the base model.  While the 10-fold SVM did a little better, the computational costs don't justify it.  I also examined 3 random forests as possible solutions but the cross validated SVMs  preformed better than the random forests.

This analysis applies Akum's razor by focusing on AUC instead of all the other metrics.  I could have just as easily focused on positive predictive value if that was more important to the client but I assumed the overall accuracy was the most important.  I applied sort of a Pareto analysis in recommending the K-5 SVM.  Yes the K-10 did better but in terms of computational costs it cost a lot more to squeeze out a little more performance.  It had hit the point of diminishing returns and so it was easier for me to recommend something that got the most bang for the buck.